### 2018-10-25 main.tex
Add files via upload

%&pdflatex
\documentclass[preprint,3p,review,authoryear,10pt]{elsarticle}

\usepackage{graphicx}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage[T1]{fontenc}
%\setlength{\parindent}{0in}
\usepackage{times}
\usepackage[british]{babel}
\usepackage{algorithm}
\usepackage{algorithmic} \renewcommand{\algorithmicrequire}{\textbf{Given:}} \renewcommand{\algorithmicensure}{\textbf{Returns:}}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mathtools}
\renewcommand \thesection{\arabic{section}}
\usepackage[caption=false]{subfig}
\usepackage[compact]{titlesec}
\usepackage{soul}
\usepackage{textcomp}
\usepackage[section]{placeins}
\usepackage{url}
\usepackage{xspace}
\usepackage{lineno}
\usepackage{cleveref}

%Added by Diarmid
\usepackage{siunitx}

% % new commands
\newcommand{\I}{\mathcal{I}}
\DeclareMathOperator*{\any}{any}
\DeclareMathOperator*{\all}{all}

\newcommand{\ddt}[1]{\frac{\partial #1}{\partial t}}
\newcommand{\ddx}[1]{\frac{\partial #1}{\partial x}}

%% for annotating paper for TODO notes etc.  Comment this out and
%% uncomment alternative for final version
\newcommand{\esf}[1]{\footnote{#1}\marginpar{\fbox{\thefootnote}}}
%\newcommand{\esf}[1]{}

\journal{Computers \& Chemical Engineering}

\begin{document}
\linenumbers
\begin{frontmatter}

\title{Posing Redox Flow Battery Scheduling Problems in Current-Density Terms for Improved Fidelity and Computability.}

\author[]{D. Roberts \corref{cor_author}}
\ead{droberts1@sheffield.ac.uk}
\author{S. Brown}

\address{Department of Chemical and Biological Engineering, The University of Sheffield, Sheffield S10 2TN}

\cortext[cor_author]{Corresponding author}

\begin{abstract}
Redox Flow Battery (RFB) systems are promising technologies for the multi-hour electrical energy storage that would be necessary in a future low carbon emission UK electrical grid based on wind and solar power. In order to perform techno-economic analyses (TEA) on these systems, it is necessary to calculate potential revenue for realistic applications, ideally under optimal operation (scheduling). In the present work, a novel schedule optimisation framework is introduced in which coulombic and voltaic components are separated, and the problem variablised in terms of current-density rather than power. This approach allows the expression of ohmic losses as a convex quadratic function of current-density in the objective function while maintaining all linear constraints. The resultant non-linear problem is more readily solvable than previous expressions in terms of power. It is demonstrated for an arbitrage application involving a vanadium RFB that the consideration of variable voltaic losses results in a significant improvement to the accuracy of revenue estimation when compared to the typical linear programming approach, where a fixed efficiency is assumed. The framework allows the removal of the assumption of a representative fixed open-cell voltage, which further improves revenue estimation, at the cost of computability. In a third case study, the framework allows a linear constraint on the working cell-voltage, resulting in current-density restriction at high state of charge (SOC) in order to avoid hydrogen and oxygen evolution. This feature would be relevant for the optimisation of Li-ion battery scheduling, where voltage is a very important consideration. The introduced framework is intended to assist technical optimisation of future RFB systems, for example in determining the optimum pump or cooling system size. The framework allows product optimisation in terms of revenue under optimal scheduling against a price signal (rather than metrics such as system cost, or lifetime cost per unit energy throughput as in previous works).

\end{abstract}

\begin{keyword}
Redox flow Battery \sep Optimisation \sep Non-linear Programming \sep Techno-Economic Analysis \sep Scheduling \sep VRFB  
%% keywords here, in the form: keyword \sep keyword

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

\end{keyword}

\end{frontmatter}

% \linenumbers

%% main text
\section{Introduction}
\label{sec:Intro}
The  growing deployment of intermittent renewable power sources such as wind and solar in industrialised nations in order to substitute fossil fuel power is a key driver of research into the development and applications of battery energy storage systems (BESS).

At present in the UK, the fluctuating output from these power sources is topped up to match demand primarily by varying the output of gas fired power plant \cite{DraxElectricInsights}. In order to meet the strict CO\textsubscript{2} emissions targets for the UK electrical supply by 2050 (and bearing in mind that nuclear power capacity is not predicted to grow until after 2030 \cite{FES2018}) it is probable that further multi-hour energy storage will be required to displace gas plant from the role of balancing.  Within the UK electrical market, the phasing out of gas power plants (most likely via increments to the carbon price) to be replaced by intermittent generation would lead to high electrical prices where demand exceeds supply, and low (or negative) pricing where supply exceeds demand.

At present, revenue from arbitrage on the wholesale electrical price is not in itself sufficient to justify BESS investment, and the deployment of grid scale BESS is dominated by projects that have won contracts for short term stability services tendered by National Grid. For example, \SI{500}{\mega\watt} of BESS capacity won contracts for the T-4 2020/21 capacity market (CM) auction, and \SI{200}{\mega\watt} of BESS capacity won enhanced frequency response (EFR) contracts. The majority of successful BESS applicants for the EFR auction also have CM contracts.

However, due to concerns that the average storage duration (i.e. energy to power ratio) of the BESS contracted for CM is shorter than the \SI{2}{\hour} that a stress event may last for, National Grid recently applied derating factors according to the energy to power ratio. Hence, a system rated at \SI{1}{\mega\watt}/\SI{2}{\mega\watt\hour} would only receive capacity payments for \SI{0.68}{\mega\watt} in the 2018/19 T-1 auction, and \SI{0.65}{\mega\watt} in the 2021/22 T-4 auction. The derating factor is relaxed to 0.96 for systems rated for 4+ hours of discharge \cite{NGDeratingFactors}.

This development disfavours the power-centric Li-ion systems and favours redox flow battery (RFB) systems, where the incremental cost of energy capacity is low, and systems are typically specified with a 3-6 \si{\hour} energy to power ratio.

In order to quantify the economic merits of RFB systems in comparison to other BESS, it is necessary to perform techno-economic analyses (TEA) for realistic applications. TEA should involve optimisation of the operation of a BESS to maximise revenue (or minimise costs). This should be done at various energy and power ratings (as these are decoupled in a non-hybrid RFB) to find the global optimum, as described by \cite{Oudalov2007}. The optimisation of operation requires determining the schedule, that is, the power output of the battery in each sub-period that maximises benefit without violating energy conservation.

For the optimisation scheduling, linear programming is a commonly used method. In the case of arbitrage, the linear programming (LP) approach involves the maximisation of the objective function for revenue, defined by:

\begin{equation}
\label{eqn:Lit_Review_LP_Obj_Func}
R = \tau \sum_{t} (P_{Dis,t} - P_{Chg,t}) p_t
\end{equation}

where $P_{Dis,t}$ and $P_{Chg,t}$ represent the bounded variables discharge and charge power between the RFB and the rest of the system in period $t$, $\tau$ the time-step parameter (length of period $t$) and $p_t$ the parameter price of electrical energy in period $t$.

The state of charge (SOC) of the BESS is then defined by:

\begin{equation}
\label{eqn: Lit_Review_LP_SOC_Constraint_Example}
\begin{gathered}
SOC_t = SOC_{t-1} + \frac{\tau P_{Chg,t} \eta_{Chg}}{E_{BESS}} - \frac{\tau P_{Dis,t}}{E_{BESS}\eta_{Dis}} \\
\end{gathered}
\end{equation}

where $E_{BESS}$ is the parameter energy capacity and $\eta_{Chg}$ and $\eta_{Dis}$ are the charge and discharge energy efficiency parameters respectively. The SOC is at all times constrained between a maximum of 1 and a minimum of 0 (or a narrower window). Linear programming is an efficient problem formulation, as the optimum value for each variable will either lie at the maximum or minimum permitted value, making a search of the interior unnecessary.

Variants of the above formulation have been used to calculate the revenue or avoided costs due to RFB systems in various applications.

\cite{Hu2010} optimised the schedule of a VRFB and a poly-sulphide bromine RFB in order to maximise arbitrage revenue on the Danish day ahead electrical spot market. As the optimisation was to be performed daily, the SOC was constrained to return to 0 at the end of each \SI{24}{\hour} period.

\cite{Johnston2015} used a LP formulation to optimise the scheduling of a VRFB installed at a wind farm. The objective was to maximise revenue on the wholsesale electrical market while satisfying the frequency response obligations (constraints) of the wind installation. \cite{Vaca2017} looked at a similar application but compared the payback period for VRFB and supercapacitor installations with a hybrid system. In both studies, a cost of degradation in \pound per unit energy throughput was applied as a coefficient to ESS power variables in the objective function. This approach discourages the performance of arbitrage below a certain threshold price differential.

A full system schedule optimisation problem was posed by \cite{Chen2012} to assess the benefit of a VRFB in a microgrid with a load, PV, wind and dispatchable micro-generation. When the micro-grid was isolated, the objective was to minimise the start-up and running costs of micro-generators used to back up the renewable power output. Integer variables were required to factor start-up costs (MILP) and additional constraints were required to guarantee system reliability.

A potentially significant simplification in all of the above studies is the assumption of a constant efficiency. In reality, the efficiency of any BESS is a function of the charge/discharge power (among other variables, such as SOC). In the above formulation, if efficiency were defined as a  function of power, it would be necessary to place this function on the denominator of the SOC increment term in the discharge expression in \Cref{eqn: Lit_Review_LP_SOC_Constraint_Example}. This complicates the problem as the constraint is now non-linear and non-quadratic. \cite{Nguyen2014} report a highly detailed nonlinear model for the charge and discharge efficiency of a vanadium RFB. In a subsequent article (\cite{Nguyen2015}) this function was incorporated in the SOC expression of a schedule optimisation problem in the aforementioned fashion. Due to the complication of the SOC constraint, the authors employ a search approach with dynamic programming to break the problem down. This is more computationally intensive than the LP method, but the authors state that an optimum schedule may be found in an acceptable time-scale as the small number of variables and the presence of constraints limit the search space.

\cite{Sarker2017} also incorporated the dependency of efficiency on power in a scheduling problem (although applied to a Li-ion battery rather than a RFB). Quadratic terms are employed to relate losses to C-rate of charge and discharge. The presence of this quadratic function on the denominator of the discharge SOC increment term is avoided by redefining the discharge power variable as upstream of losses (rather than the power delivered to the grid). The discharge efficiency function is then moved from the SOC expression and multiplied by the discharge power variable in the objective function. The charge efficiency is left in the SOC expression (redefining the charge power variable in the same way would require dividing it by the quadratic charge efficiency term in the objective function). The problem therefore becomes convex (quadratic), but nonlinear in both objective function and constraint. The authors then employ a linear approximation technique, and combine the efficiency treatment with a piece-wise linear treatment of degradation as a function of power, giving an overall mixed integer-linear programming problem (MILP).

The present work is motivated by the desire to simplify the expression of variable efficiency within a RFB schedule optimisation problem. This is done by the separation of the coulombic and voltaic components of power, so that each may be dealt with in the most appropriate manner.

\section{Model formulation}
\label{sec:Model_Formulation}

\subsection{LP Problem in Current-Density Terms}
\label{Model_Formulation_LP_Reference_Current_Density_Terms}
The  basic LP scheduling problem described in \cref{eqn:Lit_Review_LP_Obj_Func,eqn: Lit_Review_LP_SOC_Constraint_Example} was re-posed in terms of current-density in order to serve as a benchmark.  The first step was to replace the power variable(s) with charge and discharge current-density variables. Current-density was chosen over current as the former is a scale-independent metric commonly used in the RFB literature. The use of current density makes multiplication by stack area necessary to obtain absolute current, and multiplication by open-cell-volage (OCV) necessary to obtain power. Separate efficiencies representing voltaic and ohmic losses must now be specified, rather than the single energy efficiency term in \cref{eqn: Lit_Review_LP_SOC_Constraint_Example}. As current-density is defined based on the current at the RFB terminals, coulombic losses are dealt with upstream in \cref{eqn: Method_SOC_Tracker}. It is however necessary to incorporate voltaic losses in the objective function.  The objective function for revenue in period $Y$ (to be maximised) is defined by:

\begin{equation}
\label{eqn: Linear_Schedule_Objective_Function_Current_as_Variable}
R_{LP} = \frac{A.\tau OCV_{50\%}}{1\times 10^{6}}\sum_{t}p_{t}(I_{D, t}\sqrt{\bar\eta_{V}}(1 - l_{BOP}) - \frac{I_{C,t}}{\sqrt{\bar\eta_{V}}(1 - l_{BOP})}) \; \forall t \in Y
\end{equation}

where $I_{C,t}$ and $I_{D,t}$ are independent variables representing charge and discharge current density (\si{\ampere\per\square\meter}) respectively. $A$ is the stack area parameter (\si{\square\meter}), $\tau$ the timestep (\si{\hour}), $OCV_{50\%}$ the open cell voltage at 50\% state of charge (SOC) and $p_t$ the indexed electrical price parameter (\SI[sticky-per, bracket-unit-denominator = false]{}[\pounds]{\per\mega\watt\hour}). $\bar\eta_{V}$ is the representative round-trip voltaic efficiency. The  placement of the additional fractional balance of plant losses parameter $l_{BOP}$ reflects the assumption that balance of plant power consumption is primarily due to electrolyte pumping, and that pumping power is proportional to current. The validity of this assumption is discussed in \cref{analysis_of_assumptions}.

The function describing SOC given in \cref{eqn: Lit_Review_LP_SOC_Constraint_Example} is now expressable in purely coulombic terms. This is done by replacing the power variables with current-density variables, the energy efficiency parameters with a coulombic efficiency parameter, and the energy capacity of the RFB with a coulombic capacity. The resultant function for the SOC at the end of sub-period $t$ is defined by:

\begin{equation}
\label{eqn: Method_SOC_Tracker}
SOC_t = SOC_{t-1} + \frac{A.\tau}{1000C}(I_{C,t}\sqrt{\eta_C} - \frac{I_{D,t}}{\sqrt{\eta_C}}) \; \forall t \in Y
\end{equation}

where $C$ is the coulombic capacity of the RFB (\si{\ampere\hour}). It is assumed that the coulombic efficiency $\eta_C$ does not vary with SOC or current-density, and that coulombic losses for charge and discharge are symmetrical. This assumption is discussed in \cref{analysis_of_assumptions}.

The following constraints are applied to the list of SOC values generated by the above expression. Firstly the constraint on SOC range is formalised by:

\begin{equation}
\label{eqn: Method_Constraint_SOC_Range}
SOC_{min} \leq SOC_t \leq SOC_{max} \; \forall t \in Y
\end{equation}

where $SOC_{min}$ and $SOC_{max}$ are the minimum and maximum permitted SOC parameters.

So that the optimisation may be legitimately performed on consecutive periods of historical data (i.e. maintaining an energy balance) the SOC is required to return to 50\% at the end of each period. This constraint is formalised by:

\begin{equation}
\label{eqn: Method_Constraint_E_Conservation}
SOC_0 = SOC_n = 0.5
\end{equation}
where $n$ is the final sub-period in $Y$.

Finally, the current-density in and out of the RFB is constrained by:

\begin{equation}
\label{eqn: Method_Constraint_Ia}
I_{min} \leq I_{D,t}, \; I_{C,t} \leq I_{max} \; \forall t \in Y
\end{equation}

where $I_{min}$ and $I_{max}$ are the minimum and maximum permitted current-densities respectively.

\subsection{NLP treatment of voltaic losses}
\label{Model_Formulation_NLP_treatment_voltaic_losses}
A major simplification in the LP problem in terms of power is the assumption of a constant energy efficiency parameter. This assumption also applies to the LP problem in current-density terms introduced in \cref{Model_Formulation_LP_Reference_Current_Density_Terms}, although separate voltaic and coulombic efficiencies parameters are called for. In reality, ohmic (and pseudo-ohmic) over-potential is a linear function of current-density . The faradaic over-potential, which occurs when the circuit is closed, may be considered constant with respect to current-density (CITE article I used in CR). \cref{eqn: Linear_Schedule_Objective_Function_Current_as_Variable} was adapted to include these losses, by subtracting/adding the faradaic over-potential from/to the OCV during discharge/charge, and including the ohmic losses in the form of the classical $P_{Loss} = I^2R$ equation. This results in a polynomial objective function for revenue in period $Y$ defined by:

\begin{equation}
\label{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable}
R_{NLP} = \frac{A.\tau}{1\times 10^{6}}\sum_{t}p_{t}(I_{D, t}(OCV_{50\%} - V_a)(1-l_{BOP}) - \frac{I_{C, t}(OCV_{50\%} + V_a)}{1-l_{BOP}} - (I_{D,t}^{2} + I_{C,t}^{2})ASR) \; \forall t \in Y
\end{equation}

where $V_a$ and $ASR$ are scalar parameters representing faradaic over-potential (\si{\volt}) and area specific resistance (\si{\ohm\square\meter}). The placement of $V_a$ and $ASR$ reflects an assumption of symmetry in voltaic losses between charge and discharge (discussed further in \cref{analysis_of_assumptions}). The objective function is subject to the same constraints as described in \cref{Model_Formulation_LP_Reference_Current_Density_Terms}.

Although non-linear, \cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable} the objective function is convex (as it requires maximising a negative quadratic), hence a gradient based algorithmic solver is guaranteed to find the optimum from any starting point.

Under the constant coulombic efficiency assumption \cref{eqn: Method_SOC_Tracker} is linear, and so by extension are the constraints in \cref{eqn: Method_Constraint_SOC_Range,eqn: Method_Constraint_E_Conservation}.  The overall formulation hence comprises a nonlinear objective function subject to linear constraints. This is an improvement over the formulations of losses expressed in power efficiency terms by \cite{Nguyen2015,Sarker2017}, both of which use non-linear SOC constraint expressions. According to \cite{Hart2017}

\textit{"In the context of nonlinear models, those with nonlinear objective functions are typically easier to optimize than models with nonlinear constraint expressions"}

The formulation also benefits from a consistent definition of the current-density variables in terms of flow between the terminals and the external system, unlike in \cite{Sarker2017}.

The NLP problem is based on an assumption of constant coulombic efficiency, which is discussed further in \cref{analysis_of_assumptions}.

\subsection{Functional Treatment of OCV}
\label{Model_Formulation_Functional_Treatment_OCV}
A simplification in the NLP problem introduced in \cref{Model_Formulation_NLP_treatment_voltaic_losses} is the use of a fixed OCV, ($OCV_{50\%}$) where OCV is in fact a function of SOC. If the OCV is linear with respect to SOC in the permitted SOC range (which, in the case of the system described by \cite{Reed2016} it is), and the RFB transfers as much energy below 50\% SOC as above it, then this assumption may be reasonable.
However, using the current-density formulation, it is possible to functionalise the relationship between OCV and SOC. The relationship between the average OCV and the average SOC in sub-period t is approximated by:

\begin{equation}
\label{eqn: OCV_vs_SOC}
OCV_t = a\frac{SOC_t + SOC_{t-1}}{2} + b 
\end{equation}

where $a$ and $b$ are system specific parameters.

Replacing $OCV_{50\%}$ with $OCV_t$ in \cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable} and expanding $SOC_t$ and $SOC_{t-1}$ by \cref{eqn: Method_SOC_Tracker} results in an objective function (to be maximised) defined by:

\begin{equation}
    \label{eqn: NLP_true_OCV_long_form}
    \begin{split}
R_{NLP} = \frac{A.\tau}{1\times 10^{6}}\sum_{t}p_{t}(\\
(1-l_{BOP})I_{D, t}(a\frac{SOC_{t-1} + SOC_{t-2} + \frac{A.\tau}{1000C}(I_{C,t}\sqrt{\eta_C} - \frac{I_{D,t}}{\sqrt{\eta_C}} +  I_{C,t-1}\sqrt{\eta_C} - \frac{I_{D,t-1}}{\sqrt{\eta_C}})}{2} + b - V_a)\\ - \frac{1}{1-l_{BOP}}(I_{C, t}(a\frac{SOC_{t-1} +  SOC_{t-2} + \frac{A.\tau}{1000C}(I_{C,t}\sqrt{\eta_C} - \frac{I_{D,t}}{\sqrt{\eta_C}} + I_{C,t-1}\sqrt{\eta_C} - \frac{I_{D,t-1}}{\sqrt{\eta_C}})}{2} + b + V_a))\\ 
- (I_{D,t}^{2} + I_{C,t}^{2})ASR) \; \forall t \in Y
\end{split}
\end{equation}

The above function is non-convex due to the presence of bivariate quadratic terms  (e.g. $-I_{D,t}I_{C,t-1}$). Hence the problem matrix is not positive definite. However, by minor adjustment of the diagonal of the $Q$ matrix it is possible to obtain a positive semi-definite matrix. In the case of the Gurobi solver, a default diagonal adjustment of 1e-6 is allowed. If the required adjustment would be greater than this value, an error message is returned.

\subsection{Cell Voltage Constraint}
\label{Model_Formulation_Constraint_Cell_Voltage}
It has so far been assumed that the RFB may accept the maximum charge current-density right up to the maximum permitted SOC. In reality this may result in the cell-voltage being too-high (overcharging, discussed further in \cref{Results_Cell_Voltage_Constraint}. An alternative charging approach (common for Li-ion batteries where high cell voltages pose safety concerns) is to switch to a constant voltage taper charge at high SOC, where the current declines until a cut-off value is reached. In the case of an RFB, the fixed losses described above would likely make taper charging uneconomical (see the low efficiency at low current-density reported by \cite{Nguyen2014}). Rather than globally adjust the maximum permitted charge current-density or SOC (which may lead to underestimation of revenue), the objective function in \cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable} is made subject to a new linear constraint in terms of cell-voltage, defined by:

    \begin{equation}
\label{eqn: Cell_Voltage_Constraint}
OCV_t + V_a + I_{C,t}ASR \leq V_{max} \; \forall t \in Y
\end{equation}

Where $V_{max}$ is the maximum permitted cell-voltage. As $OCV_t$ is an averaged value of SOC in the sub-period $t$ (\cref{eqn: OCV_vs_SOC}), the accuracy of the model should improve as $\tau$ is reduced. A similar constraint could be implemented for discharge if required.

\subsection{RFB Specification}
\label{VRFB_specification}
The RFB is modelled here as a single cell. As the power output of a hypothetical RFB may be increased at the cost of efficiency, it is necessary to specify a rated efficiency when determining the stack area required for, say, a \SI{1}{\kilo\watt} system. This is formalised by:

\begin{equation}
\label{eqn: Rated_efficiency}
\eta_{DC} = \eta_{V,rated}\eta_{C}(1-l_{BOP})^2 = \eta_{rated}
\end{equation}

Where $\eta_V$ and $\eta_C$ are the round trip voltaic and coulombic efficiencies respectively. It is assumed that the one-way fractional balance of plant losses ($l_{BOP}$) are primarily due to electrolyte pumping, and that the pump runs on DC power.

The stack area (\si{\square\meter}) is calculated by:

\begin{equation}
\label{eqn: Simple_RFB_Cost_Model_Stack_Area}
A = \frac{P_{rated}}{I_{rated}V_0\sqrt{\eta_V}(1 - l_{BOP})}
\end{equation}

where $P_{rated}$ is the system power rating (\si{\watt}) and $I_{rated}$ is the current density (\si{\ampere\per\square\metre}) corresponding experimentally to coulombic and voltaic efficiencies that satisfy \Cref{eqn: Rated_efficiency}.

The coulombic capacity of the VRFB, required for the tracking of SOC in \Cref{eqn: Method_SOC_Tracker}, is calculated by:

\begin{equation}
\label{eqn: coulombic_capacity}
C = \frac{AI_{rated}r_{e:p}}{\sqrt{\eta_C}}
\end{equation}

Where $r_{e:p}$ is the specified energy to power ratio of the system. The presence of $\eta_C$ ensures there is sufficient chemical energy in the tank to cover coulombic losses upon discharge (square root after \cite{Darling2014}).

\section{Case Study}
\label{Case Study}

\subsection{The Application}
\label{Results_The_Application}
Pure price arbitrage on the day ahead N2EX electrical market was chosen as a case study, as the availability of future price information makes deterministic schedule optimisation relevant. Schedule optimisation was performed on 24h periods, with $\tau$ of 1h. The indexed price $p_t$ was parametrised using data obtained from the NordPool website \cite{NordPool2016}.

\subsubsection{The RFB}
\label{Model_Formulation_VRFB_Params}
\cite{Reed2016} studied various electrode flow architectures on a \SI{1}{\kilo\watt} VRFB system based on a \SI{50}{\micro\meter} Nafion NR212 membrane. In this case study, performance data were taken from the test of the highest performing stack architecture (IDD2s) at an electrolyte flow rate of 400 \si{\ml\per\minute} (per cell). This setup was tested at three current-densities: 1600, 2400 and 3200 \si{\ampere\per\square\meter}. At each test condition efficiency data were obtained as integrals over a cycle between 15\% and 85\% SOC.

In the parametrisation of the VRFB for following case studies $P_{rated}$ was set at \SI{1}{\kilo\watt}, the energy to power capacity ratio $r_{e:p}$ was set at 4 and $\eta_{rated}$ was set at 75\%. $L_{BOP}$ was set at 0.02, based on a conservative interpretation of the pumping power data reported for a very similar system by \cite{Kim2013}. A coulombic efficiency ($\eta_C$) of 0.975 was derived as the average of the values reported by \cite{Reed2016} at the three current-densities tested, which only varied between 0.974 and 0.976.  For the chosen system  the overall energy efficiency requirement in \cref{eqn: Rated_efficiency} is satisfied when the voltaic efficiency $\eta_{V,rated}$ is 0.801, which corresponds (by linear interpolation) to a current-density ($I_{rated}$) of \SI{2188}{\ampere\per\square\meter}. $OCV_{50\%}$ was taken from \cite{Kim2011}, where the OCV of a chloride electrolyte was reported at 20\%, 50\% and 80\% SOC. Although a different electrolyte (based on a mixture of sulphuric acid and hydrochloric acid) was used in \cite{Reed2016}, OCV data have not been published.

\cref{tab:VRFB_Sizing_Params} summarises the parameter values used to represent the VRFB.

\begin{table}[!pht]
\captionsetup{font=normalsize}
\centering
\caption{Parameters describing VRFB in sizing process}
\label{tab:VRFB_Sizing_Params}
\begin{tabular}{ccccccc}
\hline
Param &  Value & Units &  Source \\
\hline
$P_{rated}$ & 1 & \si{\kilo\watt} & Arb.\\
$r_{e:p}$ & 4 & - & \multirow{2}{*}{Typical}\\
$\eta_{rated}$ & 0.75 & - \\
\hline
$\eta_{V,rated}$ & 0.801 & - &\multirow{3}{*}{\cite{Reed2016}}\\
$\eta_C$ & 0.975 & -\\
$I_{rated}$ & 2188 & \si{\ampere\per\square\meter} \\
\hline
$OCV_{50\%}$ & 1.46 & \si{\volt} & \cite{Kim2011}\\
$l_{BOP}$ & 0.02 & - & \cite{Kim2013}\\

\hline
\end{tabular}
\end{table}

\subsection{Implementation}
\label{Results_Implementation}
Each optimisation problem was posed using PYOMO within a Python loop for repeat solutions on 24h sets of data. The Gurobi solver was called using the SolverFactory function. The script was run on a virtual machine (\cite{VirtualBox}) running Linux Mint, which was allocated 1024 MB RAM.

\subsection{NLP Treatment of Voltaic Losses}
\label{Results_NLP_Treatment_Voltaic_losses}
In the first case study, the performance of the NLP formulation with functionalised voltaic losses (\cref{Model_Formulation_NLP_treatment_voltaic_losses}) was assessed by comparison to the LP formulation described in \cref{Model_Formulation_LP_Reference_Current_Density_Terms}. The LP objective function in \cref{eqn: Linear_Schedule_Objective_Function_Current_as_Variable} requires the assumption of a constant voltaic efficiency term $\bar\eta_{V}$. A value of 0.842 was applied, as this is the voltaic efficiency at the mid-point of the permitted current-density range (\SI{1600}{\ampere\per\square\meter}). The NLP formulation in \cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable} requires values for area specific resistance ($ASR$) and Faradaic potential ($V_a$) in order to parametrise voltaic losses in the objective function. These were derived by multiplying $OCV_{50\%}$ by $1-\sqrt{\eta_V}$ to obtain an over-potential $\delta_v$ at the three current-density points in the experimental data (assuming losses are symmetrical in charge and discharge). $ASR$ was obtained as the gradient of the linear regression on this plot, and $V_a$ as the intercept at zero current-density.

\cref{fig:Va_ASR_Derivation} shows the regression of $\delta_V$ against current density for the data taken from \cite{Reed2016}.

\begin{figure}[!ht]
\centering
\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./ASR_deriv.pdf}
\caption{Derivation of $ASR$ (gradient) and $V_a$ (intercept) for IDD2s VRFB embodiment reported by \cite{Reed2016} at 400 \si{\ml\per\minute}.}
\label{fig:Va_ASR_Derivation}
\end{figure}

The optimisation of scheduling was performed for each day in 2017 using both the NLP formulation and the LP benchmark. \Cref{fig:IDD2s_opt_schedules} shows the optimal hourly schedule determined by each method for the 22\textsuperscript{nd} of August (this date was chosen because the resultant revenue is the median of the range obtained across the year).

\begin{figure}[!ht]
\centering
\subfloat[][]{
\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./IDD2s_schedules.pdf}
\label{fig:IDD2s_opt_schedules}
}
\subfloat[][]{
\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./IDD2s_schedules_SOC.pdf}
\label{fig:IDD2s_opt_schedules}
}
\caption{(a) Optimal VRFB schedules derived for the 22\textsuperscript{nd} August 2017 using LP and NLP optimisation formulations (discharge plotted as negative). (b) corresponding SOC values.}
\end{figure}

Both the LP and NLP solutions involve charging the VRFB when the electrical price is low (e.g. between 05:00 and 07:00) and discharging it when the price is high (e.g. 10:00 to 13:00). The LP optimal profile involves discharging or charging the system at high current-density in the four hours with the most extreme prices. The NLP optimal schedule also targets the most extreme periods, but spreads the charging/discharging out. This is the consequence of factoring variable ohmic losses in the objective function -   high current-density is penalised by higher losses. Under the LP formulation, the VRFB is predicted to generate \pounds 0.068, whereas the NLP formulation predicts a revenue of \pounds 0.063. In reality however, the voltaic losses that would occur under the LP schedule would reduce the revenue to \pounds 0.051. Across 2017, LP scheduling would have resulted in \pounds 22.61 (against prediction of \pounds 33.16) and NLP scheduling would have yielded \pounds 27.99.

In a TEA of this VRFB, use of the LP formulation would therefore lead to meaningful errors in calculation of revenue (in this case an overestimation). Use of the NLP formulation to determine scheduling in a practical application would result in significantly higher revenue than would the LP formulation. It would be possible to set a representative voltaic efficiency ($\bar\eta_V$) in the LP formulation such that the estimate of revenue would be closer to that of the NLP formulation, but this value cannot be known \textit{a priori}.

Solving \cref{eqn: Linear_Schedule_Objective_Function_Current_as_Variable} for each day in 2017 takes \SI{129}{\second} which is a minimal increase over the \SI{127}{\second} required to solve the LP objective function in \cref{eqn: Linear_Schedule_Objective_Function_Current_as_Variable}.

\subsection{NLP with Functional OCV}
\label{Results_NLP_Functional_OCV}
In the second case study, the impact of expressing OCV as a function of SOC (and hence of the current-density variables in preceding sub-periods) was investigated. Parameters $a$ and $b$ representing the linear relationship between OCV and SOC in \cref{eqn: NLP_true_OCV_long_form} were set at 0.2667 and 1.333, after linear regression on data published by \cite{Kim2011}.

\cref{fig:OCV_fixed_dynamic_schedules} shows a comparison of the optimal schedules obtained using the functionalised OCV formulation and the previous formulation (\cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable}) where OCV is fixed at the 50\% SOC value ($OCV_{50\%}$). \cref{fig:OCV} shows the OCV in the former solution.

\begin{figure}[!ht]
\centering
\subfloat[][]{
\label{fig:OCV_fixed_dynamic_schedules}
\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./OCV_fixed_dynamic_schedules.pdf}

}
\subfloat[][]{
\label{fig:OCV}
\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./OCV_fixed_dynamic_OCV.pdf}
}
\caption{(a) Optimal VRFB schedules derived for the 22\textsuperscript{nd} August 2017 using NLP formulations where OCV is fixed ($OCV_{50\%}$) and functionalised. (b) Optimal values of OCV where it is functionalised.}
\end{figure}

The performance of arbitrage between 16:00 and 20:00 is notably reduced when OCV is functionalised. In this period there are small fluctuations in price from which the RFB derived revenue when a fixed OCV was assumed. However the global optimum requires that the RFB be at a low SOC in these hours, and hence at a low OCV. The low OCV means that a higher current-density would be required to match the power output in the fixed OCV case, and hence higher voltaic losses would be incurred, eroding the revenue. On this date the impact on revenue is low; the optimal schedule with consideration of functional OCV results in a revenue of \pounds 0.062 compared to \pounds 0.063 for the base case. Across 2017 the formulation with functional OCV returns \pounds 26.40 compared to \pounds 27.81 for the fixed OCV case. The assumption of a fixed OCV (parametrised at 50\% SOC) therefore results in a small but meaningful overestimation of revenue in this application. Including functionalised OCV in the objective function increases the solve time - the 2017 data were processed in \SI{217}{\second} compared to \SI{129}{\second} for the convex objective function in \cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable}.

The non-convex objective function in \cref{eqn: NLP_true_OCV_long_form} makes the problem less robust: in the 2017 data there were two days in which the price parameter set yielded problem matrices that required greater diagonal adjustment than the default maximum permitted value of 1e-6. Solutions for the two days were obtained by increasing the Gurobi Parameter "PSDTol" to 1.7e-6. For certain applications this approach may become impractical.

\subsection{Cell-Voltage Constraint}
\label{Results_Cell_Voltage_Constraint}
In the NLP formulations given in \cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable,eqn: NLP_true_OCV_long_form} it is assumed that the RFB may be charged at maximum current-density at the maximum SOC. This was the charging strategy applied in \cite{Reed2016}, where the cell-voltage was allowed to reach \SI{1.85}{\volt} at the highest current-density, corresponding to an approximate SOC of 0.85. Overcharging may lead to oxygen evolution at the cathode and hydrogen evolution at the anode. This results in reduced coulombic efficiency and  (more importantly) the evolved gases can cause a range of problems as described by \cite{Kear2012}.  The particular cell-voltage at which gas evolution becomes problematic will vary from system to system (as it depends on the particular electrolyte chemistry). In this case study $V_{max}$ in \cref{eqn: Cell_Voltage_Constraint} was set at 1.65, just below the voltage at which \cite{Wei2017} reported evolution of macroscopic bubbles of hydrogen and oxygen (\SI{1.70}{\volt}).

Maximising \cref{eqn: NLP_true_OCV_long_form} subject to \cref{eqn: Cell_Voltage_Constraint} for the 2017 data resulted in \pounds 26.07 compared to \pounds 26.40 when the cell voltage was not constrained. The difference is small because high current-density is already discouraged by the functionalised voltaic losses in \cref{eqn: NLP_true_OCV_long_form}. \cref{fig:V_max_schedules} shows the optimal schedules with and without constraint for 5\textsuperscript{th} February 2017, where the difference in revenue was greatest (\pounds 0.018 vs. \pounds 0.020). \cref{fig:V_max_voltages} shows the corresponding cell-voltage profiles.

\begin{figure}[!ht]
\centering
\subfloat[][]{
\label{fig:V_max_schedules}
\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./V_max_schedules.pdf}
}
\subfloat[][]{
\label{fig:V_max_voltages}
\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./V_max_voltages.pdf}
}
\caption{(a) Optimal VRFB schedules derived for the 5\textsuperscript{th} February 2017 by maximising \cref{eqn: NLP_true_OCV_long_form} with no cell-voltage constraint, and subject to the cell voltage constraint given in \cref{eqn: Cell_Voltage_Constraint}. (b)Corresponding cell-voltage profiles.}
\end{figure}

In the unconstrained case the RFB charges at the maximum current-density of \SI{3200}{\ampere\per\square\meter} in hour 19 in order to capitalise on the unusually low electrical price, resulting in a cell-voltage of \SI{1.71}{\volt}. In the constrained case, the current-density is restricted to \SI{2213}{\ampere\per\square\meter} in the same period, and the RFB must charge more (and discharge less) in other periods to satisfy \cref{eqn: Method_Constraint_E_Conservation}. The impact of the constraint would be more severe if the variations in price data were greater, or the maximum voltage set at a lower value.

The application of the linear constraint on cell-voltage increases the time to process the 2017 data from \SI{217}{\second} to \SI{232}{\second}.

\subsection{Analysis of Assumptions}
\label{analysis_of_assumptions}
A number of assumptions have been made in the preceding NLP formulations in current-density terms:

\begin{enumerate}

    \item It is assumed that both coulombic efficiency and the voltaic loss parameters are invariant with SOC and symmetrical (i.e. the same for charge and discharge). Regarding the first assumption, experimental work by \cite{Nguyen2014} showed that the voltaic discharge efficiency is lower at 70\% SOC than at 35\% SOC, but that the difference decreases as power increase (data are only reported up to \SI{1.8}{\kilo\watt} for a \SI{5}{\kilo\watt} system). Regarding the latter, the same authors provide experimental evidence that the voltaic efficiency of the particular VRFB is higher for charge than discharge. The latter issue could be dealt with in the present formulation (if experimental data were made available) by including applying separate $ASR$, $V_a$ and $\eta_{C}$ terms for charge and discharge. The former issue would be more difficult to deal with, and a first step would be to perform a sensitivity study based on relevant system data.

    \item Coulombic efficiency is assumed to be constant with current-density, i.e. the coulombic losses are proportional to the useful current-density. This assumption fits the data reported by \cite{Reed2016}, but the behaviour below \SI{1600}{\ampere\per\square\meter} was not reported. In principle shunt current losses are a proportional to cell-voltage (as described by \cite{Xing2011}, and hence a function of SOC and over-voltage (via \cref{eqn: Cell_Voltage_Constraint}). The dependence of coulombic losses on current-density is therefore weaker than implied by the present assumption. As a first approximation, this function could be simplified to a fixed term.

    \item Balance of plant losses are fixed as a proportion of current-density in \Cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable}, reflecting an assumption of a continuously variable pump. This neglects the increase in pressure drop with electrolyte flow (shown in \cite{Reed2016}). The formulation may be improved by removing the $(1-l_{BOP})$ terms and adding a quadratic function of current-density. Alternatively the pump may have a  fixed output and be sized such that the maximum charge/discharge current-density may be satisfied at the highest/lowest SOC without incurring a significant additional over-potential due to mass-transport limitations.

    \item The RFB is not thermally constrained. The VRFB based on a sulphate electroylte is liable to precipitation of  V\textsubscript{2}O\textsubscript{5} from the catholyte as temperature increases (\cite{Rahman2009,Vijayakumar2011}). Although the chloride and mixed acid electrolytes developed at PNNL are more stable (\cite{Li2011,Kim2011}), temperature remains an important consideration. The framework introduced here could easily be expanded to capture the thermal performance of the system. A thermal constraint could be put in place using a similar construct to the one describing SOC, where the increase of heat in the electrolyte in sub-period $t$ is equal to the ohmic losses ($(I_{D,t}^{2} + I_{C,t}^{2})ASR$). This would allow cost benefit analyses of active cooling versus restrictions on operation at different electrolyte concentrations.

\end{enumerate}

As discussed above, both the coulombic losses and the pumping power may be better represented as fixed terms. In this case, the efficiency of the system at low power input/output will be unfavourable (as demonstrated by \cite{Nguyen2014}). An issue with representing these losses as fixed terms, is that they would be present at all times in the schedule. In order to minimise parasitic losses, a binary "on/off" variable could be added to the formulation in order to allow idling.

\section{Conclusion} 
\label{sec:Conclusion}
A novel framework for RFB schedule optimisation has been introduced in which the voltaic and coulombic components of the power output/input are separated, with the problem variablised in terms of current-density. An arbitrage case-study is used to demonstrate the increased fidelity that is achievable using the framework.

The framework is first used to describe ohmic losses as a quadratic function of current-density in the objective function, hence maintaining a convex solution space and all linear constraints. The importance of inclusion of variable ohmic losses in the arbitrage case study if demonstrated. Optimal deterministic scheduling using the introduced formulation (NLP) resulted in 24\% greater revenue across 2017 when compared to a basic LP approach that assumes a constant representative voltaic efficiency. In the context of a TEA, failure to factor in the variable voltaic losses leads to an overestimation of revenue by 47\%. The improved NLP treatment of voltaic losses only requires a minimal increase in computing time when compared to the LP problem with fixed voltaic efficiency.

The framework also facilitates the expression of OCV as a function of SOC, rather than assuming a fixed average value. The non-convex objective function that results may be solved using Gurobi, although there is a penalty in computing time. For the arbitrage case study, it is demonstrated that assuming a fixed OCV results in a small but significant overestimation of revenue. This assumption should be checked for other applications.

The framework also allows for an additional linear constraint to be placed on upper cell-voltage as a function of SOC and over-potential. This constraint restricts the current-density at which the battery may be charged when the SOC is high, which is an important consideration for safety and battery lifetime. In the present case study, the impact of the chosen maximum voltage is low, as the previous factoring of voltaic losses already discourages high power input. However, in applications with greater price variation this constraint will become more important.

A number of improvements to the particular problems presented are proposed, in order to more faithfully represent coulombic losses, the parasitic consumption of power in electrolyte pumping, and to allow idling.

It is intended that the novel framework be used for techno-economic analysis and optimisation of various RFB chemistries at a system level (i.e. up to the AC-DC boundary), for example in the optimal specification of pump rating, or cooling system for a particular application. RFB systems have previously been the subject of technical optimisation, for example the study of the trade-offs between electrolyte concentration and membrane ASR (\cite{Weber2013} \emph{this should be weber et al.}), or shunt currents and pumping losses (\cite{Viswanathan2014}. In these works, a simple metric such as levelised cost of electrical energy under a simple daily cycle is used as the objective to minimise. The framework introduced here allows optimisation in terms of revenue under optimal scheduling for any application where a price signal exists.

Although the framework was conceived with RFB systems in mind, the easy application of a cell-voltage constraint would be useful for Li-ion scheduling, where low and high cell-voltages increase cell degradation, and the latter also poses a safety concern. In the case of Li-ion batteries, where coulombic losses are negligible, the present assumption of a fixed coulombic efficiency would be entirely appropriate. The introduced treatment of voltaic losses is an improvement over that of \cite{Sarker2017} as the constraints are all linear.

\section{Acknowledgements}

This work was supported by the UK Engineering and Physical Science Research Council via grant EP/L016818/1 which funds the Centre for Doctoral Training in Energy Storage and its Applications. Drax?

\section*{References}
\bibliographystyle{elsarticle-harv}
\bibliography{Elsevier_article.bib}

\end{document}

### 2018-10-29 Paper/main.tex
added new main.tex

--- 2018-10-25 main.tex
@@
 %% uncomment alternative for final version
-\newcommand{\esf}[1]{\footnote{#1}\marginpar{\fbox{\thefootnote}}}
+\newcommand{\sol}[1]{\footnote{#1}\marginpar{\fbox{\thefootnote}}}
 %\newcommand{\esf}[1]{}
@@
 
-\title{Posing Redox Flow Battery Scheduling Problems in Current-Density Terms for Improved Fidelity and Computability.}
+\title{The importance of fidelity in the scheduling of Redox Flow Batteries.}
 
-\author[]{D. Roberts \corref{cor_author}}
-\ead{droberts1@sheffield.ac.uk}
-\author{S. Brown}
+\author{D. Roberts }
+\author[]{S. Brown \corref{cor_author}}
+\ead{s.f.brown@sheffield.ac.uk}
 
@@
 \label{sec:Intro}
-The  growing deployment of intermittent renewable power sources such as wind and solar in industrialised nations in order to substitute fossil fuel power is a key driver of research into the development and applications of battery energy storage systems (BESS).
+The  growing deployment of intermittent renewable power sources \sol{you need numbers here to underline how important this is, i.e.
+what percentage of energy, around the world, is now renewable} such as wind and solar in industrialised nations in order to substitute fossil fuel power is a key driver of research into the development and applications of battery energy storage systems \sol{why?
+what does it do?
+the argument needs to highlight the properties that batteries have and why we think are important} (BESS).
+\sol{you need to pack in the references here, you mention research so we need references to know what they looked at...}
 
 At present in the UK, the fluctuating output from these power sources is topped up to match demand primarily by varying the output of gas fired power plant \cite{DraxElectricInsights}.
-In order to meet the strict CO\textsubscript{2} emissions targets for the UK electrical supply by 2050 (and bearing in mind that nuclear power capacity is not predicted to grow until after 2030 \cite{FES2018}) it is probable that further multi-hour energy storage will be required to displace gas plant from the role of balancing.
-Within the UK electrical market, the phasing out of gas power plants (most likely via increments to the carbon price) to be replaced by intermittent generation would lead to high electrical prices where demand exceeds supply, and low (or negative) pricing where supply exceeds demand.
+In order to meet the strict CO\textsubscript{2} emissions targets for the UK electrical supply by 2050, and bearing in mind that nuclear power capacity is not predicted to grow until after 2030 \cite{FES2018}, it is probable that further multi-hour energy storage will be required to displace gas plant from the role of balancing.
+Within the UK electrical market, the phasing out of gas power plants, likely via increments to the carbon price \sol{reference?}, to be replaced by intermittent generation would lead to high electrical prices where demand exceeds supply, and low (or negative) pricing where supply exceeds demand.
 
@@
 The majority of successful BESS applicants for the EFR auction also have CM contracts.
+\sol{This is a bit parochial, the same is regarding BESS elsewhere (i.e.
+the big Tesla abttery).
+We need to widen the story out to make the paper relevant outside the UK}
 
@@
 
-This development disfavours the power-centric Li-ion systems and favours redox flow battery (RFB) systems, where the incremental cost of energy capacity is low, and systems are typically specified with a 3-6 \si{\hour} energy to power ratio.
+This development disfavours the power-centric Li-ion systems that previously won the majority of EFR contracts (Cite please) and favours redox flow battery (RFB) systems, where the incremental cost of energy capacity is low, and systems are typically specified with a 3-6 \si{\hour} energy to power ratio.
 
@@
 
-For the optimisation scheduling, linear programming is a commonly used method.
-In the case of arbitrage, the linear programming (LP) approach involves the maximisation of the objective function for revenue, defined by:
+For the optimisation of arbitrage revenue, the linear programming (LP) approach typically involves the maximisation of the objective function for revenue.
+Linear programming is an efficient problem formulation, as the optimum value for each variable will either lie at the maximum or minimum of a permitted value, making a search of the interior unnecessary.
+
+Variants of the above formulation have been used to calculate the revenue or avoided costs due to RFB systems in various applications.
+For example, \cite{Hu2010} optimised the schedule of a VRFB and a poly-sulphide bromine RFB in order to maximise arbitrage revenue on the Danish day ahead electrical spot market.
+As the optimisation was to be performed daily, the SOC is constrained to return to 0 at the end of each \SI{24}{\hour} period.
+
+\cite{Johnston2015} used a LP formulation to optimise the scheduling of a VRFB installed at a wind farm.
+The objective was to maximise revenue on the wholsesale electrical market while satisfying the frequency response obligations (constraints) of the wind installation\sol{what did they find?} .
+\cite{Vaca2017} looked at a similar application but compared the payback period for VRFB and supercapacitor installations with a hybrid system.
+In both studies, a cost of degradation in \pound per unit energy throughput was applied as a coefficient to ESS power variables in the objective function.
+This approach discourages the performance of arbitrage below a certain threshold price differential \sol{why is this note worthy?}.
+
+A full system schedule optimisation problem was posed by \cite{Chen2012} to assess the benefit of a VRFB in a microgrid with a load, PV, wind and dispatchable micro-generation.
+When the micro-grid was isolated, the objective was to minimise the start-up and running costs of micro-generators used to back up the renewable power output.
+The authors used a mixed-integer LP (MILP) approach, where the innteger variables were used required to factor start-up costs, additional constraints were required to guarantee system reliability \sol{are these reliability constraints interesting?}.
+
+A significant simplification in all of the above studies is the assumption of a constant efficiency.
+In reality, the efficiency of any BESS is a function of the charge/discharge power (among other variables, such as SOC).
+In the above formulation, if efficiency were defined as a  function of power, it would be necessary to place this function on the denominator of the SOC increment term in the discharge expression in \Cref{eqn: Lit_Review_LP_SOC_Constraint_Example}.\sol{merge the previous and coming sentences to remove the reference to the equations} This complicates the problem as the constraint is now non-linear and non-quadratic.
+\cite{Nguyen2014} report a highly detailed nonlinear\sol{be consistent in you use of 'non-linear' or 'nonlinear'} model for the charge and discharge efficiency of a vanadium RFB.
+In a subsequent article (\cite{Nguyen2015}) this function was incorporated into the SOC expression of a schedule optimisation problem in the aforementioned fashion.
+Due to the complication of the SOC constraint, the authors employ a search approach with dynamic programming to break the problem down.
+This is more computationally intensive than the LP method, but the authors state that an optimum schedule may be found in an acceptable time-scale as the small number of variables and the presence of constraints limit the search space.
+
+\cite{Sarker2017} also incorporated the dependency of efficiency on power in a scheduling problem (although applied to a Li-ion battery rather than a RFB).
+Quadratic terms are employed to relate losses to C-rate of charge and discharge.
+The presence of this quadratic function on the denominator of the discharge SOC increment term is avoided by redefining the discharge power variable as upstream of losses (rather than the power delivered to the grid).
+The discharge efficiency function is then moved from the SOC expression and multiplied by the discharge power variable in the objective function.
+The charge efficiency is left in the SOC expression (redefining the charge power variable in the same way would require dividing it by the quadratic charge efficiency term in the objective function).
+The problem therefore becomes convex (quadratic), but nonlinear in both objective function and constraint.
+The authors then employ a linear approximation technique\sol{does this need a reference}, and combine the efficiency treatment with a piece-wise linear treatment of degradation as a function of power, giving an overall mixed integer-linear programming problem (MILP).
+
+In this work, the schedule problem is formulated in terms of current-density rather than power, which allows the separation of coulombic and voltaic efficiency components.
+It will be shown that important constraints on the problem may be posed more readily in this form.\sol{you do more than this, you formulate this problem 3 times with increasing levels of complexity then apply each of these to a case study to compare their performance.
+You need to say this and be sure to say that this is all novel, otherwise it's not clear why the reviewers should be bothered}
+
+This paper is organised as follows: Section \ref{sec:Model_Formulation} presents...\sol{complete this paragraph}
+
+\section{Model formulation}
+\label{sec:Model_Formulation}
+\sol{describe what you're going to do here...}
+\subsection{LP Problem in Current-Density Terms}
+\label{Model_Formulation_LP_Reference_Current_Density_Terms}
+The typical LP formulation for the scheduling problem uses an objective function which is formulated in terms of the power as \sol{give references}:
 
@@
 where $E_{BESS}$ is the parameter energy capacity and $\eta_{Chg}$ and $\eta_{Dis}$ are the charge and discharge energy efficiency parameters respectively.
-The SOC is at all times constrained between a maximum of 1 and a minimum of 0 (or a narrower window).
-Linear programming is an efficient problem formulation, as the optimum value for each variable will either lie at the maximum or minimum permitted value, making a search of the interior unnecessary.
+The SOC is at all times normally constrained to between a maximum of 1 and a minimum of 0, though a narrower window has also used \sol{reference}.
 
-Variants of the above formulation have been used to calculate the revenue or avoided costs due to RFB systems in various applications.
+To re-pose this in terms of current-density, the power variable(s) with charge and discharge current-density variables.\sol{can you describe the relations that you've used here}
 
-\cite{Hu2010} optimised the schedule of a VRFB and a poly-sulphide bromine RFB in order to maximise arbitrage revenue on the Danish day ahead electrical spot market.
-As the optimisation was to be performed daily, the SOC was constrained to return to 0 at the end of each \SI{24}{\hour} period.
-
-\cite{Johnston2015} used a LP formulation to optimise the scheduling of a VRFB installed at a wind farm.
-The objective was to maximise revenue on the wholsesale electrical market while satisfying the frequency response obligations (constraints) of the wind installation.
-\cite{Vaca2017} looked at a similar application but compared the payback period for VRFB and supercapacitor installations with a hybrid system.
-In both studies, a cost of degradation in \pound per unit energy throughput was applied as a coefficient to ESS power variables in the objective function.
-This approach discourages the performance of arbitrage below a certain threshold price differential.
-
-A full system schedule optimisation problem was posed by \cite{Chen2012} to assess the benefit of a VRFB in a microgrid with a load, PV, wind and dispatchable micro-generation.
-When the micro-grid was isolated, the objective was to minimise the start-up and running costs of micro-generators used to back up the renewable power output.
-Integer variables were required to factor start-up costs (MILP) and additional constraints were required to guarantee system reliability.
-
-A potentially significant simplification in all of the above studies is the assumption of a constant efficiency.
-In reality, the efficiency of any BESS is a function of the charge/discharge power (among other variables, such as SOC).
-In the above formulation, if efficiency were defined as a  function of power, it would be necessary to place this function on the denominator of the SOC increment term in the discharge expression in \Cref{eqn: Lit_Review_LP_SOC_Constraint_Example}.
-This complicates the problem as the constraint is now non-linear and non-quadratic.
-\cite{Nguyen2014} report a highly detailed nonlinear model for the charge and discharge efficiency of a vanadium RFB.
-In a subsequent article (\cite{Nguyen2015}) this function was incorporated in the SOC expression of a schedule optimisation problem in the aforementioned fashion.
-Due to the complication of the SOC constraint, the authors employ a search approach with dynamic programming to break the problem down.
-This is more computationally intensive than the LP method, but the authors state that an optimum schedule may be found in an acceptable time-scale as the small number of variables and the presence of constraints limit the search space.
-
-\cite{Sarker2017} also incorporated the dependency of efficiency on power in a scheduling problem (although applied to a Li-ion battery rather than a RFB).
-Quadratic terms are employed to relate losses to C-rate of charge and discharge.
-The presence of this quadratic function on the denominator of the discharge SOC increment term is avoided by redefining the discharge power variable as upstream of losses (rather than the power delivered to the grid).
-The discharge efficiency function is then moved from the SOC expression and multiplied by the discharge power variable in the objective function.
-The charge efficiency is left in the SOC expression (redefining the charge power variable in the same way would require dividing it by the quadratic charge efficiency term in the objective function).
-The problem therefore becomes convex (quadratic), but nonlinear in both objective function and constraint.
-The authors then employ a linear approximation technique, and combine the efficiency treatment with a piece-wise linear treatment of degradation as a function of power, giving an overall mixed integer-linear programming problem (MILP).
-
-The present work is motivated by the desire to simplify the expression of variable efficiency within a RFB schedule optimisation problem.
-This is done by the separation of the coulombic and voltaic components of power, so that each may be dealt with in the most appropriate manner.
-
-\section{Model formulation}
-\label{sec:Model_Formulation}
-
-\subsection{LP Problem in Current-Density Terms}
-\label{Model_Formulation_LP_Reference_Current_Density_Terms}
-The  basic LP scheduling problem described in \cref{eqn:Lit_Review_LP_Obj_Func,eqn: Lit_Review_LP_SOC_Constraint_Example} was re-posed in terms of current-density in order to serve as a benchmark.
-The first step was to replace the power variable(s) with charge and discharge current-density variables.
 Current-density was chosen over current as the former is a scale-independent metric commonly used in the RFB literature.
-The use of current density makes multiplication by stack area necessary to obtain absolute current, and multiplication by open-cell-volage (OCV) necessary to obtain power.
+The use of current-density makes multiplication by stack area necessary to obtain absolute current, and multiplication by open-cell-volage (OCV) necessary to obtain power.
 Separate efficiencies representing voltaic and ohmic losses must now be specified, rather than the single energy efficiency term in \cref{eqn: Lit_Review_LP_SOC_Constraint_Example}.
@@
 \cref{eqn: Linear_Schedule_Objective_Function_Current_as_Variable} was adapted to include these losses, by subtracting/adding the faradaic over-potential from/to the OCV during discharge/charge, and including the ohmic losses in the form of the classical $P_{Loss} = I^2R$ equation.
-This results in a polynomial objective function for revenue in period $Y$ defined by:
+This results in a nonlinear (NLP) objective function for revenue in period $Y$ defined by:
 
@@
 
-The above function is non-convex due to the presence of bivariate quadratic terms  (e.g.
-$-I_{D,t}I_{C,t-1}$).
-Hence the problem matrix is not positive definite.
-However, by minor adjustment of the diagonal of the $Q$ matrix it is possible to obtain a positive semi-definite matrix.
-In the case of the Gurobi solver, a default diagonal adjustment of 1e-6 is allowed.
-If the required adjustment would be greater than this value, an error message is returned.
+This objective function is non-convex, due to the presence of bivariate quadratic terms  (e.g. $-I_{D,t}I_{C,t-1}$). However, it is possible to solve the problem by generating a positive semidefinite
 
@@
 Each optimisation problem was posed using PYOMO within a Python loop for repeat solutions on 24h sets of data.
-The Gurobi solver was called using the SolverFactory function.
+The gurobi solver was called via the SolverFactory module.
 The script was run on a virtual machine (\cite{VirtualBox}) running Linux Mint, which was allocated 1024 MB RAM.
 
-\subsection{NLP Treatment of Voltaic Losses}
+\subsection{NLP treatment of Voltaic Losses}
 \label{Results_NLP_Treatment_Voltaic_losses}
@@
 
-In a TEA of this VRFB, use of the LP formulation would therefore lead to meaningful errors in calculation of revenue (in this case an overestimation).
+In a techno-economic analysis of this VRFB, use of the LP formulation would therefore lead to errors in calculation of revenue (in this case an overestimation).
 Use of the NLP formulation to determine scheduling in a practical application would result in significantly higher revenue than would the LP formulation.
@@
 
-\subsection{NLP with Functional OCV}
+\subsection{NLP with functional OCV}
 \label{Results_NLP_Functional_OCV}
@@
 }
-\caption{(a) Optimal VRFB schedules derived for the 22\textsuperscript{nd} August 2017 using NLP formulations where OCV is fixed ($OCV_{50\%}$) and functionalised.
-(b) Optimal values of OCV where it is functionalised.}
+\caption{(a) Optimal VRFB schedules derived for the 22\textsuperscript{nd} August 2017 using NLP formulations where OCV is fixed $OCV_{50\%}$ and functionalised.
+(b)Optimal values of OCV where it is functionalised.}
 \end{figure}
@@
 Across 2017 the formulation with functional OCV returns \pounds 26.40 compared to \pounds 27.81 for the fixed OCV case.
-The assumption of a fixed OCV (parametrised at 50\% SOC) therefore results in a small but meaningful overestimation of revenue in this application.
+The assumption of a fixed OCV (parametrised at 50\% SOC) therefore results in a small but not insignificant overestimation of revenue in this application.
 Including functionalised OCV in the objective function increases the solve time - the 2017 data were processed in \SI{217}{\second} compared to \SI{129}{\second} for the convex objective function in \cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable}.
 
-The non-convex objective function in \cref{eqn: NLP_true_OCV_long_form} makes the problem less robust: in the 2017 data there were two days in which the price parameter set yielded problem matrices that required greater diagonal adjustment than the default maximum permitted value of 1e-6.
+The non-convex objective function in \cref{eqn: NLP_true_OCV_long_form} makes the formulation less robust: in the 2017 data there were two days in which the price parameter set yielded problem matrices that required greater diagonal adjustment than the default maximum permitted value of 1e-6.
+In these instances Gurobi returns an error message describing the diagonal adjustment that would be required.
 Solutions for the two days were obtained by increasing the Gurobi Parameter "PSDTol" to 1.7e-6.
-For certain applications this approach may become impractical.
 
-\subsection{Cell-Voltage Constraint}
+\subsection{Cell-Voltage constraint}
 \label{Results_Cell_Voltage_Constraint}
@@
 up to the AC-DC boundary), for example in the optimal specification of pump rating, or cooling system for a particular application.
-RFB systems have previously been the subject of technical optimisation, for example the study of the trade-offs between electrolyte concentration and membrane ASR (\cite{Weber2013} \emph{this should be weber et al.}), or shunt currents and pumping losses (\cite{Viswanathan2014}.
-In these works, a simple metric such as levelised cost of electrical energy under a simple daily cycle is used as the objective to minimise.
+RFB systems have previously been the subject of technical optimisation, for example the study of the trade-offs between electrolyte concentration and ASR (\cite{Weber2013} \emph{this should be weber et al.}), or shunt currents and pumping losses (\cite{Viswanathan2014}.
+In these works, a simple metric such as levelised cost of electrical energy under a simple duty cycle is used as the objective to minimise.
 The framework introduced here allows optimisation in terms of revenue under optimal scheduling for any application where a price signal exists.

### 2018-10-29 Paper/main.tex

@@
 
-where $P_{Dis,t}$ and $P_{Chg,t}$ represent the bounded variables discharge and charge power between the RFB and the rest of the system in period $t$, $\tau$ the time-step parameter (length of period $t$) and $p_t$ the parameter price of electrical energy in period $t$.
+where $P_{Dis,t}$ and $P_{Chg,t}$ represent the bounded variables discharge and charge power between the RFB and the rest of the system in period $t$, $\tau$ the time-step parameter (length of period $t$) and $p_t$ the parameter price of electrical energy in period $t$ \sol{it would be worth defining you $Y$ here and summing over $t \in Y$, this would clarify some of your later text}.
 
@@
 It is however necessary to incorporate voltaic losses in the objective function.
-The objective function for revenue in period $Y$ (to be maximised) is defined by:
+The objective function for revenue in period $Y$ \sol{we need to define $Y$ as the set of times points, t}(to be maximised) is defined by:
 
@@
 So that the optimisation may be legitimately performed on consecutive periods of historical data (i.e.
-maintaining an energy balance) the SOC is required to return to 50\% at the end of each period.
+maintaining an energy balance) the SOC is required to return to 50\% at the end of each period\sol{in the introduction you noted that other authors forced the SOC to zero, why is the difference?}.
 This constraint is formalised by:
@@
 A major simplification in the LP problem in terms of power is the assumption of a constant energy efficiency parameter.
-This assumption also applies to the LP problem in current-density terms introduced in \cref{Model_Formulation_LP_Reference_Current_Density_Terms}, although separate voltaic and coulombic efficiencies parameters are called for.
-In reality, ohmic (and pseudo-ohmic) over-potential is a linear function of current-density .
+This assumption also applies to the LP problem in current-density terms introduced in \cref{Model_Formulation_LP_Reference_Current_Density_Terms}, although separate voltaic and coulombic efficiencies parameters are accounted for for.
+In reality, ohmic (and pseudo-ohmic) over-potential is a linear function of current-density.
 The faradaic over-potential, which occurs when the circuit is closed, may be considered constant with respect to current-density (CITE article I used in CR).
 \cref{eqn: Linear_Schedule_Objective_Function_Current_as_Variable} was adapted to include these losses, by subtracting/adding the faradaic over-potential from/to the OCV during discharge/charge, and including the ohmic losses in the form of the classical $P_{Loss} = I^2R$ equation.
-This results in a nonlinear (NLP) objective function for revenue in period $Y$ defined by:
+This results in a nonlinear (NLP) objective function for revenue in period $Y$ given by:
 
@@
 The overall formulation hence comprises a nonlinear objective function subject to linear constraints.
-This is an improvement over the formulations of losses expressed in power efficiency terms by \cite{Nguyen2015,Sarker2017}, both of which use non-linear SOC constraint expressions.
-According to \cite{Hart2017}
-
-\textit{"In the context of nonlinear models, those with nonlinear objective functions are typically easier to optimize than models with nonlinear constraint expressions"}
+This is an improvement over the formulations of losses expressed in power efficiency terms by \cite{Nguyen2015,Sarker2017}, both of which use non-linear SOC constraint expressions, as it is likely to be easier to resolve \cite{Hart2017}.
 

### 2018-10-29 Paper/main.tex
Add files via upload

@@
 
-In a techno-economic analysis of this VRFB, use of the LP formulation would therefore lead to errors in calculation of revenue (in this case an overestimation).
+In a techno-economic analysis of this VRFB, use of the LP formulation would therefore lead to meaningful errors in calculation of revenue (in this case an overestimation).
 Use of the NLP formulation to determine scheduling in a practical application would result in significantly higher revenue than would the LP formulation.
@@
 The difference is small because high current-density is already discouraged by the functionalised voltaic losses in \cref{eqn: NLP_true_OCV_long_form}.
+\cref{fig:V_max_schedules_no_effect_day} shows the optimal scheduling with and without voltage constraint for the median revenue day 22\textsuperscript{nd} August 2017, and \cref{fig:V_max_voltages_no_effect_day} the corresponding cell voltage profiles, showing a maximum cell-voltage of \SI{1.62}{\volt} and hence no violation in the unconstrained schedule.
+
+\begin{figure}[!ht]
+\centering
+\subfloat[][]{
+\label{fig:V_max_schedules_no_effect_day}
+\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./V_max_schedules_no_effect_day.pdf}
+}
+\subfloat[][]{
+\label{fig:V_max_voltages_no_effect_day}
+\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./V_max_voltages_no_effect_day.pdf}
+}
+\caption{(a) Optimal VRFB schedules derived for the 22\textsuperscript{nd} August 2017 by maximising \cref{eqn: NLP_true_OCV_long_form} with no cell-voltage constraint, and subject to the cell voltage constraint given in \cref{eqn: Cell_Voltage_Constraint}.
+(b) Corresponding cell-voltage profiles.}
+\end{figure}
+
 \cref{fig:V_max_schedules} shows the optimal schedules with and without constraint for 5\textsuperscript{th} February 2017, where the difference in revenue was greatest (\pounds 0.018 vs.
@@
 \caption{(a) Optimal VRFB schedules derived for the 5\textsuperscript{th} February 2017 by maximising \cref{eqn: NLP_true_OCV_long_form} with no cell-voltage constraint, and subject to the cell voltage constraint given in \cref{eqn: Cell_Voltage_Constraint}.
-(b)Corresponding cell-voltage profiles.}
+(b) Corresponding cell-voltage profiles.}
 \end{figure}

### 2018-10-29 Paper/main.tex
updating the tex file

--- 2018-10-29 Paper/main.tex
@@
 A simplification in the NLP problem introduced in \cref{Model_Formulation_NLP_treatment_voltaic_losses} is the use of a fixed OCV, ($OCV_{50\%}$) where OCV is in fact a function of SOC.
-If the OCV is linear with respect to SOC in the permitted SOC range (which, in the case of the system described by \cite{Reed2016} it is), and the RFB transfers as much energy below 50\% SOC as above it, then this assumption may be reasonable.
+If the OCV is linear with respect to SOC in the permitted SOC range, which is the case of the system described by \cite{Reed2016} it is, and the RFB transfers as much energy below 50\% SOC as above it, then this assumption may be reasonable.
 However, using the current-density formulation, it is possible to functionalise the relationship between OCV and SOC.
-The relationship between the average OCV and the average SOC in sub-period t is approximated by:
+The relationship between the average OCV and the average SOC in sub-period $t$ is approximated by:
 
@@
 
-Replacing $OCV_{50\%}$ with $OCV_t$ in \cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable} and expanding $SOC_t$ and $SOC_{t-1}$ by \cref{eqn: Method_SOC_Tracker} results in an objective function (to be maximised) defined by:
+Replacing $OCV_{50\%}$ with $OCV_t$ in \cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable} and expanding $SOC_t$ and $SOC_{t-1}$ by \cref{eqn: Method_SOC_Tracker} results in an objective function  defined by:
 
@@
     \begin{split}
-R_{NLP} = \frac{A.\tau}{1\times 10^{6}}\sum_{t}p_{t}(\\
+\max R_{NLP} = \frac{A.\tau}{1\times 10^{6}}\sum_{t}p_{t}(\\
 (1-l_{BOP})I_{D, t}(a\frac{SOC_{t-1} + SOC_{t-2} + \frac{A.\tau}{1000C}(I_{C,t}\sqrt{\eta_C} - \frac{I_{D,t}}{\sqrt{\eta_C}} +  I_{C,t-1}\sqrt{\eta_C} - \frac{I_{D,t-1}}{\sqrt{\eta_C}})}{2} + b - V_a)\\ - \frac{1}{1-l_{BOP}}(I_{C, t}(a\frac{SOC_{t-1} +  SOC_{t-2} + \frac{A.\tau}{1000C}(I_{C,t}\sqrt{\eta_C} - \frac{I_{D,t}}{\sqrt{\eta_C}} + I_{C,t-1}\sqrt{\eta_C} - \frac{I_{D,t-1}}{\sqrt{\eta_C}})}{2} + b + V_a))\\
@@
 
-This objective function is non-convex, due to the presence of bivariate quadratic terms  (e.g. $-I_{D,t}I_{C,t-1}$). However, it is possible to solve the problem by generating a positive semidefinite
+This objective function is non-convex, due to the presence of bivariate quadratic terms (e.g.
+$-I_{D,t}I_{C,t-1}$).
+However, it is possible to solve the problem by generating a positive semidefinite\sol{I think that we should just leave it at the previous sentence, as the thing turns our to work anyway}
 
@@
 It has so far been assumed that the RFB may accept the maximum charge current-density right up to the maximum permitted SOC.
-In reality this may result in the cell-voltage being too-high (overcharging, discussed further in \cref{Results_Cell_Voltage_Constraint}.
-An alternative charging approach (common for Li-ion batteries where high cell voltages pose safety concerns) is to switch to a constant voltage taper charge at high SOC, where the current declines until a cut-off value is reached.
-In the case of an RFB, the fixed losses described above would likely make taper charging uneconomical (see the low efficiency at low current-density reported by \cite{Nguyen2014}).
-Rather than globally adjust the maximum permitted charge current-density or SOC (which may lead to underestimation of revenue), the objective function in \cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable} is made subject to a new linear constraint in terms of cell-voltage, defined by:
+In reality this may result in the cell-voltage being too-high, discussed further in \cref{Results_Cell_Voltage_Constraint}.
+An alternative charging approach, which is common for Li-ion batteries where high cell voltages pose safety concerns, is to switch to a constant voltage taper charge at high SOC, where the current declines until a cut-off value is reached.
+In the case of an RFB, the fixed losses described above would likely make taper charging uneconomical, as deomstrated by the low efficiency at low current-density reported by \cite{Nguyen2014}.
+Rather than globally adjust the maximum permitted charge current-density or SOC, which may lead to underestimation of revenue, the objective function in \cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable} is made subject to a new linear constraint in terms of cell-voltage, defined by:
 
@@
 \cite{Reed2016} studied various electrode flow architectures on a \SI{1}{\kilo\watt} VRFB system based on a \SI{50}{\micro\meter} Nafion NR212 membrane.
-In this case study, performance data were taken from the test of the highest performing stack architecture (IDD2s) at an electrolyte flow rate of 400 \si{\ml\per\minute} (per cell).
+In this case study, performance data were taken from the test of the highest performing stack architecture (IDD2s) at an electrolyte flow rate of 400 \si{\ml\per\minute} per cell.
 This setup was tested at three current-densities: 1600, 2400 and 3200 \si{\ampere\per\square\meter}.
@@
 In the parametrisation of the VRFB for following case studies $P_{rated}$ was set at \SI{1}{\kilo\watt}, the energy to power capacity ratio $r_{e:p}$ was set at 4 and $\eta_{rated}$ was set at 75\%.
-$L_{BOP}$ was set at 0.02, based on a conservative interpretation of the pumping power data reported for a very similar system by \cite{Kim2013}.
+$L_{BOP}$\sol{should this be little l?} was set at 0.02, based on a conservative interpretation of the pumping power data reported for a very similar system by \cite{Kim2013}.
 A coulombic efficiency ($\eta_C$) of 0.975 was derived as the average of the values reported by \cite{Reed2016} at the three current-densities tested, which only varied between 0.974 and 0.976.
@@
 $OCV_{50\%}$ was taken from \cite{Kim2011}, where the OCV of a chloride electrolyte was reported at 20\%, 50\% and 80\% SOC.
-Although a different electrolyte (based on a mixture of sulphuric acid and hydrochloric acid) was used in \cite{Reed2016}, OCV data have not been published.
+Although a different electrolyte, based on a mixture of sulphuric acid and hydrochloric acid, was used in \cite{Reed2016} the OCV data have not been published.
 
@@
 A value of 0.842 was applied, as this is the voltaic efficiency at the mid-point of the permitted current-density range (\SI{1600}{\ampere\per\square\meter}).
-The NLP formulation in \cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable} requires values for area specific resistance ($ASR$) and Faradaic potential ($V_a$) in order to parametrise voltaic losses in the objective function.
-These were derived by multiplying $OCV_{50\%}$ by $1-\sqrt{\eta_V}$ to obtain an over-potential $\delta_v$ at the three current-density points in the experimental data (assuming losses are symmetrical in charge and discharge).
+Similarly, the NLP formulation in \cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable} requires values for area specific resistance ($ASR$) and Faradaic potential ($V_a$) in order to parametrise voltaic losses in the objective function.
+These were derived by multiplying $OCV_{50\%}$ by $1-\sqrt{\eta_V}$ to obtain an over-potential $\delta_v$ at the three current-density points in the experimental data, assuming losses are symmetrical in charge and discharge.
 $ASR$ was obtained as the gradient of the linear regression on this plot, and $V_a$ as the intercept at zero current-density.
@@
 
-The optimisation of scheduling was performed for each day in 2017 using both the NLP formulation and the LP benchmark.
-\Cref{fig:IDD2s_opt_schedules} shows the optimal hourly schedule determined by each method for the 22\textsuperscript{nd} of August (this date was chosen because the resultant revenue is the median of the range obtained across the year).
+The optimisation of the scheduling was performed for each day in 2017 using both the NLP formulation and the LP benchmark.
+\Cref{fig:IDD2s_opt_schedules} shows the optimal hourly schedule determined by each method for the 22\textsuperscript{nd} of August, this date was chosen because the resultant revenue is the median of the range obtained across the year.
 
@@
 In reality however, the voltaic losses that would occur under the LP schedule would reduce the revenue to \pounds 0.051.
-Across 2017, LP scheduling would have resulted in \pounds 22.61 (against prediction of \pounds 33.16) and NLP scheduling would have yielded \pounds 27.99.
+Across 2017, LP scheduling would have resulted in \pounds 22.61, against a prediction of \pounds 33.16, and NLP scheduling would have yielded \pounds 27.99.
 
-In a techno-economic analysis of this VRFB, use of the LP formulation would therefore lead to errors in calculation of revenue (in this case an overestimation).
-Use of the NLP formulation to determine scheduling in a practical application would result in significantly higher revenue than would the LP formulation.
+In a techno-economic analysis of this VRFB, use of the LP formulation would therefore lead to errors in calculation of revenue, in this case an overestimation.
+Use of the NLP formulation to determine scheduling in a practical application would result in the prediction of significantly higher revenue than would the LP formulation.
 It would be possible to set a representative voltaic efficiency ($\bar\eta_V$) in the LP formulation such that the estimate of revenue would be closer to that of the NLP formulation, but this value cannot be known \textit{a priori}.
 
-Solving \cref{eqn: Linear_Schedule_Objective_Function_Current_as_Variable} for each day in 2017 takes \SI{129}{\second} which is a minimal increase over the \SI{127}{\second} required to solve the LP objective function in \cref{eqn: Linear_Schedule_Objective_Function_Current_as_Variable}.
+Importantly, the increased fidelity in the formulation results in a minimal increase in solution time; in this case, solving  \cref{eqn: Linear_Schedule_Objective_Function_Current_as_Variable} for each day in 2017 takes \SI{129}{\second} over the \SI{127}{\second} required to solve the LP objective function in \cref{eqn: Linear_Schedule_Objective_Function_Current_as_Variable}.
 
@@
 \label{Results_NLP_Functional_OCV}
-In the second case study, the impact of expressing OCV as a function of SOC (and hence of the current-density variables in preceding sub-periods) was investigated.
+In the second case study, the impact of expressing OCV as a function of SOC, and hence of the current-density variables in preceding sub-periods, was investigated.
 Parameters $a$ and $b$ representing the linear relationship between OCV and SOC in \cref{eqn: NLP_true_OCV_long_form} were set at 0.2667 and 1.333, after linear regression on data published by \cite{Kim2011}.
@@
 
-The performance of arbitrage between 16:00 and 20:00 is notably reduced when OCV is functionalised.
+The performance of arbitrage between 16:00 and 20:00 is notably reduced in \cref{fig:OCV}.
 In this period there are small fluctuations in price from which the RFB derived revenue when a fixed OCV was assumed.
@@
 
-The non-convex objective function in \cref{eqn: NLP_true_OCV_long_form} makes the formulation less robust: in the 2017 data there were two days in which the price parameter set yielded problem matrices that required greater diagonal adjustment than the default maximum permitted value of 1e-6.
+%The non-convex objective function in \cref{eqn: NLP_true_OCV_long_form} makes the formulation less robust: in the 2017 data there were two days in which the price parameter set yielded problem matrices that required greater diagonal adjustment than the default maximum permitted value of 1e-6.
 In these instances Gurobi returns an error message describing the diagonal adjustment that would be required.
@@
 Overcharging may lead to oxygen evolution at the cathode and hydrogen evolution at the anode.
-This results in reduced coulombic efficiency and  (more importantly) the evolved gases can cause a range of problems as described by \cite{Kear2012}.
-The particular cell-voltage at which gas evolution becomes problematic will vary from system to system (as it depends on the particular electrolyte chemistry).
+This results in reduced coulombic efficiency and, more importantly, the evolved gases can cause a range of problems as described by \cite{Kear2012}.
+The particular cell-voltage at which gas evolution becomes problematic will vary from system to system as it depends on the particular electrolyte chemistry.
 In this case study $V_{max}$ in \cref{eqn: Cell_Voltage_Constraint} was set at 1.65, just below the voltage at which \cite{Wei2017} reported evolution of macroscopic bubbles of hydrogen and oxygen (\SI{1.70}{\volt}).
@@
 Maximising \cref{eqn: NLP_true_OCV_long_form} subject to \cref{eqn: Cell_Voltage_Constraint} for the 2017 data resulted in \pounds 26.07 compared to \pounds 26.40 when the cell voltage was not constrained.
-The difference is small because high current-density is already discouraged by the functionalised voltaic losses in \cref{eqn: NLP_true_OCV_long_form}.
-\cref{fig:V_max_schedules} shows the optimal schedules with and without constraint for 5\textsuperscript{th} February 2017, where the difference in revenue was greatest (\pounds 0.018 vs.
-\pounds 0.020).
+The difference is small because high current-density is already discouraged by the voltaic losses in \cref{eqn: NLP_true_OCV_long_form}.
+\cref{fig:V_max_schedules} shows the optimal schedules with and without constraint for 5\textsuperscript{th} February 2017, where the difference in revenue was greatest (\pounds 0.018 compared to \pounds 0.020).
 \cref{fig:V_max_voltages} shows the corresponding cell-voltage profiles.
@@
 
-    \item It is assumed that both coulombic efficiency and the voltaic loss parameters are invariant with SOC and symmetrical (i.e.
-the same for charge and discharge).
-Regarding the first assumption, experimental work by \cite{Nguyen2014} showed that the voltaic discharge efficiency is lower at 70\% SOC than at 35\% SOC, but that the difference decreases as power increase (data are only reported up to \SI{1.8}{\kilo\watt} for a \SI{5}{\kilo\watt} system).
+    \item It is assumed that both coulombic efficiency and the voltaic loss parameters are invariant with SOC and symmetrical, i.e.
+the same for charge and discharge.
+Regarding the first assumption, experimental work by \cite{Nguyen2014} showed that the voltaic discharge efficiency is lower at 70\% SOC than at 35\% SOC, but that the difference decreases as power increases, although data are only reported up to \SI{1.8}{\kilo\watt} for a \SI{5}{\kilo\watt} system.
 Regarding the latter, the same authors provide experimental evidence that the voltaic efficiency of the particular VRFB is higher for charge than discharge.
-The latter issue could be dealt with in the present formulation (if experimental data were made available) by including applying separate $ASR$, $V_a$ and $\eta_{C}$ terms for charge and discharge.
+The latter issue could be dealt with in the present formulation, if experimental data were made available, by including applying separate $ASR$, $V_a$ and $\eta_{C}$ terms for charge and discharge.
 The former issue would be more difficult to deal with, and a first step would be to perform a sensitivity study based on relevant system data.
@@
 the coulombic losses are proportional to the useful current-density.
-This assumption fits the data reported by \cite{Reed2016}, but the behaviour below \SI{1600}{\ampere\per\square\meter} was not reported.
-In principle shunt current losses are a proportional to cell-voltage (as described by \cite{Xing2011}, and hence a function of SOC and over-voltage (via \cref{eqn: Cell_Voltage_Constraint}).
+This assumption is supported by the data reported by \cite{Reed2016}, but the behaviour below \SI{1600}{\ampere\per\square\meter} was not reported.
+In principle shunt current losses are proportional to cell-voltage, as described by \cite{Xing2011}, and hence a function of SOC and over-voltage (via \cref{eqn: Cell_Voltage_Constraint}.
 The dependence of coulombic losses on current-density is therefore weaker than implied by the present assumption.
@@
     \item The RFB is not thermally constrained.
-The VRFB based on a sulphate electroylte is liable to precipitation of  V\textsubscript{2}O\textsubscript{5} from the catholyte as temperature increases (\cite{Rahman2009,Vijayakumar2011}).
-Although the chloride and mixed acid electrolytes developed at PNNL are more stable (\cite{Li2011,Kim2011}), temperature remains an important consideration.
+The VRFB based on a sulphate electroylte is liable to precipitation of  V\textsubscript{2}O\textsubscript{5} from the catholyte as temperature increases \citep{Rahman2009,Vijayakumar2011}.
+Although the chloride and mixed acid electrolytes developed at PNNL are more stable \citep{Li2011,Kim2011}, temperature remains an important consideration.
 The framework introduced here could easily be expanded to capture the thermal performance of the system.
@@
 
-As discussed above, both the coulombic losses and the pumping power may be better represented as fixed terms.
+%As discussed above, both the coulombic losses and the pumping power may be better represented as fixed terms.
 In this case, the efficiency of the system at low power input/output will be unfavourable (as demonstrated by \cite{Nguyen2014}).
 An issue with representing these losses as fixed terms, is that they would be present at all times in the schedule.
-In order to minimise parasitic losses, a binary "on/off" variable could be added to the formulation in order to allow idling.
+In order to minimise parasitic losses, a binary variable could be added to the formulation in order to allow idling.
 
@@
 up to the AC-DC boundary), for example in the optimal specification of pump rating, or cooling system for a particular application.
-RFB systems have previously been the subject of technical optimisation, for example the study of the trade-offs between electrolyte concentration and ASR (\cite{Weber2013} \emph{this should be weber et al.}), or shunt currents and pumping losses (\cite{Viswanathan2014}.
+RFB systems have previously been the subject of technical optimisation, for example the study of the trade-offs between electrolyte concentration and ASR (\cite{Weber2013} \emph{this should be weber et al.}), or shunt currents and pumping losses \citep{Viswanathan2014}.
 In these works, a simple metric such as levelised cost of electrical energy under a simple duty cycle is used as the objective to minimise.
@@
 
-This work was supported by the UK Engineering and Physical Science Research Council via grant EP/L016818/1 which funds the Centre for Doctoral Training in Energy Storage and its Applications. Drax?
+This work was supported by the UK Engineering and Physical Science Research Council via grant EP/L016818/1 which funds the Centre for Doctoral Training in Energy Storage and its Applications.
+The authors are also grateful to Drax Power Ltd.
+for their financial support.
 

### 2018-10-30 Paper/main.tex
Update main.tex

@@
 
-In this work, the schedule problem is formulated in terms of current-density rather than power, which allows the separation of coulombic and voltaic efficiency components.
-It will be shown that important constraints on the problem may be posed more readily in this form.\sol{you do more than this, you formulate this problem 3 times with increasing levels of complexity then apply each of these to a case study to compare their performance.
-You need to say this and be sure to say that this is all novel, otherwise it's not clear why the reviewers should be bothered}
+The key innovation of the present work is the introduction of a RFB schedule optimisation framework posed in terms of current-density rather than power.
+This allows the separation of the voltaic and coulombic components of the power input/output.
+Using the introduced framework, three novel augmentations (add-ons?) to the typical scheduling problem are posed while maintaining an easily computable problem:
 
-This paper is organised as follows: Section \ref{sec:Model_Formulation} presents...\sol{complete this paragraph}
+\begin{itemize}
+    \item Quadratic expression of $I^2R$ voltaic losses in the objective function under linear constraints.
+    \item Open cell voltage (OCV) as a function of SOC
+    \item Maximum cell voltage constraint in terms of OCV and overpotential
+\end{itemize}
+
+The importance of these improvements to the fidelity of revenue estimation is demonstrated in a day-ahead price arbitrage case study.
 
@@
 \label{sec:Model_Formulation}
-\sol{describe what you're going to do here...}
-\subsection{LP Problem in Current-Density Terms}
+In the following section, a typical LP formulation (REF here?) for schedule optimisation for arbitrage revenue maximisation is first posed in current-density terms to serve as a reference.
+Then the three augmentations to the model described in \cref{sec:Intro} are introduced.
+Finally, the RFB specification method is outlined, including the assumption regarding the rated efficiency.
+\section{LP schedule Optimisation in Current-Density Terms}
 \label{Model_Formulation_LP_Reference_Current_Density_Terms}
-The typical LP formulation for the scheduling problem uses an objective function which is formulated in terms of the power as \sol{give references}:
+In the period $Y$, made up of sub-periods $t$, a typical LP schedule optimisation for arbitrage revenue requires maximisation of the objective function defined by  \sol{give references}:
 
@@
 \label{eqn:Lit_Review_LP_Obj_Func}
-R = \tau \sum_{t} (P_{Dis,t} - P_{Chg,t}) p_t
+R = \tau \sum_{t \in Y} (P_{Dis,t} - P_{Chg,t}) p_t
 \end{equation}
 
-where $P_{Dis,t}$ and $P_{Chg,t}$ represent the bounded variables discharge and charge power between the RFB and the rest of the system in period $t$, $\tau$ the time-step parameter (length of period $t$) and $p_t$ the parameter price of electrical energy in period $t$ \sol{it would be worth defining you $Y$ here and summing over $t \in Y$, this would clarify some of your later text}.
+where $P_{Dis,t}$ and $P_{Chg,t}$ represent the bounded variables discharge and charge power between the RFB and the rest of the system in period $t$, $\tau$ the time-step parameter (length of sub-period $t$) and $p_t$ the parameter price of electrical energy in period $t$.
 
@@
 
-To re-pose this in terms of current-density, the power variable(s) with charge and discharge current-density variables.\sol{can you describe the relations that you've used here}
-
+The above problem was reposed in terms of current-density by replacing the charge/discharge power variables with charge/discharge current-density variables, and multiplying the overall expression by a representative voltage in order to obtain power input/output by the classical $P = IV$ expression.
 Current-density was chosen over current as the former is a scale-independent metric commonly used in the RFB literature.
-The use of current-density makes multiplication by stack area necessary to obtain absolute current, and multiplication by open-cell-volage (OCV) necessary to obtain power.
+The use of current-density makes multiplication by stack area necessary to obtain absolute current.
 Separate efficiencies representing voltaic and ohmic losses must now be specified, rather than the single energy efficiency term in \cref{eqn: Lit_Review_LP_SOC_Constraint_Example}.
 As current-density is defined based on the current at the RFB terminals, coulombic losses are dealt with upstream in \cref{eqn: Method_SOC_Tracker}.
-It is however necessary to incorporate voltaic losses in the objective function.
-The objective function for revenue in period $Y$ \sol{we need to define $Y$ as the set of times points, t}(to be maximised) is defined by:
+It is however necessary to incorporate voltaic losses in the objective function, as these affect the power input/output.
+In current-density terms, the LP objective function for arbitrage revenue (to be maximised) is defined by:
 
@@
 \label{eqn: Linear_Schedule_Objective_Function_Current_as_Variable}
-R_{LP} = \frac{A.\tau OCV_{50\%}}{1\times 10^{6}}\sum_{t}p_{t}(I_{D, t}\sqrt{\bar\eta_{V}}(1 - l_{BOP}) - \frac{I_{C,t}}{\sqrt{\bar\eta_{V}}(1 - l_{BOP})}) \; \forall t \in Y
+R_{LP} = \frac{A.\tau OCV_{50\%}}{1\times 10^{6}}\sum_{t \in Y}p_{t}(I_{D, t}\sqrt{\bar\eta_{V}}(1 - l_{BOP}) - \frac{I_{C,t}}{\sqrt{\bar\eta_{V}}(1 - l_{BOP})})
 \end{equation}
@@
 So that the optimisation may be legitimately performed on consecutive periods of historical data (i.e.
-maintaining an energy balance) the SOC is required to return to 50\% at the end of each period\sol{in the introduction you noted that other authors forced the SOC to zero, why is the difference?}.
+maintaining an energy balance) the SOC is required to return to 50\% at the end of each period.
+This value is chosen so as not to force the RFB to operate at low or high SOC, as would be the case if the minimum permitted SOC was used, as in \cite{Hu2010}.
 This constraint is formalised by:
@@
 
-This objective function is non-convex, due to the presence of bivariate quadratic terms (e.g.
-$-I_{D,t}I_{C,t-1}$).
-However, it is possible to solve the problem by generating a positive semidefinite\sol{I think that we should just leave it at the previous sentence, as the thing turns our to work anyway}
+This objective function is non-convex, due to the presence of bivariate quadratic terms such as $-I_{D,t}I_{C,t-1}$.
 
@@
 \label{Results_The_Application}
-Pure price arbitrage on the day ahead N2EX electrical market was chosen as a case study, as the availability of future price information makes deterministic schedule optimisation relevant.
+Pure price arbitrage on the day ahead N2EX electrical market was chosen as a case study, as the availability of future price information makes deterministic schedule optimisation relevant as described by \cite{Hu2010}.
 Schedule optimisation was performed on 24h periods, with $\tau$ of 1h.
@@
 In the parametrisation of the VRFB for following case studies $P_{rated}$ was set at \SI{1}{\kilo\watt}, the energy to power capacity ratio $r_{e:p}$ was set at 4 and $\eta_{rated}$ was set at 75\%.
-$L_{BOP}$\sol{should this be little l?} was set at 0.02, based on a conservative interpretation of the pumping power data reported for a very similar system by \cite{Kim2013}.
+$l_{BOP}$ was set at 0.02, based on a conservative interpretation of the pumping power data reported for a very similar system by \cite{Kim2013}.
 A coulombic efficiency ($\eta_C$) of 0.975 was derived as the average of the values reported by \cite{Reed2016} at the three current-densities tested, which only varied between 0.974 and 0.976.
@@
 The NLP optimal schedule also targets the most extreme periods, but spreads the charging/discharging out.
-This is the consequence of factoring variable ohmic losses in the objective function -   high current-density is penalised by higher losses.
+This is the consequence of factoring variable ohmic losses in the objective function, where high current-density is penalised by higher losses.
 Under the LP formulation, the VRFB is predicted to generate \pounds 0.068, whereas the NLP formulation predicts a revenue of \pounds 0.063.
@@
 
-Importantly, the increased fidelity in the formulation results in a minimal increase in solution time; in this case, solving  \cref{eqn: Linear_Schedule_Objective_Function_Current_as_Variable} for each day in 2017 takes \SI{129}{\second} over the \SI{127}{\second} required to solve the LP objective function in \cref{eqn: Linear_Schedule_Objective_Function_Current_as_Variable}.
+Importantly, the increased fidelity in the formulation requires a minimal increase to solution time; in this case, solving  \cref{eqn: Linear_Schedule_Objective_Function_Current_as_Variable} for each day in 2017 takes \SI{129}{\second} versus the \SI{127}{\second} required to solve the LP objective function in \cref{eqn: Linear_Schedule_Objective_Function_Current_as_Variable}.
 
@@
 The particular cell-voltage at which gas evolution becomes problematic will vary from system to system as it depends on the particular electrolyte chemistry.
+
+With the introduction of the functional OCV in \cref{eqn: NLP_true_OCV_long_form}, the application of a maximum cell-voltage constraint is now rational.
 In this case study $V_{max}$ in \cref{eqn: Cell_Voltage_Constraint} was set at 1.65, just below the voltage at which \cite{Wei2017} reported evolution of macroscopic bubbles of hydrogen and oxygen (\SI{1.70}{\volt}).
@@
 The difference is small because high current-density is already discouraged by the voltaic losses in \cref{eqn: NLP_true_OCV_long_form}.
-\cref{fig:V_max_schedules} shows the optimal schedules with and without constraint for 5\textsuperscript{th} February 2017, where the difference in revenue was greatest (\pounds 0.018 compared to \pounds 0.020).
+
+\cref{fig:V_max_schedules_no_effect_day} shows the optimal scheduling with and without voltage constraint for the median revenue day 22\textsuperscript{nd} August 2017, and \cref{fig:V_max_voltages_no_effect_day} the corresponding cell voltage profiles, showing a maximum cell-voltage of \SI{1.62}{\volt} and hence no violation in the unconstrained schedule.
+
+\begin{figure}[!ht]
+\centering
+\subfloat[][]{
+\label{fig:V_max_schedules_no_effect_day}
+\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./V_max_schedules_no_effect_day.pdf}
+}
+\subfloat[][]{
+\label{fig:V_max_voltages_no_effect_day}
+\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./V_max_voltages_no_effect_day.pdf}
+}
+\caption{(a) Optimal VRFB schedules derived for the 22\textsuperscript{nd} August 2017 by maximising \cref{eqn: NLP_true_OCV_long_form} with no cell-voltage constraint, and subject to the cell voltage constraint given in \cref{eqn: Cell_Voltage_Constraint}.
+(b) Corresponding cell-voltage profiles.}
+\end{figure}
+
+\cref{fig:V_max_schedules} shows the optimal schedules with and without constraint for 5\textsuperscript{th} February 2017, where the difference in revenue was greatest (\pounds 0.018 vs.
+\pounds 0.020).
 \cref{fig:V_max_voltages} shows the corresponding cell-voltage profiles.

### 2018-10-31 Paper/main.tex
Update main.tex

@@
 
-For the optimisation of arbitrage revenue, the linear programming (LP) approach typically involves the maximisation of the objective function for revenue.
-Linear programming is an efficient problem formulation, as the optimum value for each variable will either lie at the maximum or minimum of a permitted value, making a search of the interior unnecessary.
-
-Variants of the above formulation have been used to calculate the revenue or avoided costs due to RFB systems in various applications.
-For example, \cite{Hu2010} optimised the schedule of a VRFB and a poly-sulphide bromine RFB in order to maximise arbitrage revenue on the Danish day ahead electrical spot market.
-As the optimisation was to be performed daily, the SOC is constrained to return to 0 at the end of each \SI{24}{\hour} period.
+Schedule optimisation is often performed by posing the objective function and constraints as linear algebraic terms, and using an algorithm such as Simplex to determine the maximum or minimum of the objective function.
+This approach, known as linear programming (LP) is efficient because the optimum value for each variable will either lie at the maximum or minimum of a permitted value, making a search of the interior unnecessary.
 
 \cite{Johnston2015} used a LP formulation to optimise the scheduling of a VRFB installed at a wind farm.
-The objective was to maximise revenue on the wholsesale electrical market while satisfying the frequency response obligations (constraints) of the wind installation\sol{what did they find?} .
+The objective was to maximise revenue on the wholsesale electrical market while satisfying the frequency response obligations (constraints) of the wind installation\sol{what did they find?}.
 \cite{Vaca2017} looked at a similar application but compared the payback period for VRFB and supercapacitor installations with a hybrid system.
@@
 When the micro-grid was isolated, the objective was to minimise the start-up and running costs of micro-generators used to back up the renewable power output.
-The authors used a mixed-integer LP (MILP) approach, where the innteger variables were used required to factor start-up costs, additional constraints were required to guarantee system reliability \sol{are these reliability constraints interesting?}.
+The authors used a mixed-integer LP (MILP) approach, where the integer variables were required to factor start-up costs.
+Additional global constraints were required to guarantee system reliability \sol{are these reliability constraints interesting?
+Yes, if one is interested in a whole system!}.
 
-A significant simplification in all of the above studies is the assumption of a constant efficiency.
-In reality, the efficiency of any BESS is a function of the charge/discharge power (among other variables, such as SOC).
-In the above formulation, if efficiency were defined as a  function of power, it would be necessary to place this function on the denominator of the SOC increment term in the discharge expression in \Cref{eqn: Lit_Review_LP_SOC_Constraint_Example}.\sol{merge the previous and coming sentences to remove the reference to the equations} This complicates the problem as the constraint is now non-linear and non-quadratic.
-\cite{Nguyen2014} report a highly detailed nonlinear\sol{be consistent in you use of 'non-linear' or 'nonlinear'} model for the charge and discharge efficiency of a vanadium RFB.
+\cite{Gomes2017}  studied the application of a VRFB in order to cope with stochastic deviations from the forecast exports of wind and solar power into a day ahead market.
+A two stage optimisation was applied, and LP was used to optimise the VRFB schedule in the first stage.
+
+\cite{Hu2010} optimised the schedule of a vanadium RFB (VRFB) and a poly-sulphide bromine RFB in order to maximise arbitrage revenue on the Danish day ahead electrical spot market.
+This approach was not strictly LP, as discussed in \cref{Model_Formulation_LP_Reference_Current_Density_Terms}, but the problem may be reposed in LP form.
+
+A significant simplification in the representation of the RFB in all of the above studies is the assumption of a constant efficiency.
+In reality, the efficiency of any BESS is a function of the charge/discharge power, among other variables, such as state of charge (SOC).
+In the LP formulation, if efficiency were defined as a  function of power, it would be necessary to place this function on the denominator of the  discharge SOC increment expression.
+This would complicate the problem as the constraint is now non-linear and non-quadratic.
+\cite{Nguyen2014} report a highly detailed non-linear model for the charge and discharge efficiency of a vanadium RFB.
 In a subsequent article (\cite{Nguyen2015}) this function was incorporated into the SOC expression of a schedule optimisation problem in the aforementioned fashion.
-Due to the complication of the SOC constraint, the authors employ a search approach with dynamic programming to break the problem down.
+Due to the complication of the SOC constraint, the authors employ a search approach with dynamic programming (DP) to break the problem down.
 This is more computationally intensive than the LP method, but the authors state that an optimum schedule may be found in an acceptable time-scale as the small number of variables and the presence of constraints limit the search space.
+A DP search approach is also used by \cite{Oudalov2007}, where the internal resistance of the BESS is expressed as a function of SOC in a TEA of a VRFB and a lead acid battery for industrial peak-shaving.
 
@@
 The charge efficiency is left in the SOC expression (redefining the charge power variable in the same way would require dividing it by the quadratic charge efficiency term in the objective function).
-The problem therefore becomes convex (quadratic), but nonlinear in both objective function and constraint.
+The problem therefore becomes convex (quadratic), but non-linear in both objective function and constraint.
 The authors then employ a linear approximation technique\sol{does this need a reference}, and combine the efficiency treatment with a piece-wise linear treatment of degradation as a function of power, giving an overall mixed integer-linear programming problem (MILP).
@@
 \label{sec:Model_Formulation}
-In the following section, a typical LP formulation (REF here?) for schedule optimisation for arbitrage revenue maximisation is first posed in current-density terms to serve as a reference.
+Pure price arbitrage is chosen as the application for schedule optimisation due to its simplicity, allowing a focus on the treatment of the RFB within the problem.
+In the following section, an LP formulation for schedule optimisation for maximum arbitrage revenue is first posed in current-density terms to serve as a reference.
 Then the three augmentations to the model described in \cref{sec:Intro} are introduced.
@@
 \label{Model_Formulation_LP_Reference_Current_Density_Terms}
-In the period $Y$, made up of sub-periods $t$, a typical LP schedule optimisation for arbitrage revenue requires maximisation of the objective function defined by  \sol{give references}:
+In the literature on RFB schedule optimisation, scheduling problems are posed using either power (\cite{Chen2012,Hu2010,Nguyen2015}) or energy (\cite{Johnston2015,Vaca2017}) as the time-indexed variables.
+This is not an important difference, as the former only requires that the timestep $\tau$ be included as a coefficient.
+Here, the power convention is adopted, as it allows the timestep to be varied independently.
+The objective function for pure arbitrage revenue described by \cite{Hu2010} is adopted here, but adapted so that separate charge and discharge power variables are explicitly defined.
+Specifying a single power variable with negative values representing charge and positive discharge is not practical for LP optimisation, as the solver will drive the variable to extremes which is likely to be sub-optimal.
+
+In the period $Y$, made up of sub-periods $t$, the schedule is optimised by maximising the revenue defined by:
 
@@
 
-where $P_{Dis,t}$ and $P_{Chg,t}$ represent the bounded variables discharge and charge power between the RFB and the rest of the system in period $t$, $\tau$ the time-step parameter (length of sub-period $t$) and $p_t$ the parameter price of electrical energy in period $t$.
+where $P_{Dis,t}$ and $P_{Chg,t}$ represent the bounded variables discharge and charge power (\si{\watt}) between the RFB and the rest of the system in sub-period $t$.
+$p_t$ is the price of electrical energy (\SI[sticky-per, bracket-unit-denominator = false]{}[\pounds]{\per\mega\watt\hour}) in sub-period $t$, and $\tau$ the time-step parameter  (\si{\hour}).
 
-The state of charge (SOC) of the BESS is then defined by:
+The SOC of the RFB is defined after \cite{Gomes2017} by:
 
@@
 where $E_{BESS}$ is the parameter energy capacity and $\eta_{Chg}$ and $\eta_{Dis}$ are the charge and discharge energy efficiency parameters respectively.
-The SOC is at all times normally constrained to between a maximum of 1 and a minimum of 0, though a narrower window has also used \sol{reference}.
+The SOC must at all times be constrained between a maximum of 1 and a minimum of 0, though a narrower window may be used \cite{Vaca2017}.
 
@@
 where $I_{C,t}$ and $I_{D,t}$ are independent variables representing charge and discharge current density (\si{\ampere\per\square\meter}) respectively.
-$A$ is the stack area parameter (\si{\square\meter}), $\tau$ the timestep (\si{\hour}), $OCV_{50\%}$ the open cell voltage at 50\% state of charge (SOC) and $p_t$ the indexed electrical price parameter (\SI[sticky-per, bracket-unit-denominator = false]{}[\pounds]{\per\mega\watt\hour}).
+$A$ is the stack area parameter (\si{\square\meter}) and $OCV_{50\%}$ the open cell voltage at 50\% SOC.
 $\bar\eta_{V}$ is the representative round-trip voltaic efficiency.
@@
 
-So that the optimisation may be legitimately performed on consecutive periods of historical data (i.e.
-maintaining an energy balance) the SOC is required to return to 50\% at the end of each period.
-This value is chosen so as not to force the RFB to operate at low or high SOC, as would be the case if the minimum permitted SOC was used, as in \cite{Hu2010}.
+So that the optimisation may be legitimately performed on consecutive periods of historical data, it is necessary that the SOC at the end of the period be returned to the starting value.
+\cite{Hu2010} set this value at 0, but in order to avoid imposing low SOC behaviour, it is here set at 0.5.
 This constraint is formalised by:
@@
 \cref{eqn: Linear_Schedule_Objective_Function_Current_as_Variable} was adapted to include these losses, by subtracting/adding the faradaic over-potential from/to the OCV during discharge/charge, and including the ohmic losses in the form of the classical $P_{Loss} = I^2R$ equation.
-This results in a nonlinear (NLP) objective function for revenue in period $Y$ given by:
+This results in a non-linear (NLP) objective function for revenue in period $Y$ given by:
 
@@
 Under the constant coulombic efficiency assumption \cref{eqn: Method_SOC_Tracker} is linear, and so by extension are the constraints in \cref{eqn: Method_Constraint_SOC_Range,eqn: Method_Constraint_E_Conservation}.
-The overall formulation hence comprises a nonlinear objective function subject to linear constraints.
+The overall formulation hence comprises a non-linear objective function subject to linear constraints.
 This is an improvement over the formulations of losses expressed in power efficiency terms by \cite{Nguyen2015,Sarker2017}, both of which use non-linear SOC constraint expressions, as it is likely to be easier to resolve \cite{Hart2017}.
@@
 An alternative charging approach, which is common for Li-ion batteries where high cell voltages pose safety concerns, is to switch to a constant voltage taper charge at high SOC, where the current declines until a cut-off value is reached.
-In the case of an RFB, the fixed losses described above would likely make taper charging uneconomical, as deomstrated by the low efficiency at low current-density reported by \cite{Nguyen2014}.
+In the case of an RFB, the fixed losses described above would likely make taper charging uneconomical, as demonstrated by the low efficiency at low current-density reported by \cite{Nguyen2014}.
 Rather than globally adjust the maximum permitted charge current-density or SOC, which may lead to underestimation of revenue, the objective function in \cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable} is made subject to a new linear constraint in terms of cell-voltage, defined by:
@@
 The dependence of coulombic losses on current-density is therefore weaker than implied by the present assumption.
-As a first approximation, this function could be simplified to a fixed term.
+As a first approximation, a fixed loss term could be placed in the SOC expression in \cref{eqn: Method_SOC_Tracker} instead of the $\sqrt{\eta_C}$ coefficients.
 
@@
     \item The RFB is not thermally constrained.
-The VRFB based on a sulphate electroylte is liable to precipitation of  V\textsubscript{2}O\textsubscript{5} from the catholyte as temperature increases \citep{Rahman2009,Vijayakumar2011}.
+The VRFB based on a sulphate electrolyte is liable to precipitation of  V\textsubscript{2}O\textsubscript{5} from the catholyte as temperature increases \citep{Rahman2009,Vijayakumar2011}.
 Although the chloride and mixed acid electrolytes developed at PNNL are more stable \citep{Li2011,Kim2011}, temperature remains an important consideration.
@@
 
-%As discussed above, both the coulombic losses and the pumping power may be better represented as fixed terms.
+As discussed above, both the coulombic losses and the pumping power may be better represented as fixed terms.
 In this case, the efficiency of the system at low power input/output will be unfavourable (as demonstrated by \cite{Nguyen2014}).
@@
 
-The framework also allows for an additional linear constraint to be placed on upper cell-voltage as a function of SOC and over-potential.
+The fidelity of the RFB representation is further increased by placing a linear constraint on cell-voltage as a function of SOC and over-potential.
 This constraint restricts the current-density at which the battery may be charged when the SOC is high, which is an important consideration for safety and battery lifetime.

### 2018-11-02 Paper/main.tex
Update main.tex

@@
 \usepackage{siunitx}
+\usepackage{hyperref}
+    \hypersetup{colorlinks=true,linkcolor=green}
 
@@
 
-At present, revenue from arbitrage on the wholesale electrical price is not in itself sufficient to justify BESS investment, and the deployment of grid scale BESS is dominated by projects that have won contracts for short term stability services tendered by National Grid.
+At present, revenue from arbitrage on the wholesale electrical price is not in itself sufficient to justify BESS investment \cite{Johnston2015}.
+and the deployment of grid scale BESS is dominated by projects that have won contracts for short term stability services tendered by National Grid.
 For example, \SI{500}{\mega\watt} of BESS capacity won contracts for the T-4 2020/21 capacity market (CM) auction, and \SI{200}{\mega\watt} of BESS capacity won enhanced frequency response (EFR) contracts.
@@
 \sol{This is a bit parochial, the same is regarding BESS elsewhere (i.e.
-the big Tesla abttery).
+the big Tesla battery).
 We need to widen the story out to make the paper relevant outside the UK}
@@
 
-Schedule optimisation is often performed by posing the objective function and constraints as linear algebraic terms, and using an algorithm such as Simplex to determine the maximum or minimum of the objective function.
-This approach, known as linear programming (LP) is efficient because the optimum value for each variable will either lie at the maximum or minimum of a permitted value, making a search of the interior unnecessary.
+Schedule optimisation is often performed by posing a linear programming (LP) problem \cite{}.
+In this method, the objective function and the constraints on the system are expressed as linear algebraic terms.
+On the resulting geometric space, the optimum is guaranteed to lie on a vertex, hence a costly search of the whole surface is not required.
 
 \cite{Johnston2015} used a LP formulation to optimise the scheduling of a VRFB installed at a wind farm.
-The objective was to maximise revenue on the wholsesale electrical market while satisfying the frequency response obligations (constraints) of the wind installation\sol{what did they find?}.
-\cite{Vaca2017} looked at a similar application but compared the payback period for VRFB and supercapacitor installations with a hybrid system.
-In both studies, a cost of degradation in \pound per unit energy throughput was applied as a coefficient to ESS power variables in the objective function.
-This approach discourages the performance of arbitrage below a certain threshold price differential \sol{why is this note worthy?}.
+The objective was to maximise revenue on the wholesesale electrical market and the frequency response market, while always maintaining an energy reserve in order to meet .
+By repeating this process at different energy and power ratings, the authors found that the economically optimal VRFB rating for the \SI{50}{\mega\watt} wind farm was \SI{5.3}{\mega\watt} and \SI{3}{\mega\watt\hour}.
+This illustrates the point made above regarding the poor economics of wholesale market arbitrage - the optimal RFB rating is dictated by the frequency market revenue and the reserve energy requirement.
+Increasing the capacity of the VRFB in order to perform additional wholesale market arbitrage is not economical.
+The authors also suggested that a hybrid system may be suitable, as the economically optimal VRFB operates in a narrow SOC window when providing frequency regulation, with the remaining SOC held in reserve for large frequency deviations that occur infrequently.
 
-A full system schedule optimisation problem was posed by \cite{Chen2012} to assess the benefit of a VRFB in a microgrid with a load, PV, wind and dispatchable micro-generation.
+This thread was picked up by \cite{Vaca2017} who compared a hybrid system consisting of a VRFB and a supercapacitor with the individual systems.
+They found that the hybrid system was the optimal choice, with the VRFB providing the energy capacity and the supercapacitor the low cost power capacity.
+
+It is worth noting that both \cite{Johnston2015,Vaca2017} assume a higher cost of energy capacity than power capacity for the VRFB.
+This contradicts economic modelling of system costs by \cite{Viswanathan2014,Ha2015Estimatingstorage}, where the energy cost is markedly lower.
+
+\cite{Johnston2015} and \cite{Vaca2017} apply a cost of degradation in \pound per unit energy throughput as a coefficient to ESS power variables in the objective function.
+This approach captures an important consideration in electrical price arbitrage using ESS - that there is a threshhold price differential below which the degradation cost outweighs the revenue.
+In both works the capital cost of the ESS and the degradation cost are included, but the residual value at project end is not counted.
+\cite{Dufo-Lopez2009} present a more detailed economic model for various ESS used to increase the value of wind power sold into a day-ahead market.
+The residual value of the ESS at the end of the project is pro-rated according to the number of cycles performed as a fraction of the expected cycle life, and discounted to a net present value.
+The schedule optimisation in this work is not as thorough as the LP approaches of  \cite{Johnston2015,Vaca2017}; the day is divided into two peak periods where the ESS is discharged and two off-peak where it is charged, and the boundaries of the periods may be varied iteratively to find an apparent optimum.
+
+A full system LP schedule optimisation problem was posed by \cite{Chen2012} to assess the benefit of a VRFB in a microgrid with a load, PV, wind and dispatchable micro-generation.
 When the micro-grid was isolated, the objective was to minimise the start-up and running costs of micro-generators used to back up the renewable power output.
@@
 \cite{Gomes2017}  studied the application of a VRFB in order to cope with stochastic deviations from the forecast exports of wind and solar power into a day ahead market.
-A two stage optimisation was applied, and LP was used to optimise the VRFB schedule in the first stage.
+A two stage optimisation was applied, and LP was used to optimise the VRFB schedule based on the forecast  in the first stage.
 
@@
 This allows the separation of the voltaic and coulombic components of the power input/output.
-Using the introduced framework, three novel augmentations (add-ons?) to the typical scheduling problem are posed while maintaining an easily computable problem:
+Using the introduced framework, three novel augmentations \emph{(add-ons?)} to the typical scheduling problem are posed while maintaining an easily computable problem:
 

### 2018-11-06 Paper/main.tex
Update main.tex

@@
 \usepackage{lineno}
+\usepackage{hyperref}
+    \hypersetup{colorlinks=true,linkcolor=green}
 \usepackage{cleveref}
@@
 \usepackage{siunitx}
-\usepackage{hyperref}
-    \hypersetup{colorlinks=true,linkcolor=green}
 
@@
 \label{sec:Intro}
-The  growing deployment of intermittent renewable power sources \sol{you need numbers here to underline how important this is, i.e.
-what percentage of energy, around the world, is now renewable} such as wind and solar in industrialised nations in order to substitute fossil fuel power is a key driver of research into the development and applications of battery energy storage systems \sol{why?
-what does it do?
-the argument needs to highlight the properties that batteries have and why we think are important} (BESS).
-\sol{you need to pack in the references here, you mention research so we need references to know what they looked at...}
+The need to reduce carbon dioxide emissions is motivating increasing levels of renewable energy deployment across industrialised nations, and costs for both solar photovoltaic technologies and wind turbines have dropped considerably in recent years.
+\sol{More citations here}
 
-At present in the UK, the fluctuating output from these power sources is topped up to match demand primarily by varying the output of gas fired power plant \cite{DraxElectricInsights}.
-In order to meet the strict CO\textsubscript{2} emissions targets for the UK electrical supply by 2050, and bearing in mind that nuclear power capacity is not predicted to grow until after 2030 \cite{FES2018}, it is probable that further multi-hour energy storage will be required to displace gas plant from the role of balancing.
-Within the UK electrical market, the phasing out of gas power plants, likely via increments to the carbon price \sol{reference?}, to be replaced by intermittent generation would lead to high electrical prices where demand exceeds supply, and low (or negative) pricing where supply exceeds demand.
+At a national level, the fluctuating output from these power sources is topped up to match demand primarily by varying output from fossil fuel power thermal plant.
+In the UK, this is specifically achieved by gas power plant \cite{DraxElectricInsights}.
+In order to meet the strict CO$_2$ emissions targets for the UK electrical supply by 2050, and bearing in mind that nuclear power capacity is not predicted to grow until after 2030, either carbon capture or significantly greater renewable power penetration will be required of the supply side\cite{FES2018}.
 
-At present, revenue from arbitrage on the wholesale electrical price is not in itself sufficient to justify BESS investment \cite{Johnston2015}.
-and the deployment of grid scale BESS is dominated by projects that have won contracts for short term stability services tendered by National Grid.
-For example, \SI{500}{\mega\watt} of BESS capacity won contracts for the T-4 2020/21 capacity market (CM) auction, and \SI{200}{\mega\watt} of BESS capacity won enhanced frequency response (EFR) contracts.
-The majority of successful BESS applicants for the EFR auction also have CM contracts.
-\sol{This is a bit parochial, the same is regarding BESS elsewhere (i.e.
-the big Tesla battery).
-We need to widen the story out to make the paper relevant outside the UK}
+In economic terms, the price of electricity at a given moment is set by the operating costs and  of the most expensive generator in the merit order \cite{Staffell}.
+As wind and solar have zero marginal cost, increasing penetration of these power sources will reduce electrical price according to a stochastic distribution.
+At the same time, increasing operating costs for dispatchable fossil generation due to reduced capacity factors and carbon taxation will increase the price in some periods.
 
-However, due to concerns that the average storage duration (i.e.
-energy to power ratio) of the BESS contracted for CM is shorter than the \SI{2}{\hour} that a stress event may last for, National Grid recently applied derating factors according to the energy to power ratio.
-Hence, a system rated at \SI{1}{\mega\watt}/\SI{2}{\mega\watt\hour} would only receive capacity payments for \SI{0.68}{\mega\watt} in the 2018/19 T-1 auction, and \SI{0.65}{\mega\watt} in the 2021/22 T-4 auction.
-The derating factor is relaxed to 0.96 for systems rated for 4+ hours of discharge \cite{NGDeratingFactors}.
+Such price variability presents an increasing opportunity for arbitrage by energy storage systems (ESS).
+At present however, revenue from arbitrage on the electrical price alone is not sufficient to justify BESS investment, even under optimistic economic assumptions \cite{Hu2010}.
+Hence an important aspect of BESS economics is revenue stacking, where multiple revenue streams are captured by the same system.
+For example by providing grid reliability services as well as performing market arbitrage.
+In the UK this is illustrated by the fact that the majority of successful BESS applicants for enhanced frequency response (EFR) contracts also have capacity market contracts \cite{NGDeratingFactors}.
 
-This development disfavours the power-centric Li-ion systems that previously won the majority of EFR contracts (Cite please) and favours redox flow battery (RFB) systems, where the incremental cost of energy capacity is low, and systems are typically specified with a 3-6 \si{\hour} energy to power ratio.
+\cite{Sabihuddin2015} classify the services that BESS may provide by the continuous charge/discharge duration, from power quality and regulation ($<$ \SI{1}{\minute}), through bridging power (\SI{1}{\minute} to \SI{1}{\hour}) to energy management ($>$ \SI{1}{hour}).
+In order to satisfy a given duration a BESS must be specified with the corresponding energy to power ratio.
 
-In order to quantify the economic merits of RFB systems in comparison to other BESS, it is necessary to perform techno-economic analyses (TEA) for realistic applications.
+At the shortest timescale, BESS may receive payments from the grid operator to respond to fluctuations in the AC grid frequency, as in the aforementioned EFR contracts.
+Such frequency regulation services attract Li-ion systems designed for power, such as the Hornsdale Power Reserve project in South Australia, which is specified for \SI{1}{\hour}\SI{17}{\minute} at its \SI{100}{\mega\watt} rated power output \cite{DOEGlobalESSDatabase}.
+This energy to power ratio is representative of global Li-ion installation data as collated by \cite{Newbery2018}.
+
+An example of multi-hour energy management is the use of ESS to discharge into peak power demand in a process known as peak shaving.
+In California, where BESS are being considered as replacement for gas powered peaker plants, a project may receive capacity payments in addition to the sale of energy.
+However, the regulator de-rates the payment according to a 4 hour target duration, hence a \SI{1}{\mega\watt}/\SI{2}{\mega\watt\hour} BESS would only receive credit for  \SI{0.5}{\mega\watt} of power capacity \cite{NRELPeakingESS}.
+
+A similar de-rating has also been applied recently in the UK capacity market (CM).
+This was done due to concerns that the average storage duration of the BESS contracted for CM is shorter than the \SI{2}{\hour} that a stress event may last for.
+\cite{NGDeratingFactors}.
+
+The de-rating developments would disfavour the Li-ion systems that have been deployed so far and favour systems with higher energy to power ratios.
+Redox flow battery (RFB) systems and promising candidates for multi-hour applications and revenue stacking, as the incremental cost of energy capacity is low, and systems deployed so far have an average energy to power ratio of 3.5 \cite{Newbery2018}.
+
+In order to quantify the economic merits of RFB systems in comparison to other BESS, or indeed ESS, it is necessary to perform techno-economic analyses (TEA) for particular applications.
 TEA should involve optimisation of the operation of a BESS to maximise revenue (or minimise costs).
@@
 
-Schedule optimisation is often performed by posing a linear programming (LP) problem \cite{}.
+Where a price signal exists optimisation is commonly performed by posing a linear programming (LP) problem \emph{do I dump here a list of all the LP references i will refer to below?}.
 In this method, the objective function and the constraints on the system are expressed as linear algebraic terms.
-On the resulting geometric space, the optimum is guaranteed to lie on a vertex, hence a costly search of the whole surface is not required.
+On the resulting geometric surface, the optimum is guaranteed to lie on a vertex, hence a time-consuming search of the whole surface is not required.
 
@@
 \label{sec:Model_Formulation}
+
 Pure price arbitrage is chosen as the application for schedule optimisation due to its simplicity, allowing a focus on the treatment of the RFB within the problem.
@@
 Then the three augmentations to the model described in \cref{sec:Intro} are introduced.
-Finally, the RFB specification method is outlined, including the assumption regarding the rated efficiency.
+Finally, the RFB specification process is defined.
+
 \section{LP schedule Optimisation in Current-Density Terms}
 \label{Model_Formulation_LP_Reference_Current_Density_Terms}
+
 In the literature on RFB schedule optimisation, scheduling problems are posed using either power (\cite{Chen2012,Hu2010,Nguyen2015}) or energy (\cite{Johnston2015,Vaca2017}) as the time-indexed variables.
@@
 \label{Model_Formulation_NLP_treatment_voltaic_losses}
+
 A major simplification in the LP problem in terms of power is the assumption of a constant energy efficiency parameter.
@@
 In reality, ohmic (and pseudo-ohmic) over-potential is a linear function of current-density.
-The faradaic over-potential, which occurs when the circuit is closed, may be considered constant with respect to current-density (CITE article I used in CR).
-\cref{eqn: Linear_Schedule_Objective_Function_Current_as_Variable} was adapted to include these losses, by subtracting/adding the faradaic over-potential from/to the OCV during discharge/charge, and including the ohmic losses in the form of the classical $P_{Loss} = I^2R$ equation.
+The faradaic, or activation over-potential, which occurs when the circuit is closed, may be considered constant with respect to current-density \cite{Aaron2011}.
+\Cref{eqn: Linear_Schedule_Objective_Function_Current_as_Variable} was adapted to include these losses, by subtracting/adding the faradaic over-potential from/to the OCV during discharge/charge, and including the ohmic losses in the form of the classical $P_{Loss} = I^2R$ equation.
 This results in a non-linear (NLP) objective function for revenue in period $Y$ given by:
@@
 where $V_a$ and $ASR$ are scalar parameters representing faradaic over-potential (\si{\volt}) and area specific resistance (\si{\ohm\square\meter}).
-The placement of $V_a$ and $ASR$ reflects an assumption of symmetry in voltaic losses between charge and discharge (discussed further in \cref{analysis_of_assumptions}).
+The placement of $V_a$ and $ASR$ reflects an assumption of symmetry in voltaic losses between charge and discharge, which is discussed further in \cref{analysis_of_assumptions}.
 The objective function is subject to the same constraints as described in \cref{Model_Formulation_LP_Reference_Current_Density_Terms}.
 
-Although non-linear, \cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable} the objective function is convex (as it requires maximising a negative quadratic), hence a gradient based algorithmic solver is guaranteed to find the optimum from any starting point.
+Although non-linear, \cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable} the objective function is convex, as it requires maximising a negative quadratic, hence a gradient based algorithmic solver is guaranteed to find the optimum from any starting point.
 
@@
 Both the LP and NLP solutions involve charging the VRFB when the electrical price is low (e.g.
-between 05:00 and 07:00) and discharging it when the price is high (e.g.
+05:00 to 07:00) and discharging it when the price is high (e.g.
 10:00 to 13:00).
@@
 This assumption is supported by the data reported by \cite{Reed2016}, but the behaviour below \SI{1600}{\ampere\per\square\meter} was not reported.
-In principle shunt current losses are proportional to cell-voltage, as described by \cite{Xing2011}, and hence a function of SOC and over-voltage (via \cref{eqn: Cell_Voltage_Constraint}.
-The dependence of coulombic losses on current-density is therefore weaker than implied by the present assumption.
+In principle shunt current losses are proportional to cell-voltage, as described by \cite{Xing2011}, and hence a function of SOC and over-voltage (via \cref{eqn: Cell_Voltage_Constraint}.\emph{wait, is it the terminal voltage or the OCV that drives the shunt current?} The dependence of coulombic losses on current-density is therefore weaker than implied by the present assumption.
 As a first approximation, a fixed loss term could be placed in the SOC expression in \cref{eqn: Method_SOC_Tracker} instead of the $\sqrt{\eta_C}$ coefficients.

### 2018-11-07 Paper/main.tex
Update main.tex

@@
 \label{sec:Intro}
-The need to reduce carbon dioxide emissions is motivating increasing levels of renewable energy deployment across industrialised nations, and costs for both solar photovoltaic technologies and wind turbines have dropped considerably in recent years.
-\sol{More citations here}
+Worldwide, the need to reduce CO$_2$ emissions has stimulated the development of a wide range of renewable energy sources.
+Presently, renewable energy is estimated to account for 10\% of global energy consumption, and this is forecast to grow to 12\% by 2023.
+The growth of renewable energy is forecast to be highest in the electrical power sector, from 24\% in 2017 to 30\% in 2023.
+Although hydro power presently accounts for 50\% of renewable power generation, the  intermittent power sources wind and solar are forecast to contribute most to this growth\cite{IEA2018}.
+A number of large European economies, such as Germany, UK, Italy and Spain obtained 23\% to 30\% of there electrical energy from non-hydro renewable power in 2017 \cite{BP2018}.
 
-At a national level, the fluctuating output from these power sources is topped up to match demand primarily by varying output from fossil fuel power thermal plant.
-In the UK, this is specifically achieved by gas power plant \cite{DraxElectricInsights}.
-In order to meet the strict CO$_2$ emissions targets for the UK electrical supply by 2050, and bearing in mind that nuclear power capacity is not predicted to grow until after 2030, either carbon capture or significantly greater renewable power penetration will be required of the supply side\cite{FES2018}.
+The fluctuating output from wind and solar is topped up to match demand primarily by varying output from fossil fuel power thermal plant.
+In the UK, this is specifically achieved by gas turbines and to a lesser extent coal plant \cite{DraxElectricInsights}.
+In order to meet the strict CO$_2$ emissions targets for the UK electrical supply by 2050, and bearing in mind that nuclear power capacity is not predicted to grow until after 2030, either carbon capture or significantly greater renewable power penetration will be required at the supply side \cite{FES2018}.
 
-In economic terms, the price of electricity at a given moment is set by the operating costs and  of the most expensive generator in the merit order \cite{Staffell}.
+In economic terms, the price of electricity at a given moment is set by the operating costs and finance requirements of the least economic generator that must be called upon \cite{Staffell2016}.
 As wind and solar have zero marginal cost, increasing penetration of these power sources will reduce electrical price according to a stochastic distribution.
-At the same time, increasing operating costs for dispatchable fossil generation due to reduced capacity factors and carbon taxation will increase the price in some periods.
+At the same time, the capacity factors of the dispatchable fossil plant will decrease, and hence a premium is added to the operating cost in order to cover CAPEX and other costs.
+In Europe, the operating cost of fossil generation is further increased by the required purchase of credits under the Emmisions Trading System \cite{Staffell2018}.
+The overall effect of these developments would be an increase in price variability, although this has been mitigated in practice by markets for capacity, where generators may supplement income from energy sales with payments for availability of power capacity \cite{ENGIE}.
 
-Such price variability presents an increasing opportunity for arbitrage by energy storage systems (ESS).
+Such price variability presents an opportunity for arbitrage by energy storage systems (ESS), which may charge and discharge in periods of low and high price respectively.
 At present however, revenue from arbitrage on the electrical price alone is not sufficient to justify BESS investment, even under optimistic economic assumptions \cite{Hu2010}.
@@
 Such frequency regulation services attract Li-ion systems designed for power, such as the Hornsdale Power Reserve project in South Australia, which is specified for \SI{1}{\hour}\SI{17}{\minute} at its \SI{100}{\mega\watt} rated power output \cite{DOEGlobalESSDatabase}.
-This energy to power ratio is representative of global Li-ion installation data as collated by \cite{Newbery2018}.
+This energy to power ratio is representative of global grid-scale Li-ion installation data as collated by \cite{Newbery2018}.
 
@@
 The de-rating developments would disfavour the Li-ion systems that have been deployed so far and favour systems with higher energy to power ratios.
-Redox flow battery (RFB) systems and promising candidates for multi-hour applications and revenue stacking, as the incremental cost of energy capacity is low, and systems deployed so far have an average energy to power ratio of 3.5 \cite{Newbery2018}.
+Redox flow battery (RFB) systems and promising candidates for multi-hour applications and revenue stacking, as the incremental cost of energy capacity is low.
+Systems deployed so far have an average energy to power ratio of 3.5 \cite{Newbery2018}.
 
@@
 By repeating this process at different energy and power ratings, the authors found that the economically optimal VRFB rating for the \SI{50}{\mega\watt} wind farm was \SI{5.3}{\mega\watt} and \SI{3}{\mega\watt\hour}.
-This illustrates the point made above regarding the poor economics of wholesale market arbitrage - the optimal RFB rating is dictated by the frequency market revenue and the reserve energy requirement.
+This supports the point made above regarding the poor economics of wholesale market arbitrage - the optimal RFB rating is dictated by the frequency market revenue and the reserve energy requirement.
 Increasing the capacity of the VRFB in order to perform additional wholesale market arbitrage is not economical.
@@
 
-\cite{Hu2010} optimised the schedule of a vanadium RFB (VRFB) and a poly-sulphide bromine RFB in order to maximise arbitrage revenue on the Danish day ahead electrical spot market.
+\cite{Hu2010} optimised the schedule of a VRFB and a poly-sulphide bromine RFB in order to maximise arbitrage revenue on the Danish day ahead electrical spot market.
 This approach was not strictly LP, as discussed in \cref{Model_Formulation_LP_Reference_Current_Density_Terms}, but the problem may be reposed in LP form.
@@
 
+Accurate treatment of efficiency is important, as losses impact the revenue available from arbitrage.
 The key innovation of the present work is the introduction of a RFB schedule optimisation framework posed in terms of current-density rather than power.
@@
 }
-\caption{(a) Optimal VRFB schedules derived for the 22\textsuperscript{nd} August 2017 using NLP formulations where OCV is fixed $OCV_{50\%}$ and functionalised.
+\caption{(a) Optimal VRFB schedules derived for the 22\textsuperscript{nd} August 2017 using NLP formulations where OCV is fixed as $OCV_{50\%}$ and functionalised.
 (b)Optimal values of OCV where it is functionalised.}
@@
 However the global optimum requires that the RFB be at a low SOC in these hours, and hence at a low OCV.
-The low OCV means that a higher current-density would be required to match the power output in the fixed OCV case, and hence higher voltaic losses would be incurred, eroding the revenue.
+The low OCV means that a higher current-density would be required to obtain a given pwoer outpute, and hence higher voltaic losses would be incurred, eroding the revenue.
 On this date the impact on revenue is low; the optimal schedule with consideration of functional OCV results in a revenue of \pounds 0.062 compared to \pounds 0.063 for the base case.

### 2018-11-07 Paper/main.tex
Update main.tex

@@
 \label{sec:Intro}
-Worldwide, the need to reduce CO$_2$ emissions has stimulated the development of a wide range of renewable energy sources.
+Worldwide, the need to reduce CO$_2$ emissions is stimulatling the development of a wide range of renewable energy sources.
 Presently, renewable energy is estimated to account for 10\% of global energy consumption, and this is forecast to grow to 12\% by 2023.
@@
 At the same time, the capacity factors of the dispatchable fossil plant will decrease, and hence a premium is added to the operating cost in order to cover CAPEX and other costs.
-In Europe, the operating cost of fossil generation is further increased by the required purchase of credits under the Emmisions Trading System \cite{Staffell2018}.
+Across the EU, the operating cost of fossil generation is further increased by the required purchase of credits under the Emissions Trading System \cite{Staffell2018}.
 The overall effect of these developments would be an increase in price variability, although this has been mitigated in practice by markets for capacity, where generators may supplement income from energy sales with payments for availability of power capacity \cite{ENGIE}.
 
-Such price variability presents an opportunity for arbitrage by energy storage systems (ESS), which may charge and discharge in periods of low and high price respectively.
-At present however, revenue from arbitrage on the electrical price alone is not sufficient to justify BESS investment, even under optimistic economic assumptions \cite{Hu2010}.
+Electrical price variability presents an opportunity for arbitrage by energy storage systems (ESS), which may charge and discharge in periods of low and high price respectively.
+At present however, this revenue alone is not sufficient to justify BESS investment, even under optimistic economic assumptions \cite{Hu2010}.
 Hence an important aspect of BESS economics is revenue stacking, where multiple revenue streams are captured by the same system.
@@
 
-A similar de-rating has also been applied recently in the UK capacity market (CM).
-This was done due to concerns that the average storage duration of the BESS contracted for CM is shorter than the \SI{2}{\hour} that a stress event may last for.
+A similar de-rating has also been applied recently in the UK capacity market.
+This was done due to concerns that the average storage duration of the BESS contracted for the service was shorter than the \SI{2}{\hour} that a stress event may last for.
 \cite{NGDeratingFactors}.
@@
 \subfloat[][]{
-\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./IDD2s_schedules.pdf}
+\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./figures/V_losses_schedules.pdf}
 \label{fig:IDD2s_opt_schedules}
@@
 \subfloat[][]{
-\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./IDD2s_schedules_SOC.pdf}
+\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./figures/V_losses_SOC.pdf}
 \label{fig:IDD2s_opt_schedules}
@@
 \label{fig:OCV_fixed_dynamic_schedules}
-\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./OCV_fixed_dynamic_schedules.pdf}
+\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./figures/OCV_fixed_dynamic_schedules.pdf}
 
@@
 \label{fig:OCV}
-\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./OCV_fixed_dynamic_OCV.pdf}
+\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./figures/OCV_fixed_dynamic_OCV.pdf}
 }
@@
 
-\subsection{Cell-Voltage constraint}
+\subsection{NLP with functional OCV and Cell-Voltage constraint}
 \label{Results_Cell_Voltage_Constraint}
-In the NLP formulations given in \cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable,eqn: NLP_true_OCV_long_form} it is assumed that the RFB may be charged at maximum current-density at the maximum SOC.
-This was the charging strategy applied in \cite{Reed2016}, where the cell-voltage was allowed to reach \SI{1.85}{\volt} at the highest current-density, corresponding to an approximate SOC of 0.85.
+In the NLP formulations given in \cref{Model_Formulation_NLP_treatment_voltaic_losses,Model_Formulation_Functional_Treatment_OCV} it is assumed that the RFB may be charged at maximum current-density at the maximum SOC.
+This was the charging strategy applied by \cite{Reed2016}, where the cell-voltage was allowed to reach \SI{1.85}{\volt} at the highest current-density, corresponding to an approximate SOC of 0.85.
 Overcharging may lead to oxygen evolution at the cathode and hydrogen evolution at the anode.
@@
 
-With the introduction of the functional OCV in \cref{eqn: NLP_true_OCV_long_form}, the application of a maximum cell-voltage constraint is now rational.
+With the introduction of the functional OCV in \cref{eqn: NLP_true_OCV_long_form}, the application of a maximum cell-voltage constraint is now possible.
 In this case study $V_{max}$ in \cref{eqn: Cell_Voltage_Constraint} was set at 1.65, just below the voltage at which \cite{Wei2017} reported evolution of macroscopic bubbles of hydrogen and oxygen (\SI{1.70}{\volt}).
@@
 \label{fig:V_max_schedules_no_effect_day}
-\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./V_max_schedules_no_effect_day.pdf}
+\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./figures/V_max_schedules_no_effect_day.pdf}
 }
@@
 \label{fig:V_max_voltages_no_effect_day}
-\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./V_max_voltages_no_effect_day.pdf}
+\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./figures/V_max_voltages_no_effect_day.pdf}
 }
@@
 \label{fig:V_max_schedules}
-\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./V_max_schedules.pdf}
+\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./figures/V_max_schedules.pdf}
 }
@@
 \label{fig:V_max_voltages}
-\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./V_max_voltages.pdf}
+\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./figures/V_max_voltages.pdf}
 }
@@
 The application of the linear constraint on cell-voltage increases the time to process the 2017 data from \SI{217}{\second} to \SI{232}{\second}.
+
+An important aspect of RFB economics is the variability of daily price profile across the year.
+Figure XX shows composite day-ahead price data for the winter and summer seasons in 2017.
+In summer months, the availability of cheap solar power depresses the price around midday, and the peak evening demand is low, resulting in two peaks of similar height.
+In winter, the solar price depression is not so marked, and the price in the evening peak becomes higher, hence the profile is more like a single peak.
+
+\cref{fig:winter_day_schedules,fig:summer_day_schedules} show the optimal LP and NLP schedules for example winter and summer days from 2017.
+Although the revenue obtained is \pounds 0.074 for both days under the NLP optimal schedule, the required throughput of the battery is greater on the summer day.
+This has an impact on the error of the LP revenue estimation - for the winter day the revenue is estimated at \pound 0.088, whereas on the summmer day it is estimated at \pound 0.099.
+The error in the estimation of revenue under the LP optimisation is dependent on energy throughput.
+
+\begin{figure}[!ht]
+\centering
+\subfloat[][]{
+\label{fig:winter_day_schedules}
+\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./figures/winter_day_schedules.pdf}
+}
+\subfloat[][]{
+\label{fig:summer_day_schedules}
+\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./figures/summer_day_schedules.pdf}
+}
+\caption{Optimal schedules by LP and NLP optimisation for (a) Winter day with pronounced evening price peak (b) Summer day with afternoon price depression associated with solar power}
+\end{figure}
 

### 2018-11-14 Paper/main.tex
Update main.tex

@@
 
-\cref{fig:winter_day_schedules,fig:summer_day_schedules} show the optimal LP and NLP schedules for example winter and summer days from 2017.
+\cref{fig:winter_day_schedules,fig:summer_day_schedules} show the optimal LP and NLP (\cref{Model_Formulation_Constraint_Cell_Voltage}) schedules for example winter and summer days from 2017.
 Although the revenue obtained is \pounds 0.074 for both days under the NLP optimal schedule, the required throughput of the battery is greater on the summer day.

### 2018-11-16 ESS_PV_sizing_framework.py
Add files via upload

@@ -1,0 +1,263 @@
+﻿# coding: utf-8
+"""This script imports price, load and solar data for an electrical retail customer, then uses LP to minimise costs by
+optimal ESS scheduling. PV power goes by default to suppply load, with any excess being split between export and ESS
+charging, with a variable, r, denoting the fraction going to the ESS.
+
+The optimisation is performed within a triple loop, to probe:
+* PV installed power
+* ESS installed power
+* ESS installed energy
+Author: Diarmid Roberts. Date: 4th October 2018"""
+########################################################################################################################
+# Import required tools
+########################################################################################################################
+
+# import off-the-shelf scripting modules
+from __future__ import division     # Without this, rounding errors occur in python 2.7, but apparently not in 3.4
+import pandas as pd                 # CSV handling
+import datetime                     # For time stamping optimisation process
+
+#Import price and load data
+price_data = pd.read_csv('price.csv')
+load_data = pd.read_csv('load.csv')
+solar_data = pd.read_csv('solar.csv')
+# import PYOMO
+from pyomo.environ import *
+from pyomo.opt import SolverFactory
+
+########################################################################################################################
+# User entered parameters
+########################################################################################################################
+# Constraints on installed ratings of ESS and PV
+AvailableArea = 1800                # m2 available for PV installation
+#MaxPowerPV = AvailableArea * 0.150  # Convert area to peak power assuming a certain kWp/m2
+MaxPowerPV = 150                    # Set to be double max load, so export not greater than current peak load
+MaxPowerESS = 30                    # Set a maximum for ESS power rating (kW)
+MaxDurationESS = 4                  # Maximum energy to power ratio of ESS (h)
+EffESS = 0.9                        # Round trip energy efficiency of ESS
+FiT = 0.042                         # FiT for PV (£/kWh)
+ExportTariff = 0.052                # FiT export Tariff (£/kWh)
+
+# Define granularity of sizing matrix (fraction of max install)
+StepESSPower = 0.5
+StepESSDuration = 0.2
+StepSolar = 0.5
+
+########################################################################################################################
+# Optimisation Framework
+########################################################################################################################
+# This loop runs through different PV installation sizes
+df_by_PV_install = pd.DataFrame()
+for s in range(0,int(1+1/StepSolar)):
+    PVRating = s * StepSolar * MaxPowerPV
+    print('PV', PVRating, 'kW')
+    # this loop goes through different ESS power ratings
[…]

### 2018-11-16 pv_ess_behind_meter/ESS_PV_sizing_framework.py
Create ESS_PV_sizing_framework.py

@@ -1,0 +1,257 @@
+# coding: utf-8
+"""This script imports price, load and solar data for an electrical retail customer, then uses LP to minimise costs by
+optimal ESS scheduling. PV power goes by default to suppply load, with any excess being split between export and ESS
+charging, with a variable, r, denoting the fraction going to the ESS.
+
+The optimisation is performed within a triple loop, to probe:
+* PV installed power
+* ESS installed power
+* ESS installed energy
+Author: Diarmid Roberts. Date: 4th October 2018"""
+########################################################################################################################
+# Import required tools
+########################################################################################################################
+
+# import off-the-shelf scripting modules
+from __future__ import division     # Without this, rounding errors occur in python 2.7, but apparently not in 3.4
+import pandas as pd                 # CSV handling
+import datetime                     # For time stamping optimisation process
+
+#Import price and load data
+price_data = pd.read_csv('price.csv')
+load_data = pd.read_csv('load.csv')
+solar_data = pd.read_csv('solar.csv')
+# import PYOMO
+from pyomo.environ import *
+from pyomo.opt import SolverFactory
+
+########################################################################################################################
+# User entered parameters
+########################################################################################################################
+# Constraints on installed ratings of ESS and PV
+AvailableArea = 1800                # m2 available for PV installation
+#MaxPowerPV = AvailableArea * 0.150  # Convert area to peak power assuming a certain kWp/m2
+MaxPowerPV = 150                    # Set to be double max load, so export not greater than current peak load
+MaxPowerESS = 30                    # Set a maximum for ESS power rating (kW)
+MaxDurationESS = 4                  # Maximum energy to power ratio of ESS (h)
+EffESS = 0.9                        # Round trip energy efficiency of ESS
+FiT = 0.042                         # FiT for PV (£/kWh)
+ExportTariff = 0.052                # FiT export Tariff (£/kWh)
+
+# Define granularity of sizing matrix (fraction of max install)
+StepESSPower = 0.5
+StepESSDuration = 0.2
+StepSolar = 0.5
+
+########################################################################################################################
+# Optimisation Framework
+########################################################################################################################
+# This loop runs through different PV installation sizes
+df_by_PV_install = pd.DataFrame()
+for s in range(0,int(1+1/StepSolar)):
+    PVRating = s * StepSolar * MaxPowerPV
+    print('PV', PVRating, 'kW')
+    # this loop goes through different ESS power ratings
[…]

### 2018-11-16 ESS_PV_sizing_framework.py
Delete ESS_PV_sizing_framework.py

@@ -1,263 +1,0 @@
-﻿# coding: utf-8
-"""This script imports price, load and solar data for an electrical retail customer, then uses LP to minimise costs by
-optimal ESS scheduling. PV power goes by default to suppply load, with any excess being split between export and ESS
-charging, with a variable, r, denoting the fraction going to the ESS.
-
-The optimisation is performed within a triple loop, to probe:
-* PV installed power
-* ESS installed power
-* ESS installed energy
-Author: Diarmid Roberts. Date: 4th October 2018"""
-########################################################################################################################
-# Import required tools
-########################################################################################################################
-
-# import off-the-shelf scripting modules
-from __future__ import division     # Without this, rounding errors occur in python 2.7, but apparently not in 3.4
-import pandas as pd                 # CSV handling
-import datetime                     # For time stamping optimisation process
-
-#Import price and load data
-price_data = pd.read_csv('price.csv')
-load_data = pd.read_csv('load.csv')
-solar_data = pd.read_csv('solar.csv')
-# import PYOMO
-from pyomo.environ import *
-from pyomo.opt import SolverFactory
-
-########################################################################################################################
-# User entered parameters
-########################################################################################################################
-# Constraints on installed ratings of ESS and PV
-AvailableArea = 1800                # m2 available for PV installation
-#MaxPowerPV = AvailableArea * 0.150  # Convert area to peak power assuming a certain kWp/m2
-MaxPowerPV = 150                    # Set to be double max load, so export not greater than current peak load
-MaxPowerESS = 30                    # Set a maximum for ESS power rating (kW)
-MaxDurationESS = 4                  # Maximum energy to power ratio of ESS (h)
-EffESS = 0.9                        # Round trip energy efficiency of ESS
-FiT = 0.042                         # FiT for PV (£/kWh)
-ExportTariff = 0.052                # FiT export Tariff (£/kWh)
-
-# Define granularity of sizing matrix (fraction of max install)
-StepESSPower = 0.5
-StepESSDuration = 0.2
-StepSolar = 0.5
-
-########################################################################################################################
-# Optimisation Framework
-########################################################################################################################
-# This loop runs through different PV installation sizes
-df_by_PV_install = pd.DataFrame()
-for s in range(0,int(1+1/StepSolar)):
-    PVRating = s * StepSolar * MaxPowerPV
-    print('PV', PVRating, 'kW')
-    # this loop goes through different ESS power ratings
[…]

### 2018-11-19 Paper/main.tex

@@
 
-\title{The importance of fidelity in the scheduling of Redox Flow Batteries.}
+\title{Improving the Fidelity of Redox Flow Battery Schedule Optimisation for more Accurate Techno-economic Analyses.}
 
@@
 \begin{abstract}
-Redox Flow Battery (RFB) systems are promising technologies for the multi-hour electrical energy storage that would be necessary in a future low carbon emission UK electrical grid based on wind and solar power.
-In order to perform techno-economic analyses (TEA) on these systems, it is necessary to calculate potential revenue for realistic applications, ideally under optimal operation (scheduling).
-In the present work, a novel schedule optimisation framework is introduced in which coulombic and voltaic components are separated, and the problem variablised in terms of current-density rather than power.
-This approach allows the expression of ohmic losses as a convex quadratic function of current-density in the objective function while maintaining all linear constraints.
-The resultant non-linear problem is more readily solvable than previous expressions in terms of power.
-It is demonstrated for an arbitrage application involving a vanadium RFB that the consideration of variable voltaic losses results in a significant improvement to the accuracy of revenue estimation when compared to the typical linear programming approach, where a fixed efficiency is assumed.
-The framework allows the removal of the assumption of a representative fixed open-cell voltage, which further improves revenue estimation, at the cost of computability.
-In a third case study, the framework allows a linear constraint on the working cell-voltage, resulting in current-density restriction at high state of charge (SOC) in order to avoid hydrogen and oxygen evolution.
-This feature would be relevant for the optimisation of Li-ion battery scheduling, where voltage is a very important consideration.
-The introduced framework is intended to assist technical optimisation of future RFB systems, for example in determining the optimum pump or cooling system size.
-The framework allows product optimisation in terms of revenue under optimal scheduling against a price signal (rather than metrics such as system cost, or lifetime cost per unit energy throughput as in previous works).
+Redox Flow Battery (RFB) systems are promising technologies for the multi-hour electrical energy storage that would be necessary in a future low carbon emission  electrical grids based on wind and solar power.
+In order to perform techno-economic analyses (TEA) on these systems it is necessary to calculate potential revenue for realistic applications, ideally under optimal operation (scheduling).
+In this work a novel schedule optimisation framework is introduced in which coulombic and voltaic components are separated, and the problem variablised in terms of current-density rather than power.
+This framework allows the expression of ohmic losses as a convex quadratic function of current-density in the objective function while maintaining all linear constraints.
+It is demonstrated for an arbitrage application involving a vanadium RFB that the consideration of variable voltaic losses results in a significant improvement to the accuracy of revenue estimation when compared to the typical linear programming (LP) approach where a fixed efficiency is assumed.
+The quadratic formulation is more readily solvable than alternative approaches reported in the literature, with minimal increase to solve time compared to the LP approach.
+The fidelity of the RFB representation is further increased by expressing open cell voltage as a function of state of charge (SOC).
+At low SOC, the current-density required for a given power is greater, and hence the ohmic losses are higher.
+In a third development, a linear constraint is applied to the working cell-voltage, resulting in current-density restriction at high SOC in order to avoid hydrogen and oxygen evolution.
+This feature would also be useful for schedule optimisation of Li-ion batteries, where voltage is an important consideration.
+The introduced framework is intended to assist techno-economic optimisation of future RFB systems, for example in determining the optimum pump or cooling system size for a particular application.
+The framework allows system optimisation in terms of revenue under optimal scheduling against a price signal rather than metrics such as system cost, or lifetime cost per unit energy throughput as in previous works.
 
@@
 \begin{keyword}
-Redox flow Battery \sep Optimisation \sep Non-linear Programming \sep Techno-Economic Analysis \sep Scheduling \sep VRFB  
+Redox Flow Battery \sep Optimisation \sep Non-linear Programming \sep Techno-Economic Analysis \sep Scheduling \sep VRFB  
 %% keywords here, in the form: keyword \sep keyword
@@
 \label{sec:Intro}
-Worldwide, the need to reduce CO$_2$ emissions is stimulatling the development of a wide range of renewable energy sources.
+Worldwide, the need to reduce CO$_2$ emissions is stimulating the development of a wide range of renewable energy sources.
 Presently, renewable energy is estimated to account for 10\% of global energy consumption, and this is forecast to grow to 12\% by 2023.
@@
 This should be done at various energy and power ratings (as these are decoupled in a non-hybrid RFB) to find the global optimum, as described by \cite{Oudalov2007}.
-The optimisation of operation requires determining the schedule, that is, the power output of the battery in each sub-period that maximises benefit without violating energy conservation.
+The optimisation of operation requires determining the schedule, that is the power output of the battery in each sub-period, that maximises benefit without violating energy conservation.
 
-Where a price signal exists optimisation is commonly performed by posing a linear programming (LP) problem \emph{do I dump here a list of all the LP references i will refer to below?}.
+Where a price signal exists optimisation is commonly performed by posing a linear programming (LP) problem.
 In this method, the objective function and the constraints on the system are expressed as linear algebraic terms.
@@
 \cite{Hu2010} optimised the schedule of a VRFB and a poly-sulphide bromine RFB in order to maximise arbitrage revenue on the Danish day ahead electrical spot market.
-This approach was not strictly LP, as discussed in \cref{Model_Formulation_LP_Reference_Current_Density_Terms}, but the problem may be reposed in LP form.
+This approach was not strictly LP, but the problem may be reposed in LP form.
 
@@
 The problem therefore becomes convex (quadratic), but non-linear in both objective function and constraint.
-The authors then employ a linear approximation technique\sol{does this need a reference}, and combine the efficiency treatment with a piece-wise linear treatment of degradation as a function of power, giving an overall mixed integer-linear programming problem (MILP).
+The authors then employ a linear approximation technique\sol{does this need a reference?
+DR: do you mean for the technique used?}, and combine the efficiency treatment with a piece-wise linear treatment of degradation as a function of power, giving an overall mixed integer-linear programming problem (MILP).
 
@@
 This allows the separation of the voltaic and coulombic components of the power input/output.
-Using the introduced framework, three novel augmentations \emph{(add-ons?)} to the typical scheduling problem are posed while maintaining an easily computable problem:
+Using the introduced framework, three improvements to the RFB representation in the schedule optimisation are posed while maintaining an easily computable problem:
 
@@
 
-The importance of these improvements to the fidelity of revenue estimation is demonstrated in a day-ahead price arbitrage case study.
+The importance of these improvements to the accuracy of revenue estimation is demonstrated in a day-ahead price arbitrage case study.
 
@@
 
-The SOC of the RFB is defined after \cite{Gomes2017} by:
+The SOC of the RFB at time $t$ is defined after \cite{Gomes2017} by:
 
@@
 The low OCV means that a higher current-density would be required to obtain a given pwoer outpute, and hence higher voltaic losses would be incurred, eroding the revenue.
-On this date the impact on revenue is low; the optimal schedule with consideration of functional OCV results in a revenue of \pounds 0.062 compared to \pounds 0.063 for the base case.
-Across 2017 the formulation with functional OCV returns \pounds 26.40 compared to \pounds 27.81 for the fixed OCV case.
+On this date the impact on revenue is low; the optimal schedule with consideration of functional OCV results in a revenue of \pounds \SI{0.062}{\per\kilo\watt} compared to \pounds \SI{0.063}{\per\kilo\watt} for the base case.
+Across 2017 the formulation with functional OCV returns \pounds \SI{26.40}{\per\kilo\watt} compared to \pounds \SI{27.81}{\per\kilo\watt} for the fixed OCV case.
 The assumption of a fixed OCV (parametrised at 50\% SOC) therefore results in a small but not insignificant overestimation of revenue in this application.
@@
 
-\cref{fig:V_max_schedules_no_effect_day} shows the optimal scheduling with and without voltage constraint for the median revenue day 22\textsuperscript{nd} August 2017, and \cref{fig:V_max_voltages_no_effect_day} the corresponding cell voltage profiles, showing a maximum cell-voltage of \SI{1.62}{\volt} and hence no violation in the unconstrained schedule.
-
-\begin{figure}[!ht]
-\centering
-\subfloat[][]{
-\label{fig:V_max_schedules_no_effect_day}
-\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./figures/V_max_schedules_no_effect_day.pdf}
-}
-\subfloat[][]{
-\label{fig:V_max_voltages_no_effect_day}
-\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./figures/V_max_voltages_no_effect_day.pdf}
-}
-\caption{(a) Optimal VRFB schedules derived for the 22\textsuperscript{nd} August 2017 by maximising \cref{eqn: NLP_true_OCV_long_form} with no cell-voltage constraint, and subject to the cell voltage constraint given in \cref{eqn: Cell_Voltage_Constraint}.
-(b) Corresponding cell-voltage profiles.}
-\end{figure}
-
-\cref{fig:V_max_schedules} shows the optimal schedules with and without constraint for 5\textsuperscript{th} February 2017, where the difference in revenue was greatest (\pounds 0.018 vs.
-\pounds 0.020).
+\cref{fig:V_max_schedules} shows the optimal schedules with and without constraint for 5\textsuperscript{th} February 2017, where the difference in revenue was greatest (\pounds \SI{0.18}{\per\kilo\watt} vs.
+\pounds \SI{0.20}{\per\kilo\watt}).
 \cref{fig:V_max_voltages} shows the corresponding cell-voltage profiles.
@@
 
-In the unconstrained case the RFB charges at the maximum current-density of \SI{3200}{\ampere\per\square\meter} in hour 19 in order to capitalise on the unusually low electrical price, resulting in a cell-voltage of \SI{1.71}{\volt}.
+In the unconstrained case the RFB charges at the maximum current-density of \SI{3200}{\ampere\per\square\meter} in hour 20 in order to capitalise on the unusually low electrical price, resulting in a cell-voltage of \SI{1.71}{\volt}.
 In the constrained case, the current-density is restricted to \SI{2213}{\ampere\per\square\meter} in the same period, and the RFB must charge more (and discharge less) in other periods to satisfy \cref{eqn: Method_Constraint_E_Conservation}.
@@
 The application of the linear constraint on cell-voltage increases the time to process the 2017 data from \SI{217}{\second} to \SI{232}{\second}.
-
+\subsection{Impact of price profile on optimisation results}
+\label{Results_impact_of_price_profile}
 An important aspect of RFB economics is the variability of daily price profile across the year.
-Figure XX shows composite day-ahead price data for the winter and summer seasons in 2017.
+\Cref{fig:DJF_2017_N2EX,fig:JJA_2017_N2EX} show composite day-ahead price data for the winter and summer seasons in 2017.
 In summer months, the availability of cheap solar power depresses the price around midday, and the peak evening demand is low, resulting in two peaks of similar height.
@@
 
+\begin{figure}[!ht]
+\centering
+\subfloat[][]{
+\label{fig:JJA_2017_N2EX}
+\includegraphics[trim = 13.8cm 3.5cm 10.3cm 4.5cm, clip, width = .4\textwidth]{./figures/JJA_2017.pdf}
+}
+\subfloat[][]{
+\label{fig:DJF_2017_N2EX}
+\includegraphics[trim = 13.8cm 3.5cm 10.3cm 4.5cm, clip, width = .4\textwidth]{./figures/DJF_2017.pdf}
+}
+\caption{Composite hourly N2EX day-ahead price data for a) June, July and August 2017 and b) December,January, February 2017}
+\end{figure}
+
 \cref{fig:winter_day_schedules,fig:summer_day_schedules} show the optimal LP and NLP (\cref{Model_Formulation_Constraint_Cell_Voltage}) schedules for example winter and summer days from 2017.
-Although the revenue obtained is \pounds 0.074 for both days under the NLP optimal schedule, the required throughput of the battery is greater on the summer day.
-This has an impact on the error of the LP revenue estimation - for the winter day the revenue is estimated at \pound 0.088, whereas on the summmer day it is estimated at \pound 0.099.
-The error in the estimation of revenue under the LP optimisation is dependent on energy throughput.
+Although the revenue obtained is \pounds \SI{0.074}{\per\kilo\watt} for both days under the NLP optimal schedule, the throughput of the RFB is greater on the summer day.
+This has an impact on the error of the LP revenue estimation; for the winter day the revenue is estimated at \pounds \SI{0.088}{\per\kilo\watt}, whereas on the summmer day it is estimated at \pounds \SI{0.099}{\per\kilo\watt}.
+The error in the estimation of revenue using the LP optimisation is therefore dependent on the application.
 

### 2018-11-19 Paper/main.tex
Update main.tex

@@
 In reality however, the voltaic losses that would occur under the LP schedule would reduce the revenue to \pounds 0.051.
-Across 2017, LP scheduling would have resulted in \pounds 22.61, against a prediction of \pounds 33.16, and NLP scheduling would have yielded \pounds 27.99.
+Across 2017, LP scheduling would have resulted in \pounds \SI{22.61}{\per\kilo\watt}, against a prediction of \pounds \SI{33.16}{\per\kilo\watt}, and NLP scheduling would have yielded \pounds \SI{27.99}{\per\kilo\watt}.
 
@@
 
-Maximising \cref{eqn: NLP_true_OCV_long_form} subject to \cref{eqn: Cell_Voltage_Constraint} for the 2017 data resulted in \pounds 26.07 compared to \pounds 26.40 when the cell voltage was not constrained.
+Maximising \cref{eqn: NLP_true_OCV_long_form} subject to \cref{eqn: Cell_Voltage_Constraint} for the 2017 data resulted in \pounds \SI{26.07}{\per\kilo\watt} compared to \pounds \SI{26.40}{\per\kilo\watt} when the cell voltage was not constrained.
 The difference is small because high current-density is already discouraged by the voltaic losses in \cref{eqn: NLP_true_OCV_long_form}.
@@
 This assumption is supported by the data reported by \cite{Reed2016}, but the behaviour below \SI{1600}{\ampere\per\square\meter} was not reported.
-In principle shunt current losses are proportional to cell-voltage, as described by \cite{Xing2011}, and hence a function of SOC and over-voltage (via \cref{eqn: Cell_Voltage_Constraint}.\emph{wait, is it the terminal voltage or the OCV that drives the shunt current?} The dependence of coulombic losses on current-density is therefore weaker than implied by the present assumption.
+In principle shunt current losses are proportional to cell-voltage, as described by \cite{Xing2011}, and hence a function of SOC and over-voltage (via \cref{eqn: Cell_Voltage_Constraint}.
+The dependence of coulombic losses on current-density is therefore weaker than implied by the present assumption.
 As a first approximation, a fixed loss term could be placed in the SOC expression in \cref{eqn: Method_SOC_Tracker} instead of the $\sqrt{\eta_C}$ coefficients.
@@
 As discussed above, both the coulombic losses and the pumping power may be better represented as fixed terms.
-In this case, the efficiency of the system at low power input/output will be unfavourable (as demonstrated by \cite{Nguyen2014}).
+In this case, the efficiency of the system at low power input/output will be unfavourable as demonstrated by \cite{Nguyen2014}.
 An issue with representing these losses as fixed terms, is that they would be present at all times in the schedule.
@@
 up to the AC-DC boundary), for example in the optimal specification of pump rating, or cooling system for a particular application.
-RFB systems have previously been the subject of technical optimisation, for example the study of the trade-offs between electrolyte concentration and ASR (\cite{Weber2013} \emph{this should be weber et al.}), or shunt currents and pumping losses \citep{Viswanathan2014}.
+RFB systems have previously been the subject of technical optimisation, for example the study of the trade-offs between electrolyte concentration and ASR \citep{Weber2013}, or shunt currents and pumping losses \citep{Viswanathan2014}.
 In these works, a simple metric such as levelised cost of electrical energy under a simple duty cycle is used as the objective to minimise.

### 2018-11-19 Paper/main.tex
Update main.tex

@@
 This is the consequence of factoring variable ohmic losses in the objective function, where high current-density is penalised by higher losses.
-Under the LP formulation, the VRFB is predicted to generate \pounds 0.068, whereas the NLP formulation predicts a revenue of \pounds 0.063.
+Under the LP formulation, the VRFB is predicted to generate \pounds \SI{0.068}{\per\kilo\watt}, whereas the NLP formulation predicts a revenue of \pounds \SI{0.063}{\per\kilo\watt}.
 In reality however, the voltaic losses that would occur under the LP schedule would reduce the revenue to \pounds 0.051.

### 2018-11-21 Paper/main.tex
shortened introduction and added comments

@@
 \newcommand{\sol}[1]{\footnote{#1}\marginpar{\fbox{\thefootnote}}}
-%\newcommand{\esf}[1]{}
 
@@
 Worldwide, the need to reduce CO$_2$ emissions is stimulating the development of a wide range of renewable energy sources.
-Presently, renewable energy is estimated to account for 10\% of global energy consumption, and this is forecast to grow to 12\% by 2023.
+Presently, renewable energy is estimated to account for 10\% of global energy consumption, and this is forecast to grow to 12\% by 2023 \sol{you need a reference for this}.
 The growth of renewable energy is forecast to be highest in the electrical power sector, from 24\% in 2017 to 30\% in 2023.
-Although hydro power presently accounts for 50\% of renewable power generation, the  intermittent power sources wind and solar are forecast to contribute most to this growth\cite{IEA2018}.
-A number of large European economies, such as Germany, UK, Italy and Spain obtained 23\% to 30\% of there electrical energy from non-hydro renewable power in 2017 \cite{BP2018}.
+Although hydro power presently accounts for 50\% of renewable power generation, the  intermittent power sources wind and solar are forecast to contribute most to this growth\cite{IEA2018}\sol{is this reference for all of the data before this?
+it would be better to have a couple more supporting this...}.
+Indeed, a number of large European economies, such as Germany, UK, Italy and Spain obtained 23\% to 30\% of there electrical energy from non-hydro renewable power in 2017 \cite{BP2018}.
 
-The fluctuating output from wind and solar is topped up to match demand primarily by varying output from fossil fuel power thermal plant.
-In the UK, this is specifically achieved by gas turbines and to a lesser extent coal plant \cite{DraxElectricInsights}.
-In order to meet the strict CO$_2$ emissions targets for the UK electrical supply by 2050, and bearing in mind that nuclear power capacity is not predicted to grow until after 2030, either carbon capture or significantly greater renewable power penetration will be required at the supply side \cite{FES2018}.
+To maintain stability of the network, the fluctuating output from wind and solar is supplemented to match demand primarily by varying output from fossil fuel power thermal plant.
+In the UK, this is specifically achieved by gas turbines and to a lesser extent coal plant \cite{DraxElectricInsights}.% In order to meet the strict CO$_2$ emissions targets for the UK electrical supply by 2050, and bearing in mind that nuclear power capacity is not predicted to grow until after 2030, either carbon capture or significantly greater renewable power penetration will be required at the supply side \cite{FES2018}.
 
 In economic terms, the price of electricity at a given moment is set by the operating costs and finance requirements of the least economic generator that must be called upon \cite{Staffell2016}.
-As wind and solar have zero marginal cost, increasing penetration of these power sources will reduce electrical price according to a stochastic distribution.
+As wind and solar have zero marginal cost, increasing penetration of these power sources will reduce electrical price according to a stochastic distribution\sol{I'm not sure what this last bit means}.
 At the same time, the capacity factors of the dispatchable fossil plant will decrease, and hence a premium is added to the operating cost in order to cover CAPEX and other costs.
 Across the EU, the operating cost of fossil generation is further increased by the required purchase of credits under the Emissions Trading System \cite{Staffell2018}.
-The overall effect of these developments would be an increase in price variability, although this has been mitigated in practice by markets for capacity, where generators may supplement income from energy sales with payments for availability of power capacity \cite{ENGIE}.
+The overall effect of these developments would be an increase in price variability, although this has been mitigated in practice by markets for capacity, where generators may supplement income from energy sales with payments for availability of power capacity \cite{ENGIE}.\sol{can we cut this paragraph down}
 
@@
 
-At the shortest timescale, BESS may receive payments from the grid operator to respond to fluctuations in the AC grid frequency, as in the aforementioned EFR contracts.
+%At the shortest timescale, BESS may receive payments from the grid operator to respond to fluctuations in the AC grid frequency, as in the aforementioned EFR contracts.
 Such frequency regulation services attract Li-ion systems designed for power, such as the Hornsdale Power Reserve project in South Australia, which is specified for \SI{1}{\hour}\SI{17}{\minute} at its \SI{100}{\mega\watt} rated power output \cite{DOEGlobalESSDatabase}.
@@
 
-An example of multi-hour energy management is the use of ESS to discharge into peak power demand in a process known as peak shaving.
-In California, where BESS are being considered as replacement for gas powered peaker plants, a project may receive capacity payments in addition to the sale of energy.
+%An example of multi-hour energy management is the use of ESS to discharge into peak power demand in a process known as peak shaving.
+In California, where BESS are being considered as replacement for gas plant, a project may receive capacity payments in addition to the sale of energy.
 However, the regulator de-rates the payment according to a 4 hour target duration, hence a \SI{1}{\mega\watt}/\SI{2}{\mega\watt\hour} BESS would only receive credit for  \SI{0.5}{\mega\watt} of power capacity \cite{NRELPeakingESS}.
 
-A similar de-rating has also been applied recently in the UK capacity market.
+%A similar de-rating has also been applied recently in the UK capacity market.
 This was done due to concerns that the average storage duration of the BESS contracted for the service was shorter than the \SI{2}{\hour} that a stress event may last for.
@@
 
-The de-rating developments would disfavour the Li-ion systems that have been deployed so far and favour systems with higher energy to power ratios.
-Redox flow battery (RFB) systems and promising candidates for multi-hour applications and revenue stacking, as the incremental cost of energy capacity is low.
+%The de-rating developments would disfavour the Li-ion systems that have been deployed so far and favour systems with higher energy to power ratios.
+Redox flow battery (RFB) systems are promising candidates for multi-hour applications and revenue stacking, as the incremental cost of energy capacity is low.
 Systems deployed so far have an average energy to power ratio of 3.5 \cite{Newbery2018}.
-
 In order to quantify the economic merits of RFB systems in comparison to other BESS, or indeed ESS, it is necessary to perform techno-economic analyses (TEA) for particular applications.
@@
 This should be done at various energy and power ratings (as these are decoupled in a non-hybrid RFB) to find the global optimum, as described by \cite{Oudalov2007}.
-The optimisation of operation requires determining the schedule, that is the power output of the battery in each sub-period, that maximises benefit without violating energy conservation.
+The optimisation of operation requires the scheduling, that is the power output of the battery in each sub-period, that maximises benefit without violating energy conservation.
 
-Where a price signal exists optimisation is commonly performed by posing a linear programming (LP) problem.
+%Where a price signal exists optimisation is commonly performed by posing a linear programming (LP) problem.
 In this method, the objective function and the constraints on the system are expressed as linear algebraic terms.
@@
 The authors also suggested that a hybrid system may be suitable, as the economically optimal VRFB operates in a narrow SOC window when providing frequency regulation, with the remaining SOC held in reserve for large frequency deviations that occur infrequently.
-
-This thread was picked up by \cite{Vaca2017} who compared a hybrid system consisting of a VRFB and a supercapacitor with the individual systems.
+\cite{Vaca2017} compared a hybrid system consisting of a VRFB and a supercapacitor with the individual systems.
 They found that the hybrid system was the optimal choice, with the VRFB providing the energy capacity and the supercapacitor the low cost power capacity.
 
-It is worth noting that both \cite{Johnston2015,Vaca2017} assume a higher cost of energy capacity than power capacity for the VRFB.
+It is worth noting that both \cite{Johnston2015} and \cite{Vaca2017} assume a higher cost of energy capacity than power capacity for the VRFB.
 This contradicts economic modelling of system costs by \cite{Viswanathan2014,Ha2015Estimatingstorage}, where the energy cost is markedly lower.
 
-\cite{Johnston2015} and \cite{Vaca2017} apply a cost of degradation in \pound per unit energy throughput as a coefficient to ESS power variables in the objective function.
+%\cite{Johnston2015} and \cite{Vaca2017} apply a cost of degradation in \pound per unit energy throughput as a coefficient to ESS power variables in the objective function.
 This approach captures an important consideration in electrical price arbitrage using ESS - that there is a threshhold price differential below which the degradation cost outweighs the revenue.
@@
 \cite{Nguyen2014} report a highly detailed non-linear model for the charge and discharge efficiency of a vanadium RFB.
-In a subsequent article (\cite{Nguyen2015}) this function was incorporated into the SOC expression of a schedule optimisation problem in the aforementioned fashion.
+In subsequent work \citep{Nguyen2015} this function was incorporated into the SOC expression of a schedule optimisation problem in the aforementioned fashion.
 Due to the complication of the SOC constraint, the authors employ a search approach with dynamic programming (DP) to break the problem down.
@@
 The importance of these improvements to the accuracy of revenue estimation is demonstrated in a day-ahead price arbitrage case study.
+
+This work is organised as follows\sol{can you describe the organisation of the paper for me i.e. 'Section 2 presents'...}
 
@@
 Under the LP formulation, the VRFB is predicted to generate \pounds \SI{0.068}{\per\kilo\watt}, whereas the NLP formulation predicts a revenue of \pounds \SI{0.063}{\per\kilo\watt}.
-In reality however, the voltaic losses that would occur under the LP schedule would reduce the revenue to \pounds 0.051.
+In reality however, the voltaic losses that would occur under the LP schedule would reduce the revenue to \pounds \SI{0.051}{\per\kilo\watt}.
 Across 2017, LP scheduling would have resulted in \pounds \SI{22.61}{\per\kilo\watt}, against a prediction of \pounds \SI{33.16}{\per\kilo\watt}, and NLP scheduling would have yielded \pounds \SI{27.99}{\per\kilo\watt}.

### 2018-11-21 Paper/main.tex
Update main.tex

@@
 This feature would also be useful for schedule optimisation of Li-ion batteries, where voltage is an important consideration.
-The introduced framework is intended to assist techno-economic optimisation of future RFB systems, for example in determining the optimum pump or cooling system size for a particular application.
-The framework allows system optimisation in terms of revenue under optimal scheduling against a price signal rather than metrics such as system cost, or lifetime cost per unit energy throughput as in previous works.
+The introduced framework may be further employed to assist in the optimal design of future RFB systems for specific applications.
 
@@
 Worldwide, the need to reduce CO$_2$ emissions is stimulating the development of a wide range of renewable energy sources.
-Presently, renewable energy is estimated to account for 10\% of global energy consumption, and this is forecast to grow to 12\% by 2023 \sol{you need a reference for this}.
+%Presently, renewable energy is estimated to account for 10\% of global energy consumption, and this is forecast to grow to 12\% by 2023 \sol{you need a reference for this}.
+
 The growth of renewable energy is forecast to be highest in the electrical power sector, from 24\% in 2017 to 30\% in 2023.
-Although hydro power presently accounts for 50\% of renewable power generation, the  intermittent power sources wind and solar are forecast to contribute most to this growth\cite{IEA2018}\sol{is this reference for all of the data before this?
+Although hydro power presently accounts for 50\% of renewable power generation, the  intermittent power sources wind and solar are forecast to contribute most to this growth\cite{IEA2018}%\sol{is this reference for all of the data before this?
 it would be better to have a couple more supporting this...}.
-Indeed, a number of large European economies, such as Germany, UK, Italy and Spain obtained 23\% to 30\% of there electrical energy from non-hydro renewable power in 2017 \cite{BP2018}.
+Indeed, in a number of large European economies, such as Germany, UK, Italy and Spain 23\% to 30\% of electrical energy was obtained from non-hydro renewable power in 2017 \cite{BP2018}.
+In the US, Texas obtained around 17\% of electrical energy supply from wind power in 2017, whereas California obtained 9\% from wind and 10\% from solar\cite{ERCOT2017,California2018}.
 
-To maintain stability of the network, the fluctuating output from wind and solar is supplemented to match demand primarily by varying output from fossil fuel power thermal plant.
+To maintain grid stability in these regions, the fluctuating output from wind and solar is supplemented to match demand primarily by varying output from fossil fuel power thermal plant.
 In the UK, this is specifically achieved by gas turbines and to a lesser extent coal plant \cite{DraxElectricInsights}.% In order to meet the strict CO$_2$ emissions targets for the UK electrical supply by 2050, and bearing in mind that nuclear power capacity is not predicted to grow until after 2030, either carbon capture or significantly greater renewable power penetration will be required at the supply side \cite{FES2018}.
 
-In economic terms, the price of electricity at a given moment is set by the operating costs and finance requirements of the least economic generator that must be called upon \cite{Staffell2016}.
-As wind and solar have zero marginal cost, increasing penetration of these power sources will reduce electrical price according to a stochastic distribution\sol{I'm not sure what this last bit means}.
+The price of electricity at a given moment is set by the operating costs and finance requirements of the least economic generator that must be called upon to supply it \cite{Staffell2016}.
+As wind and solar power have zero marginal cost, the output of such installations depresses the electrical price intermittently.
 At the same time, the capacity factors of the dispatchable fossil plant will decrease, and hence a premium is added to the operating cost in order to cover CAPEX and other costs.
 Across the EU, the operating cost of fossil generation is further increased by the required purchase of credits under the Emissions Trading System \cite{Staffell2018}.
-The overall effect of these developments would be an increase in price variability, although this has been mitigated in practice by markets for capacity, where generators may supplement income from energy sales with payments for availability of power capacity \cite{ENGIE}.\sol{can we cut this paragraph down}
+The overall effect of these developments would be an increase in price variability, although this has been mitigated in some markets by capacity payments, made on the basis of power availability rather than energy sold \cite{ENGIE}.
 
@@
 
-A full system LP schedule optimisation problem was posed by \cite{Chen2012} to assess the benefit of a VRFB in a microgrid with a load, PV, wind and dispatchable micro-generation.
+A full system LP schedule optimisation problem was posed by \cite{Chen2012} to assess the benefit of a VRFB in a micro-grid with a load, PV, wind and dispatchable micro-generation.
 When the micro-grid was isolated, the objective was to minimise the start-up and running costs of micro-generators used to back up the renewable power output.
 The authors used a mixed-integer LP (MILP) approach, where the integer variables were required to factor start-up costs.
-Additional global constraints were required to guarantee system reliability \sol{are these reliability constraints interesting?
-Yes, if one is interested in a whole system!}.
+Additional global constraints were required to guarantee reliability of the micro-grid.
 
@@
 The problem therefore becomes convex (quadratic), but non-linear in both objective function and constraint.
-The authors then employ a linear approximation technique\sol{does this need a reference?
-DR: do you mean for the technique used?}, and combine the efficiency treatment with a piece-wise linear treatment of degradation as a function of power, giving an overall mixed integer-linear programming problem (MILP).
+The authors then employ a linear approximation technique and combine the efficiency treatment with a piece-wise linear treatment of degradation as a function of power, giving an overall mixed integer-linear programming problem (MILP).
 
@@
 \begin{itemize}
-    \item Quadratic expression of $I^2R$ voltaic losses in the objective function under linear constraints.
+    \item Quadratic expression of $I^2R$ ohmic losses in the objective function under linear constraints.
     \item Open cell voltage (OCV) as a function of SOC
@@
 
-The importance of these improvements to the accuracy of revenue estimation is demonstrated in a day-ahead price arbitrage case study.
-
-This work is organised as follows\sol{can you describe the organisation of the paper for me i.e. 'Section 2 presents'...}
+The formulation of each of these elements and the corresponding treatment of the VFB are described in \cref{sec:Model_Formulation}.
+The impact of the improvements to the optimisation formulation are demonstrated in \cref{Case Study}, using the case study of a \SI{4}{\hour} rated VRFB performing arbitrage against the day-ahead N2EX electricity price.
 
@@
 In the following section, an LP formulation for schedule optimisation for maximum arbitrage revenue is first posed in current-density terms to serve as a reference.
-Then the three augmentations to the model described in \cref{sec:Intro} are introduced.
-Finally, the RFB specification process is defined.
+Then the three improvements to the model described in \cref{sec:Intro} are introduced.
+Finally, the process for specifying the RFB within the problem is defined.
 
-\section{LP schedule Optimisation in Current-Density Terms}
+\subsection{LP schedule Optimisation in Current-Density Terms}
 \label{Model_Formulation_LP_Reference_Current_Density_Terms}

### 2018-11-21 Paper/main.tex
Update main.tex

@@
 In this work a novel schedule optimisation framework is introduced in which coulombic and voltaic components are separated, and the problem variablised in terms of current-density rather than power.
-This framework allows the expression of ohmic losses as a convex quadratic function of current-density in the objective function while maintaining all linear constraints.
+This framework allows the expression of ohmic losses as a convex quadratic function of current-density in the objective function while maintaining all linear constraints (NLP).
 It is demonstrated for an arbitrage application involving a vanadium RFB that the consideration of variable voltaic losses results in a significant improvement to the accuracy of revenue estimation when compared to the typical linear programming (LP) approach where a fixed efficiency is assumed.
-The quadratic formulation is more readily solvable than alternative approaches reported in the literature, with minimal increase to solve time compared to the LP approach.
+The introduced NLP formulation is more readily computable than alternative approaches reported in the literature, with minimal increase to solve time compared to the LP approach.
 The fidelity of the RFB representation is further increased by expressing open cell voltage as a function of state of charge (SOC).
@@
 \begin{itemize}
-    \item Quadratic expression of $I^2R$ ohmic losses in the objective function under linear constraints.
+    \item Quadratic expression of ohmic losses ($I^2R$) in the objective function under linear constraints.
     \item Open cell voltage (OCV) as a function of SOC

### 2018-11-27 MILP_extension/MILP.tex
Create MILP.tex

--- 2018-11-21 Paper/main.tex
@@
 \linenumbers
-\begin{frontmatter}
 
-\title{Improving the Fidelity of Redox Flow Battery Schedule Optimisation for more Accurate Techno-economic Analyses.}
+\title{MILP for idling to minimise coulombic losses}
 
-\author{D. Roberts }
-\author[]{S. Brown \corref{cor_author}}
-\ead{s.f.brown@sheffield.ac.uk}
+\section{MILP formulation with fixed coulombic losses}
+In the NLP formulation introduced in XX, it is assumed that the coulombic efficiency of the RFB is constant, i.e.
+the coulombic losses are proportional to the useful current density.
+\textit{What I say in assumptions section of paper...}
 
-\address{Department of Chemical and Biological Engineering, The University of Sheffield, Sheffield S10 2TN}
-
-\cortext[cor_author]{Corresponding author}
-
-\begin{abstract}
-Redox Flow Battery (RFB) systems are promising technologies for the multi-hour electrical energy storage that would be necessary in a future low carbon emission  electrical grids based on wind and solar power.
-In order to perform techno-economic analyses (TEA) on these systems it is necessary to calculate potential revenue for realistic applications, ideally under optimal operation (scheduling).
-In this work a novel schedule optimisation framework is introduced in which coulombic and voltaic components are separated, and the problem variablised in terms of current-density rather than power.
-This framework allows the expression of ohmic losses as a convex quadratic function of current-density in the objective function while maintaining all linear constraints.
-It is demonstrated for an arbitrage application involving a vanadium RFB that the consideration of variable voltaic losses results in a significant improvement to the accuracy of revenue estimation when compared to the typical linear programming (LP) approach where a fixed efficiency is assumed.
-The quadratic formulation is more readily solvable than alternative approaches reported in the literature, with minimal increase to solve time compared to the LP approach.
-The fidelity of the RFB representation is further increased by expressing open cell voltage as a function of state of charge (SOC).
-At low SOC, the current-density required for a given power is greater, and hence the ohmic losses are higher.
-In a third development, a linear constraint is applied to the working cell-voltage, resulting in current-density restriction at high SOC in order to avoid hydrogen and oxygen evolution.
-This feature would also be useful for schedule optimisation of Li-ion batteries, where voltage is an important consideration.
-The introduced framework may be further employed to assist in the optimal design of future RFB systems for specific applications.
-
-\end{abstract}
-
-\begin{keyword}
-Redox Flow Battery \sep Optimisation \sep Non-linear Programming \sep Techno-Economic Analysis \sep Scheduling \sep VRFB  
-%% keywords here, in the form: keyword \sep keyword
-
-%% MSC codes here, in the form: \MSC code \sep code
-%% or \MSC[2008] code \sep code (2000 is the default)
-
-\end{keyword}
-
-\end{frontmatter}
-
-% \linenumbers
-
-%% main text
-\section{Introduction}
-\label{sec:Intro}
-Worldwide, the need to reduce CO$_2$ emissions is stimulating the development of a wide range of renewable energy sources.
-%Presently, renewable energy is estimated to account for 10\% of global energy consumption, and this is forecast to grow to 12\% by 2023 \sol{you need a reference for this}.
-
-The growth of renewable energy is forecast to be highest in the electrical power sector, from 24\% in 2017 to 30\% in 2023.
-Although hydro power presently accounts for 50\% of renewable power generation, the  intermittent power sources wind and solar are forecast to contribute most to this growth\cite{IEA2018}%\sol{is this reference for all of the data before this?
-it would be better to have a couple more supporting this...}.
-Indeed, in a number of large European economies, such as Germany, UK, Italy and Spain 23\% to 30\% of electrical energy was obtained from non-hydro renewable power in 2017 \cite{BP2018}.
-In the US, Texas obtained around 17\% of electrical energy supply from wind power in 2017, whereas California obtained 9\% from wind and 10\% from solar\cite{ERCOT2017,California2018}.
-
-To maintain grid stability in these regions, the fluctuating output from wind and solar is supplemented to match demand primarily by varying output from fossil fuel power thermal plant.
-In the UK, this is specifically achieved by gas turbines and to a lesser extent coal plant \cite{DraxElectricInsights}.% In order to meet the strict CO$_2$ emissions targets for the UK electrical supply by 2050, and bearing in mind that nuclear power capacity is not predicted to grow until after 2030, either carbon capture or significantly greater renewable power penetration will be required at the supply side \cite{FES2018}.
-
-The price of electricity at a given moment is set by the operating costs and finance requirements of the least economic generator that must be called upon to supply it \cite{Staffell2016}.
-As wind and solar power have zero marginal cost, the output of such installations depresses the electrical price intermittently.
-At the same time, the capacity factors of the dispatchable fossil plant will decrease, and hence a premium is added to the operating cost in order to cover CAPEX and other costs.
-Across the EU, the operating cost of fossil generation is further increased by the required purchase of credits under the Emissions Trading System \cite{Staffell2018}.
-The overall effect of these developments would be an increase in price variability, although this has been mitigated in some markets by capacity payments, made on the basis of power availability rather than energy sold \cite{ENGIE}.
-
-Electrical price variability presents an opportunity for arbitrage by energy storage systems (ESS), which may charge and discharge in periods of low and high price respectively.
-At present however, this revenue alone is not sufficient to justify BESS investment, even under optimistic economic assumptions \cite{Hu2010}.
-Hence an important aspect of BESS economics is revenue stacking, where multiple revenue streams are captured by the same system.
-For example by providing grid reliability services as well as performing market arbitrage.
-In the UK this is illustrated by the fact that the majority of successful BESS applicants for enhanced frequency response (EFR) contracts also have capacity market contracts \cite{NGDeratingFactors}.
-
-\cite{Sabihuddin2015} classify the services that BESS may provide by the continuous charge/discharge duration, from power quality and regulation ($<$ \SI{1}{\minute}), through bridging power (\SI{1}{\minute} to \SI{1}{\hour}) to energy management ($>$ \SI{1}{hour}).
-In order to satisfy a given duration a BESS must be specified with the corresponding energy to power ratio.
-
-%At the shortest timescale, BESS may receive payments from the grid operator to respond to fluctuations in the AC grid frequency, as in the aforementioned EFR contracts.
-Such frequency regulation services attract Li-ion systems designed for power, such as the Hornsdale Power Reserve project in South Australia, which is specified for \SI{1}{\hour}\SI{17}{\minute} at its \SI{100}{\mega\watt} rated power output \cite{DOEGlobalESSDatabase}.
-This energy to power ratio is representative of global grid-scale Li-ion installation data as collated by \cite{Newbery2018}.
-
-%An example of multi-hour energy management is the use of ESS to discharge into peak power demand in a process known as peak shaving.
-In California, where BESS are being considered as replacement for gas plant, a project may receive capacity payments in addition to the sale of energy.
-However, the regulator de-rates the payment according to a 4 hour target duration, hence a \SI{1}{\mega\watt}/\SI{2}{\mega\watt\hour} BESS would only receive credit for  \SI{0.5}{\mega\watt} of power capacity \cite{NRELPeakingESS}.
-
-%A similar de-rating has also been applied recently in the UK capacity market.
-This was done due to concerns that the average storage duration of the BESS contracted for the service was shorter than the \SI{2}{\hour} that a stress event may last for.
-\cite{NGDeratingFactors}.
-
-%The de-rating developments would disfavour the Li-ion systems that have been deployed so far and favour systems with higher energy to power ratios.
-Redox flow battery (RFB) systems are promising candidates for multi-hour applications and revenue stacking, as the incremental cost of energy capacity is low.
-Systems deployed so far have an average energy to power ratio of 3.5 \cite{Newbery2018}.
-In order to quantify the economic merits of RFB systems in comparison to other BESS, or indeed ESS, it is necessary to perform techno-economic analyses (TEA) for particular applications.
-TEA should involve optimisation of the operation of a BESS to maximise revenue (or minimise costs).
-This should be done at various energy and power ratings (as these are decoupled in a non-hybrid RFB) to find the global optimum, as described by \cite{Oudalov2007}.
-The optimisation of operation requires the scheduling, that is the power output of the battery in each sub-period, that maximises benefit without violating energy conservation.
-
-%Where a price signal exists optimisation is commonly performed by posing a linear programming (LP) problem.
-In this method, the objective function and the constraints on the system are expressed as linear algebraic terms.
-On the resulting geometric surface, the optimum is guaranteed to lie on a vertex, hence a time-consuming search of the whole surface is not required.
-
-\cite{Johnston2015} used a LP formulation to optimise the scheduling of a VRFB installed at a wind farm.
-The objective was to maximise revenue on the wholesesale electrical market and the frequency response market, while always maintaining an energy reserve in order to meet .
-By repeating this process at different energy and power ratings, the authors found that the economically optimal VRFB rating for the \SI{50}{\mega\watt} wind farm was \SI{5.3}{\mega\watt} and \SI{3}{\mega\watt\hour}.
-This supports the point made above regarding the poor economics of wholesale market arbitrage - the optimal RFB rating is dictated by the frequency market revenue and the reserve energy requirement.
-Increasing the capacity of the VRFB in order to perform additional wholesale market arbitrage is not economical.
-The authors also suggested that a hybrid system may be suitable, as the economically optimal VRFB operates in a narrow SOC window when providing frequency regulation, with the remaining SOC held in reserve for large frequency deviations that occur infrequently.
-\cite{Vaca2017} compared a hybrid system consisting of a VRFB and a supercapacitor with the individual systems.
-They found that the hybrid system was the optimal choice, with the VRFB providing the energy capacity and the supercapacitor the low cost power capacity.
-
-It is worth noting that both \cite{Johnston2015} and \cite{Vaca2017} assume a higher cost of energy capacity than power capacity for the VRFB.
-This contradicts economic modelling of system costs by \cite{Viswanathan2014,Ha2015Estimatingstorage}, where the energy cost is markedly lower.
-
-%\cite{Johnston2015} and \cite{Vaca2017} apply a cost of degradation in \pound per unit energy throughput as a coefficient to ESS power variables in the objective function.
-This approach captures an important consideration in electrical price arbitrage using ESS - that there is a threshhold price differential below which the degradation cost outweighs the revenue.
-In both works the capital cost of the ESS and the degradation cost are included, but the residual value at project end is not counted.
-\cite{Dufo-Lopez2009} present a more detailed economic model for various ESS used to increase the value of wind power sold into a day-ahead market.
-The residual value of the ESS at the end of the project is pro-rated according to the number of cycles performed as a fraction of the expected cycle life, and discounted to a net present value.
-The schedule optimisation in this work is not as thorough as the LP approaches of  \cite{Johnston2015,Vaca2017}; the day is divided into two peak periods where the ESS is discharged and two off-peak where it is charged, and the boundaries of the periods may be varied iteratively to find an apparent optimum.
-
-A full system LP schedule optimisation problem was posed by \cite{Chen2012} to assess the benefit of a VRFB in a micro-grid with a load, PV, wind and dispatchable micro-generation.
-When the micro-grid was isolated, the objective was to minimise the start-up and running costs of micro-generators used to back up the renewable power output.
-The authors used a mixed-integer LP (MILP) approach, where the integer variables were required to factor start-up costs.
-Additional global constraints were required to guarantee reliability of the micro-grid.
-
-\cite{Gomes2017}  studied the application of a VRFB in order to cope with stochastic deviations from the forecast exports of wind and solar power into a day ahead market.
-A two stage optimisation was applied, and LP was used to optimise the VRFB schedule based on the forecast  in the first stage.
-
-\cite{Hu2010} optimised the schedule of a VRFB and a poly-sulphide bromine RFB in order to maximise arbitrage revenue on the Danish day ahead electrical spot market.
-This approach was not strictly LP, but the problem may be reposed in LP form.
-
-A significant simplification in the representation of the RFB in all of the above studies is the assumption of a constant efficiency.
-In reality, the efficiency of any BESS is a function of the charge/discharge power, among other variables, such as state of charge (SOC).
-In the LP formulation, if efficiency were defined as a  function of power, it would be necessary to place this function on the denominator of the  discharge SOC increment expression.
-This would complicate the problem as the constraint is now non-linear and non-quadratic.
-\cite{Nguyen2014} report a highly detailed non-linear model for the charge and discharge efficiency of a vanadium RFB.
-In subsequent work \citep{Nguyen2015} this function was incorporated into the SOC expression of a schedule optimisation problem in the aforementioned fashion.
-Due to the complication of the SOC constraint, the authors employ a search approach with dynamic programming (DP) to break the problem down.
-This is more computationally intensive than the LP method, but the authors state that an optimum schedule may be found in an acceptable time-scale as the small number of variables and the presence of constraints limit the search space.
-A DP search approach is also used by \cite{Oudalov2007}, where the internal resistance of the BESS is expressed as a function of SOC in a TEA of a VRFB and a lead acid battery for industrial peak-shaving.
-
-\cite{Sarker2017} also incorporated the dependency of efficiency on power in a scheduling problem (although applied to a Li-ion battery rather than a RFB).
-Quadratic terms are employed to relate losses to C-rate of charge and discharge.
-The presence of this quadratic function on the denominator of the discharge SOC increment term is avoided by redefining the discharge power variable as upstream of losses (rather than the power delivered to the grid).
-The discharge efficiency function is then moved from the SOC expression and multiplied by the discharge power variable in the objective function.
-The charge efficiency is left in the SOC expression (redefining the charge power variable in the same way would require dividing it by the quadratic charge efficiency term in the objective function).
-The problem therefore becomes convex (quadratic), but non-linear in both objective function and constraint.
-The authors then employ a linear approximation technique and combine the efficiency treatment with a piece-wise linear treatment of degradation as a function of power, giving an overall mixed integer-linear programming problem (MILP).
-
-Accurate treatment of efficiency is important, as losses impact the revenue available from arbitrage.
-The key innovation of the present work is the introduction of a RFB schedule optimisation framework posed in terms of current-density rather than power.
-This allows the separation of the voltaic and coulombic components of the power input/output.
-Using the introduced framework, three improvements to the RFB representation in the schedule optimisation are posed while maintaining an easily computable problem:
-
-\begin{itemize}
-    \item Quadratic expression of $I^2R$ ohmic losses in the objective function under linear constraints.
-    \item Open cell voltage (OCV) as a function of SOC
-    \item Maximum cell voltage constraint in terms of OCV and overpotential
-\end{itemize}
-
-The formulation of each of these elements and the corresponding treatment of the VFB are described in \cref{sec:Model_Formulation}.
-The impact of the improvements to the optimisation formulation are demonstrated in \cref{Case Study}, using the case study of a \SI{4}{\hour} rated VRFB performing arbitrage against the day-ahead N2EX electricity price.
-
-\section{Model formulation}
-\label{sec:Model_Formulation}
-
-Pure price arbitrage is chosen as the application for schedule optimisation due to its simplicity, allowing a focus on the treatment of the RFB within the problem.
-In the following section, an LP formulation for schedule optimisation for maximum arbitrage revenue is first posed in current-density terms to serve as a reference.
-Then the three improvements to the model described in \cref{sec:Intro} are introduced.
-Finally, the process for specifying the RFB within the problem is defined.
-
-\subsection{LP schedule Optimisation in Current-Density Terms}
-\label{Model_Formulation_LP_Reference_Current_Density_Terms}
-
-In the literature on RFB schedule optimisation, scheduling problems are posed using either power (\cite{Chen2012,Hu2010,Nguyen2015}) or energy (\cite{Johnston2015,Vaca2017}) as the time-indexed variables.
-This is not an important difference, as the former only requires that the timestep $\tau$ be included as a coefficient.
-Here, the power convention is adopted, as it allows the timestep to be varied independently.
-The objective function for pure arbitrage revenue described by \cite{Hu2010} is adopted here, but adapted so that separate charge and discharge power variables are explicitly defined.
-Specifying a single power variable with negative values representing charge and positive discharge is not practical for LP optimisation, as the solver will drive the variable to extremes which is likely to be sub-optimal.
-
-In the period $Y$, made up of sub-periods $t$, the schedule is optimised by maximising the revenue defined by:
+In order to address this concern, the SOC at time $t$ was redefined by:
 
 \begin{equation}
-\label{eqn:Lit_Review_LP_Obj_Func}
-R = \tau \sum_{t \in Y} (P_{Dis,t} - P_{Chg,t}) p_t
+\label{eqn: Method_MINLP_SOC_Tracker}
+SOC_t = SOC_{t-1} +\frac{A.\tau}{1000C}(I_{C, t} - I_{D, t} - \delta_tl_C) \forall t \in Y
 \end{equation}
 
-where $P_{Dis,t}$ and $P_{Chg,t}$ represent the bounded variables discharge and charge power (\si{\watt}) between the RFB and the rest of the system in sub-period $t$.
-$p_t$ is the price of electrical energy (\SI[sticky-per, bracket-unit-denominator = false]{}[\pounds]{\per\mega\watt\hour}) in sub-period $t$, and $\tau$ the time-step parameter  (\si{\hour}).
+Where $\delta_t$ is a binary indicator variable with 0 indicating an idle state and 1 an active state, and $l_C$ the fixed coulombic loss term (\si{\ampere}).
+It is assumed that the volume of electrolyte in the stack is negligible compared to the volume in the reservoirs.
 
-The SOC of the RFB at time $t$ is defined after \cite{Gomes2017} by:
+The condition that the charging or discharging of the RFB requires the system to be in the active state is formalised by:
 
 \begin{equation}
-\label{eqn: Lit_Review_LP_SOC_Constraint_Example}
-\begin{gathered}
-SOC_t = SOC_{t-1} + \frac{\tau P_{Chg,t} \eta_{Chg}}{E_{BESS}} - \frac{\tau P_{Dis,t}}{E_{BESS}\eta_{Dis}} \\
-\end{gathered}
-\end{equation}
-
-where $E_{BESS}$ is the parameter energy capacity and $\eta_{Chg}$ and $\eta_{Dis}$ are the charge and discharge energy efficiency parameters respectively.
-The SOC must at all times be constrained between a maximum of 1 and a minimum of 0, though a narrower window may be used \cite{Vaca2017}.
-
-The above problem was reposed in terms of current-density by replacing the charge/discharge power variables with charge/discharge current-density variables, and multiplying the overall expression by a representative voltage in order to obtain power input/output by the classical $P = IV$ expression.
-Current-density was chosen over current as the former is a scale-independent metric commonly used in the RFB literature.
-The use of current-density makes multiplication by stack area necessary to obtain absolute current.
-Separate efficiencies representing voltaic and ohmic losses must now be specified, rather than the single energy efficiency term in \cref{eqn: Lit_Review_LP_SOC_Constraint_Example}.
-As current-density is defined based on the current at the RFB terminals, coulombic losses are dealt with upstream in \cref{eqn: Method_SOC_Tracker}.
-It is however necessary to incorporate voltaic losses in the objective function, as these affect the power input/output.
-In current-density terms, the LP objective function for arbitrage revenue (to be maximised) is defined by:
-
-\begin{equation}
-\label{eqn: Linear_Schedule_Objective_Function_Current_as_Variable}
-R_{LP} = \frac{A.\tau OCV_{50\%}}{1\times 10^{6}}\sum_{t \in Y}p_{t}(I_{D, t}\sqrt{\bar\eta_{V}}(1 - l_{BOP}) - \frac{I_{C,t}}{\sqrt{\bar\eta_{V}}(1 - l_{BOP})})
-\end{equation}
-
-where $I_{C,t}$ and $I_{D,t}$ are independent variables representing charge and discharge current density (\si{\ampere\per\square\meter}) respectively.
-$A$ is the stack area parameter (\si{\square\meter}) and $OCV_{50\%}$ the open cell voltage at 50\% SOC.
-$\bar\eta_{V}$ is the representative round-trip voltaic efficiency.
-The  placement of the additional fractional balance of plant losses parameter $l_{BOP}$ reflects the assumption that balance of plant power consumption is primarily due to electrolyte pumping, and that pumping power is proportional to current.
-The validity of this assumption is discussed in \cref{analysis_of_assumptions}.
-
-The function describing SOC given in \cref{eqn: Lit_Review_LP_SOC_Constraint_Example} is now expressable in purely coulombic terms.
-This is done by replacing the power variables with current-density variables, the energy efficiency parameters with a coulombic efficiency parameter, and the energy capacity of the RFB with a coulombic capacity.
-The resultant function for the SOC at the end of sub-period $t$ is defined by:
-
-\begin{equation}
-\label{eqn: Method_SOC_Tracker}
-SOC_t = SOC_{t-1} + \frac{A.\tau}{1000C}(I_{C,t}\sqrt{\eta_C} - \frac{I_{D,t}}{\sqrt{\eta_C}}) \; \forall t \in Y
-\end{equation}
-
-where $C$ is the coulombic capacity of the RFB (\si{\ampere\hour}).
-It is assumed that the coulombic efficiency $\eta_C$ does not vary with SOC or current-density, and that coulombic losses for charge and discharge are symmetrical.
-This assumption is discussed in \cref{analysis_of_assumptions}.
-
-The following constraints are applied to the list of SOC values generated by the above expression. Firstly the constraint on SOC range is formalised by:
-
-\begin{equation}
-\label{eqn: Method_Constraint_SOC_Range}
-SOC_{min} \leq SOC_t \leq SOC_{max} \; \forall t \in Y
-\end{equation}
-
-where $SOC_{min}$ and $SOC_{max}$ are the minimum and maximum permitted SOC parameters.
-
-So that the optimisation may be legitimately performed on consecutive periods of historical data, it is necessary that the SOC at the end of the period be returned to the starting value.
-\cite{Hu2010} set this value at 0, but in order to avoid imposing low SOC behaviour, it is here set at 0.5.
-This constraint is formalised by:
-
-\begin{equation}
-\label{eqn: Method_Constraint_E_Conservation}
-SOC_0 = SOC_n = 0.5
-\end{equation}
-where $n$ is the final sub-period in $Y$.
-
-Finally, the current-density in and out of the RFB is constrained by:
-
-\begin{equation}
-\label{eqn: Method_Constraint_Ia}
-I_{min} \leq I_{D,t}, \; I_{C,t} \leq I_{max} \; \forall t \in Y
-\end{equation}
-
-where $I_{min}$ and $I_{max}$ are the minimum and maximum permitted current-densities respectively.
-
-\subsection{NLP treatment of voltaic losses}
-\label{Model_Formulation_NLP_treatment_voltaic_losses}
-
-A major simplification in the LP problem in terms of power is the assumption of a constant energy efficiency parameter.
-This assumption also applies to the LP problem in current-density terms introduced in \cref{Model_Formulation_LP_Reference_Current_Density_Terms}, although separate voltaic and coulombic efficiencies parameters are accounted for for.
-In reality, ohmic (and pseudo-ohmic) over-potential is a linear function of current-density.
-The faradaic, or activation over-potential, which occurs when the circuit is closed, may be considered constant with respect to current-density \cite{Aaron2011}.
-\Cref{eqn: Linear_Schedule_Objective_Function_Current_as_Variable} was adapted to include these losses, by subtracting/adding the faradaic over-potential from/to the OCV during discharge/charge, and including the ohmic losses in the form of the classical $P_{Loss} = I^2R$ equation.
-This results in a non-linear (NLP) objective function for revenue in period $Y$ given by:
-
-\begin{equation}
-\label{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable}
-R_{NLP} = \frac{A.\tau}{1\times 10^{6}}\sum_{t}p_{t}(I_{D, t}(OCV_{50\%} - V_a)(1-l_{BOP}) - \frac{I_{C, t}(OCV_{50\%} + V_a)}{1-l_{BOP}} - (I_{D,t}^{2} + I_{C,t}^{2})ASR) \; \forall t \in Y
-\end{equation}
-
-where $V_a$ and $ASR$ are scalar parameters representing faradaic over-potential (\si{\volt}) and area specific resistance (\si{\ohm\square\meter}).
-The placement of $V_a$ and $ASR$ reflects an assumption of symmetry in voltaic losses between charge and discharge, which is discussed further in \cref{analysis_of_assumptions}.
-The objective function is subject to the same constraints as described in \cref{Model_Formulation_LP_Reference_Current_Density_Terms}.
-
-Although non-linear, \cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable} the objective function is convex, as it requires maximising a negative quadratic, hence a gradient based algorithmic solver is guaranteed to find the optimum from any starting point.
-
-Under the constant coulombic efficiency assumption \cref{eqn: Method_SOC_Tracker} is linear, and so by extension are the constraints in \cref{eqn: Method_Constraint_SOC_Range,eqn: Method_Constraint_E_Conservation}.
-The overall formulation hence comprises a non-linear objective function subject to linear constraints.
-This is an improvement over the formulations of losses expressed in power efficiency terms by \cite{Nguyen2015,Sarker2017}, both of which use non-linear SOC constraint expressions, as it is likely to be easier to resolve \cite{Hart2017}.
-
-The formulation also benefits from a consistent definition of the current-density variables in terms of flow between the terminals and the external system, unlike in \cite{Sarker2017}.
-
-The NLP problem is based on an assumption of constant coulombic efficiency, which is discussed further in \cref{analysis_of_assumptions}.
-
-\subsection{Functional Treatment of OCV}
-\label{Model_Formulation_Functional_Treatment_OCV}
-A simplification in the NLP problem introduced in \cref{Model_Formulation_NLP_treatment_voltaic_losses} is the use of a fixed OCV, ($OCV_{50\%}$) where OCV is in fact a function of SOC.
-If the OCV is linear with respect to SOC in the permitted SOC range, which is the case of the system described by \cite{Reed2016} it is, and the RFB transfers as much energy below 50\% SOC as above it, then this assumption may be reasonable.
-However, using the current-density formulation, it is possible to functionalise the relationship between OCV and SOC.
-The relationship between the average OCV and the average SOC in sub-period $t$ is approximated by:
-
-\begin{equation}
-\label{eqn: OCV_vs_SOC}
-OCV_t = a\frac{SOC_t + SOC_{t-1}}{2} + b 
-\end{equation}
-
-where $a$ and $b$ are system specific parameters.
-
-Replacing $OCV_{50\%}$ with $OCV_t$ in \cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable} and expanding $SOC_t$ and $SOC_{t-1}$ by \cref{eqn: Method_SOC_Tracker} results in an objective function  defined by:
-
-\begin{equation}
-    \label{eqn: NLP_true_OCV_long_form}
 \begin{split}
-\max R_{NLP} = \frac{A.\tau}{1\times 10^{6}}\sum_{t}p_{t}(\\
-(1-l_{BOP})I_{D, t}(a\frac{SOC_{t-1} + SOC_{t-2} + \frac{A.\tau}{1000C}(I_{C,t}\sqrt{\eta_C} - \frac{I_{D,t}}{\sqrt{\eta_C}} +  I_{C,t-1}\sqrt{\eta_C} - \frac{I_{D,t-1}}{\sqrt{\eta_C}})}{2} + b - V_a)\\ - \frac{1}{1-l_{BOP}}(I_{C, t}(a\frac{SOC_{t-1} +  SOC_{t-2} + \frac{A.\tau}{1000C}(I_{C,t}\sqrt{\eta_C} - \frac{I_{D,t}}{\sqrt{\eta_C}} + I_{C,t-1}\sqrt{\eta_C} - \frac{I_{D,t-1}}{\sqrt{\eta_C}})}{2} + b + V_a))\\
-- (I_{D,t}^{2} + I_{C,t}^{2})ASR) \; \forall t \in Y
+    \label{eqn: On_off_condition}
+    I_{C,t} > 0 \implies \delta_t = 1\\
+    I_{D,t} > 0 \implies \delta_t = 1
     \end{split}
@@
 
-This objective function is non-convex, due to the presence of bivariate quadratic terms such as $-I_{D,t}I_{C,t-1}$.
-
-\subsection{Cell Voltage Constraint}
-\label{Model_Formulation_Constraint_Cell_Voltage}
-It has so far been assumed that the RFB may accept the maximum charge current-density right up to the maximum permitted SOC.
-In reality this may result in the cell-voltage being too-high, discussed further in \cref{Results_Cell_Voltage_Constraint}.
-An alternative charging approach, which is common for Li-ion batteries where high cell voltages pose safety concerns, is to switch to a constant voltage taper charge at high SOC, where the current declines until a cut-off value is reached.
-In the case of an RFB, the fixed losses described above would likely make taper charging uneconomical, as demonstrated by the low efficiency at low current-density reported by \cite{Nguyen2014}.
-Rather than globally adjust the maximum permitted charge current-density or SOC, which may lead to underestimation of revenue, the objective function in \cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable} is made subject to a new linear constraint in terms of cell-voltage, defined by:
+The constraints that enforce these conditions are defined by:
 
 \begin{equation}
-\label{eqn: Cell_Voltage_Constraint}
-OCV_t + V_a + I_{C,t}ASR \leq V_{max} \; \forall t \in Y
+\begin{split}
+    \label{eqn: on_off_constraint}
+    I_{C,t} - I_{Max}\delta_{C,t} \leq 0\\ 
+    I_{D,t} - I_{Max}\delta_{D,t} \leq 0
+    \end{split}
 \end{equation}
 
-Where $V_{max}$ is the maximum permitted cell-voltage.
-As $OCV_t$ is an averaged value of SOC in the sub-period $t$ (\cref{eqn: OCV_vs_SOC}), the accuracy of the model should improve as $\tau$ is reduced.
-A similar constraint could be implemented for discharge if required.
+ after \cite{Williams2013}.
 
-\subsection{RFB Specification}
-\label{VRFB_specification}
-The RFB is modelled here as a single cell.
-As the power output of a hypothetical RFB may be increased at the cost of efficiency, it is necessary to specify a rated efficiency when determining the stack area required for, say, a \SI{1}{\kilo\watt} system.
-This is formalised by:
-
-\begin{equation}
-\label{eqn: Rated_efficiency}
-\eta_{DC} = \eta_{V,rated}\eta_{C}(1-l_{BOP})^2 = \eta_{rated}
-\end{equation}
-
-Where $\eta_V$ and $\eta_C$ are the round trip voltaic and coulombic efficiencies respectively.
-It is assumed that the one-way fractional balance of plant losses ($l_{BOP}$) are primarily due to electrolyte pumping, and that the pump runs on DC power.
-
-The stack area (\si{\square\meter}) is calculated by:
-
-\begin{equation}
-\label{eqn: Simple_RFB_Cost_Model_Stack_Area}
-A = \frac{P_{rated}}{I_{rated}V_0\sqrt{\eta_V}(1 - l_{BOP})}
-\end{equation}
-
-where $P_{rated}$ is the system power rating (\si{\watt}) and $I_{rated}$ is the current density (\si{\ampere\per\square\metre}) corresponding experimentally to coulombic and voltaic efficiencies that satisfy \Cref{eqn: Rated_efficiency}.
-
-The coulombic capacity of the VRFB, required for the tracking of SOC in \Cref{eqn: Method_SOC_Tracker}, is calculated by:
-
-\begin{equation}
-\label{eqn: coulombic_capacity}
-C = \frac{AI_{rated}r_{e:p}}{\sqrt{\eta_C}}
-\end{equation}
-
-Where $r_{e:p}$ is the specified energy to power ratio of the system.
-The presence of $\eta_C$ ensures there is sufficient chemical energy in the tank to cover coulombic losses upon discharge (square root after \cite{Darling2014}).
-
-\section{Case Study}
-\label{Case Study}
-
-\subsection{The Application}
-\label{Results_The_Application}
-Pure price arbitrage on the day ahead N2EX electrical market was chosen as a case study, as the availability of future price information makes deterministic schedule optimisation relevant as described by \cite{Hu2010}.
-Schedule optimisation was performed on 24h periods, with $\tau$ of 1h.
-The indexed price $p_t$ was parametrised using data obtained from the NordPool website \cite{NordPool2016}.
-
-\subsubsection{The RFB}
-\label{Model_Formulation_VRFB_Params}
-\cite{Reed2016} studied various electrode flow architectures on a \SI{1}{\kilo\watt} VRFB system based on a \SI{50}{\micro\meter} Nafion NR212 membrane.
-In this case study, performance data were taken from the test of the highest performing stack architecture (IDD2s) at an electrolyte flow rate of 400 \si{\ml\per\minute} per cell.
-This setup was tested at three current-densities: 1600, 2400 and 3200 \si{\ampere\per\square\meter}.
-At each test condition efficiency data were obtained as integrals over a cycle between 15\% and 85\% SOC.
-
-In the parametrisation of the VRFB for following case studies $P_{rated}$ was set at \SI{1}{\kilo\watt}, the energy to power capacity ratio $r_{e:p}$ was set at 4 and $\eta_{rated}$ was set at 75\%.
-$l_{BOP}$ was set at 0.02, based on a conservative interpretation of the pumping power data reported for a very similar system by \cite{Kim2013}.
-A coulombic efficiency ($\eta_C$) of 0.975 was derived as the average of the values reported by \cite{Reed2016} at the three current-densities tested, which only varied between 0.974 and 0.976.
-For the chosen system the overall energy efficiency requirement in \cref{eqn: Rated_efficiency} is satisfied when the voltaic efficiency $\eta_{V,rated}$ is 0.801, which corresponds (by linear interpolation) to a current-density ($I_{rated}$) of \SI{2188}{\ampere\per\square\meter}.
-$OCV_{50\%}$ was taken from \cite{Kim2011}, where the OCV of a chloride electrolyte was reported at 20\%, 50\% and 80\% SOC.
-Although a different electrolyte, based on a mixture of sulphuric acid and hydrochloric acid, was used in \cite{Reed2016} the OCV data have not been published.
-
-\cref{tab:VRFB_Sizing_Params} summarises the parameter values used to represent the VRFB.
-
-\begin{table}[!pht]
-\captionsetup{font=normalsize}
-\centering
-\caption{Parameters describing VRFB in sizing process}
-\label{tab:VRFB_Sizing_Params}
-\begin{tabular}{ccccccc}
-\hline
-Param &  Value & Units &  Source \\
-\hline
-$P_{rated}$ & 1 & \si{\kilo\watt} & Arb.\\
-$r_{e:p}$ & 4 & - & \multirow{2}{*}{Typical}\\
-$\eta_{rated}$ & 0.75 & - \\
-\hline
-$\eta_{V,rated}$ & 0.801 & - &\multirow{3}{*}{\cite{Reed2016}}\\
-$\eta_C$ & 0.975 & -\\
-$I_{rated}$ & 2188 & \si{\ampere\per\square\meter} \\
-\hline
-$OCV_{50\%}$ & 1.46 & \si{\volt} & \cite{Kim2011}\\
-$l_{BOP}$ & 0.02 & - & \cite{Kim2013}\\
-
-\hline
-\end{tabular}
-\end{table}
-
-\subsection{Implementation}
-\label{Results_Implementation}
-Each optimisation problem was posed using PYOMO within a Python loop for repeat solutions on 24h sets of data.
-The gurobi solver was called via the SolverFactory module.
-The script was run on a virtual machine (\cite{VirtualBox}) running Linux Mint, which was allocated 1024 MB RAM.
-
-\subsection{NLP treatment of Voltaic Losses}
-\label{Results_NLP_Treatment_Voltaic_losses}
-In the first case study, the performance of the NLP formulation with functionalised voltaic losses (\cref{Model_Formulation_NLP_treatment_voltaic_losses}) was assessed by comparison to the LP formulation described in \cref{Model_Formulation_LP_Reference_Current_Density_Terms}.
-The LP objective function in \cref{eqn: Linear_Schedule_Objective_Function_Current_as_Variable} requires the assumption of a constant voltaic efficiency term $\bar\eta_{V}$.
-A value of 0.842 was applied, as this is the voltaic efficiency at the mid-point of the permitted current-density range (\SI{1600}{\ampere\per\square\meter}).
-Similarly, the NLP formulation in \cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable} requires values for area specific resistance ($ASR$) and Faradaic potential ($V_a$) in order to parametrise voltaic losses in the objective function.
-These were derived by multiplying $OCV_{50\%}$ by $1-\sqrt{\eta_V}$ to obtain an over-potential $\delta_v$ at the three current-density points in the experimental data, assuming losses are symmetrical in charge and discharge.
-$ASR$ was obtained as the gradient of the linear regression on this plot, and $V_a$ as the intercept at zero current-density.
-
-\cref{fig:Va_ASR_Derivation} shows the regression of $\delta_V$ against current density for the data taken from \cite{Reed2016}.
+ The expression of coulombic losses as a fixed term results in the restriction of operation in the low current-density range, where the coulombic losses would cancel out the voltaic loss reductions.
+\cref{fig:MINLP_on_EffC} shows a comparison of the optimal schedules obtained using the NLP formulation in XX and the MINLP formulation.
 
@@
 \centering
-\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./ASR_deriv.pdf}
-\caption{Derivation of $ASR$ (gradient) and $V_a$ (intercept) for IDD2s VRFB embodiment reported by \cite{Reed2016} at 400 \si{\ml\per\minute}.}
-\label{fig:Va_ASR_Derivation}
+\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./figures/MINLP_schedule.pdf}
+\caption{Optimal schedules obtained by NLP and MINLP optimisation for N2EX electrical price on 22$^{nd}$ August 2017.}
+\label{fig:MINLP_on_EffC}
 \end{figure}
 
-The optimisation of the scheduling was performed for each day in 2017 using both the NLP formulation and the LP benchmark.
-\Cref{fig:IDD2s_opt_schedules} shows the optimal hourly schedule determined by each method for the 22\textsuperscript{nd} of August, this date was chosen because the resultant revenue is the median of the range obtained across the year.
-
-\begin{figure}[!ht]
-\centering
-\subfloat[][]{
-\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./figures/V_losses_schedules.pdf}
-\label{fig:IDD2s_opt_schedules}
-}
-\subfloat[][]{
-\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./figures/V_losses_SOC.pdf}
-\label{fig:IDD2s_opt_schedules}
-}
-\caption{(a) Optimal VRFB schedules derived for the 22\textsuperscript{nd} August 2017 using LP and NLP optimisation formulations (discharge plotted as negative). (b) corresponding SOC values.}
-\end{figure}
-
-Both the LP and NLP solutions involve charging the VRFB when the electrical price is low (e.g.
-05:00 to 07:00) and discharging it when the price is high (e.g.
-10:00 to 13:00).
-The LP optimal profile involves discharging or charging the system at high current-density in the four hours with the most extreme prices.
-The NLP optimal schedule also targets the most extreme periods, but spreads the charging/discharging out.
-This is the consequence of factoring variable ohmic losses in the objective function, where high current-density is penalised by higher losses.
-Under the LP formulation, the VRFB is predicted to generate \pounds \SI{0.068}{\per\kilo\watt}, whereas the NLP formulation predicts a revenue of \pounds \SI{0.063}{\per\kilo\watt}.
-In reality however, the voltaic losses that would occur under the LP schedule would reduce the revenue to \pounds \SI{0.051}{\per\kilo\watt}.
-Across 2017, LP scheduling would have resulted in \pounds \SI{22.61}{\per\kilo\watt}, against a prediction of \pounds \SI{33.16}{\per\kilo\watt}, and NLP scheduling would have yielded \pounds \SI{27.99}{\per\kilo\watt}.
-
-In a techno-economic analysis of this VRFB, use of the LP formulation would therefore lead to errors in calculation of revenue, in this case an overestimation.
-Use of the NLP formulation to determine scheduling in a practical application would result in the prediction of significantly higher revenue than would the LP formulation.
-It would be possible to set a representative voltaic efficiency ($\bar\eta_V$) in the LP formulation such that the estimate of revenue would be closer to that of the NLP formulation, but this value cannot be known \textit{a priori}.
-
-Importantly, the increased fidelity in the formulation requires a minimal increase to solution time; in this case, solving  \cref{eqn: Linear_Schedule_Objective_Function_Current_as_Variable} for each day in 2017 takes \SI{129}{\second} versus the \SI{127}{\second} required to solve the LP objective function in \cref{eqn: Linear_Schedule_Objective_Function_Current_as_Variable}.
-
-\subsection{NLP with functional OCV}
-\label{Results_NLP_Functional_OCV}
-In the second case study, the impact of expressing OCV as a function of SOC, and hence of the current-density variables in preceding sub-periods, was investigated.
-Parameters $a$ and $b$ representing the linear relationship between OCV and SOC in \cref{eqn: NLP_true_OCV_long_form} were set at 0.2667 and 1.333, after linear regression on data published by \cite{Kim2011}.
-
-\cref{fig:OCV_fixed_dynamic_schedules} shows a comparison of the optimal schedules obtained using the functionalised OCV formulation and the previous formulation (\cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable}) where OCV is fixed at the 50\% SOC value ($OCV_{50\%}$).
-\cref{fig:OCV} shows the OCV in the former solution.
-
-\begin{figure}[!ht]
-\centering
-\subfloat[][]{
-\label{fig:OCV_fixed_dynamic_schedules}
-\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./figures/OCV_fixed_dynamic_schedules.pdf}
-
-}
-\subfloat[][]{
-\label{fig:OCV}
-\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./figures/OCV_fixed_dynamic_OCV.pdf}
-}
-\caption{(a) Optimal VRFB schedules derived for the 22\textsuperscript{nd} August 2017 using NLP formulations where OCV is fixed as $OCV_{50\%}$ and functionalised.
-(b)Optimal values of OCV where it is functionalised.}
-\end{figure}
-
-The performance of arbitrage between 16:00 and 20:00 is notably reduced in \cref{fig:OCV}.
-In this period there are small fluctuations in price from which the RFB derived revenue when a fixed OCV was assumed.
-However the global optimum requires that the RFB be at a low SOC in these hours, and hence at a low OCV.
-The low OCV means that a higher current-density would be required to obtain a given pwoer outpute, and hence higher voltaic losses would be incurred, eroding the revenue.
-On this date the impact on revenue is low; the optimal schedule with consideration of functional OCV results in a revenue of \pounds \SI{0.062}{\per\kilo\watt} compared to \pounds \SI{0.063}{\per\kilo\watt} for the base case.
-Across 2017 the formulation with functional OCV returns \pounds \SI{26.40}{\per\kilo\watt} compared to \pounds \SI{27.81}{\per\kilo\watt} for the fixed OCV case.
-The assumption of a fixed OCV (parametrised at 50\% SOC) therefore results in a small but not insignificant overestimation of revenue in this application.
-Including functionalised OCV in the objective function increases the solve time - the 2017 data were processed in \SI{217}{\second} compared to \SI{129}{\second} for the convex objective function in \cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable}.
-
-%The non-convex objective function in \cref{eqn: NLP_true_OCV_long_form} makes the formulation less robust: in the 2017 data there were two days in which the price parameter set yielded problem matrices that required greater diagonal adjustment than the default maximum permitted value of 1e-6.
-In these instances Gurobi returns an error message describing the diagonal adjustment that would be required.
-Solutions for the two days were obtained by increasing the Gurobi Parameter "PSDTol" to 1.7e-6.
-
-\subsection{NLP with functional OCV and Cell-Voltage constraint}
-\label{Results_Cell_Voltage_Constraint}
-In the NLP formulations given in \cref{Model_Formulation_NLP_treatment_voltaic_losses,Model_Formulation_Functional_Treatment_OCV} it is assumed that the RFB may be charged at maximum current-density at the maximum SOC.
-This was the charging strategy applied by \cite{Reed2016}, where the cell-voltage was allowed to reach \SI{1.85}{\volt} at the highest current-density, corresponding to an approximate SOC of 0.85.
-Overcharging may lead to oxygen evolution at the cathode and hydrogen evolution at the anode.
-This results in reduced coulombic efficiency and, more importantly, the evolved gases can cause a range of problems as described by \cite{Kear2012}.
-The particular cell-voltage at which gas evolution becomes problematic will vary from system to system as it depends on the particular electrolyte chemistry.
-
-With the introduction of the functional OCV in \cref{eqn: NLP_true_OCV_long_form}, the application of a maximum cell-voltage constraint is now possible.
-In this case study $V_{max}$ in \cref{eqn: Cell_Voltage_Constraint} was set at 1.65, just below the voltage at which \cite{Wei2017} reported evolution of macroscopic bubbles of hydrogen and oxygen (\SI{1.70}{\volt}).
-
-Maximising \cref{eqn: NLP_true_OCV_long_form} subject to \cref{eqn: Cell_Voltage_Constraint} for the 2017 data resulted in \pounds \SI{26.07}{\per\kilo\watt} compared to \pounds \SI{26.40}{\per\kilo\watt} when the cell voltage was not constrained.
-The difference is small because high current-density is already discouraged by the voltaic losses in \cref{eqn: NLP_true_OCV_long_form}.
-
-\cref{fig:V_max_schedules} shows the optimal schedules with and without constraint for 5\textsuperscript{th} February 2017, where the difference in revenue was greatest (\pounds \SI{0.18}{\per\kilo\watt} vs.
-\pounds \SI{0.20}{\per\kilo\watt}).
-\cref{fig:V_max_voltages} shows the corresponding cell-voltage profiles.
-
-\begin{figure}[!ht]
-\centering
-\subfloat[][]{
-\label{fig:V_max_schedules}
-\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./figures/V_max_schedules.pdf}
-}
-\subfloat[][]{
-\label{fig:V_max_voltages}
-\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./figures/V_max_voltages.pdf}
-}
-\caption{(a) Optimal VRFB schedules derived for the 5\textsuperscript{th} February 2017 by maximising \cref{eqn: NLP_true_OCV_long_form} with no cell-voltage constraint, and subject to the cell voltage constraint given in \cref{eqn: Cell_Voltage_Constraint}.
-(b) Corresponding cell-voltage profiles.}
-\end{figure}
-
-In the unconstrained case the RFB charges at the maximum current-density of \SI{3200}{\ampere\per\square\meter} in hour 20 in order to capitalise on the unusually low electrical price, resulting in a cell-voltage of \SI{1.71}{\volt}.
-In the constrained case, the current-density is restricted to \SI{2213}{\ampere\per\square\meter} in the same period, and the RFB must charge more (and discharge less) in other periods to satisfy \cref{eqn: Method_Constraint_E_Conservation}.
-The impact of the constraint would be more severe if the variations in price data were greater, or the maximum voltage set at a lower value.
-
-The application of the linear constraint on cell-voltage increases the time to process the 2017 data from \SI{217}{\second} to \SI{232}{\second}.
-\subsection{Impact of price profile on optimisation results}
-\label{Results_impact_of_price_profile}
-An important aspect of RFB economics is the variability of daily price profile across the year.
-\Cref{fig:DJF_2017_N2EX,fig:JJA_2017_N2EX} show composite day-ahead price data for the winter and summer seasons in 2017.
-In summer months, the availability of cheap solar power depresses the price around midday, and the peak evening demand is low, resulting in two peaks of similar height.
-In winter, the solar price depression is not so marked, and the price in the evening peak becomes higher, hence the profile is more like a single peak.
-
-\begin{figure}[!ht]
-\centering
-\subfloat[][]{
-\label{fig:JJA_2017_N2EX}
-\includegraphics[trim = 13.8cm 3.5cm 10.3cm 4.5cm, clip, width = .4\textwidth]{./figures/JJA_2017.pdf}
-}
-\subfloat[][]{
-\label{fig:DJF_2017_N2EX}
-\includegraphics[trim = 13.8cm 3.5cm 10.3cm 4.5cm, clip, width = .4\textwidth]{./figures/DJF_2017.pdf}
-}
-\caption{Composite hourly N2EX day-ahead price data for a) June, July and August 2017 and b) December,January, February 2017}
-\end{figure}
-
-\cref{fig:winter_day_schedules,fig:summer_day_schedules} show the optimal LP and NLP (\cref{Model_Formulation_Constraint_Cell_Voltage}) schedules for example winter and summer days from 2017.
-Although the revenue obtained is \pounds \SI{0.074}{\per\kilo\watt} for both days under the NLP optimal schedule, the throughput of the RFB is greater on the summer day.
-This has an impact on the error of the LP revenue estimation; for the winter day the revenue is estimated at \pounds \SI{0.088}{\per\kilo\watt}, whereas on the summmer day it is estimated at \pounds \SI{0.099}{\per\kilo\watt}.
-The error in the estimation of revenue using the LP optimisation is therefore dependent on the application.
-
-\begin{figure}[!ht]
-\centering
-\subfloat[][]{
-\label{fig:winter_day_schedules}
-\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./figures/winter_day_schedules.pdf}
-}
-\subfloat[][]{
-\label{fig:summer_day_schedules}
-\includegraphics[trim = 2cm 8cm 2cm 9cm, clip, width = .5\textwidth]{./figures/summer_day_schedules.pdf}
-}
-\caption{Optimal schedules by LP and NLP optimisation for (a) Winter day with pronounced evening price peak (b) Summer day with afternoon price depression associated with solar power}
-\end{figure}
-
-\subsection{Analysis of Assumptions}
-\label{analysis_of_assumptions}
-A number of assumptions have been made in the preceding NLP formulations in current-density terms:
-
-\begin{enumerate}
-
-    \item It is assumed that both coulombic efficiency and the voltaic loss parameters are invariant with SOC and symmetrical, i.e.
-the same for charge and discharge.
-Regarding the first assumption, experimental work by \cite{Nguyen2014} showed that the voltaic discharge efficiency is lower at 70\% SOC than at 35\% SOC, but that the difference decreases as power increases, although data are only reported up to \SI{1.8}{\kilo\watt} for a \SI{5}{\kilo\watt} system.
-Regarding the latter, the same authors provide experimental evidence that the voltaic efficiency of the particular VRFB is higher for charge than discharge.
-The latter issue could be dealt with in the present formulation, if experimental data were made available, by including applying separate $ASR$, $V_a$ and $\eta_{C}$ terms for charge and discharge.
-The former issue would be more difficult to deal with, and a first step would be to perform a sensitivity study based on relevant system data.
-
-    \item Coulombic efficiency is assumed to be constant with current-density, i.e.
-the coulombic losses are proportional to the useful current-density.
-This assumption is supported by the data reported by \cite{Reed2016}, but the behaviour below \SI{1600}{\ampere\per\square\meter} was not reported.
-In principle shunt current losses are proportional to cell-voltage, as described by \cite{Xing2011}, and hence a function of SOC and over-voltage (via \cref{eqn: Cell_Voltage_Constraint}.
-The dependence of coulombic losses on current-density is therefore weaker than implied by the present assumption.
-As a first approximation, a fixed loss term could be placed in the SOC expression in \cref{eqn: Method_SOC_Tracker} instead of the $\sqrt{\eta_C}$ coefficients.
-
-    \item Balance of plant losses are fixed as a proportion of current-density in \Cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable}, reflecting an assumption of a continuously variable pump.
-This neglects the increase in pressure drop with electrolyte flow (shown in \cite{Reed2016}).
-The formulation may be improved by removing the $(1-l_{BOP})$ terms and adding a quadratic function of current-density.
-Alternatively the pump may have a  fixed output and be sized such that the maximum charge/discharge current-density may be satisfied at the highest/lowest SOC without incurring a significant additional over-potential due to mass-transport limitations.
-
-    \item The RFB is not thermally constrained.
-The VRFB based on a sulphate electrolyte is liable to precipitation of  V\textsubscript{2}O\textsubscript{5} from the catholyte as temperature increases \citep{Rahman2009,Vijayakumar2011}.
-Although the chloride and mixed acid electrolytes developed at PNNL are more stable \citep{Li2011,Kim2011}, temperature remains an important consideration.
-The framework introduced here could easily be expanded to capture the thermal performance of the system.
-A thermal constraint could be put in place using a similar construct to the one describing SOC, where the increase of heat in the electrolyte in sub-period $t$ is equal to the ohmic losses ($(I_{D,t}^{2} + I_{C,t}^{2})ASR$).
-This would allow cost benefit analyses of active cooling versus restrictions on operation at different electrolyte concentrations.
-
-\end{enumerate}
-
-As discussed above, both the coulombic losses and the pumping power may be better represented as fixed terms.
-In this case, the efficiency of the system at low power input/output will be unfavourable as demonstrated by \cite{Nguyen2014}.
-An issue with representing these losses as fixed terms, is that they would be present at all times in the schedule.
-In order to minimise parasitic losses, a binary variable could be added to the formulation in order to allow idling.
-
-\section{Conclusion} 
-\label{sec:Conclusion}
-A novel framework for RFB schedule optimisation has been introduced in which the voltaic and coulombic components of the power output/input are separated, with the problem variablised in terms of current-density.
-An arbitrage case-study is used to demonstrate the increased fidelity that is achievable using the framework.
-
-The framework is first used to describe ohmic losses as a quadratic function of current-density in the objective function, hence maintaining a convex solution space and all linear constraints.
-The importance of inclusion of variable ohmic losses in the arbitrage case study if demonstrated.
-Optimal deterministic scheduling using the introduced formulation (NLP) resulted in 24\% greater revenue across 2017 when compared to a basic LP approach that assumes a constant representative voltaic efficiency.
-In the context of a TEA, failure to factor in the variable voltaic losses leads to an overestimation of revenue by 47\%.
-The improved NLP treatment of voltaic losses only requires a minimal increase in computing time when compared to the LP problem with fixed voltaic efficiency.
-
-The framework also facilitates the expression of OCV as a function of SOC, rather than assuming a fixed average value.
-The non-convex objective function that results may be solved using Gurobi, although there is a penalty in computing time.
-For the arbitrage case study, it is demonstrated that assuming a fixed OCV results in a small but significant overestimation of revenue.
-This assumption should be checked for other applications.
-
-The fidelity of the RFB representation is further increased by placing a linear constraint on cell-voltage as a function of SOC and over-potential.
-This constraint restricts the current-density at which the battery may be charged when the SOC is high, which is an important consideration for safety and battery lifetime.
-In the present case study, the impact of the chosen maximum voltage is low, as the previous factoring of voltaic losses already discourages high power input.
-However, in applications with greater price variation this constraint will become more important.
-
-A number of improvements to the particular problems presented are proposed, in order to more faithfully represent coulombic losses, the parasitic consumption of power in electrolyte pumping, and to allow idling.
-
-It is intended that the novel framework be used for techno-economic analysis and optimisation of various RFB chemistries at a system level (i.e.
-up to the AC-DC boundary), for example in the optimal specification of pump rating, or cooling system for a particular application.
-RFB systems have previously been the subject of technical optimisation, for example the study of the trade-offs between electrolyte concentration and ASR \citep{Weber2013}, or shunt currents and pumping losses \citep{Viswanathan2014}.
-In these works, a simple metric such as levelised cost of electrical energy under a simple duty cycle is used as the objective to minimise.
-The framework introduced here allows optimisation in terms of revenue under optimal scheduling for any application where a price signal exists.
-
-Although the framework was conceived with RFB systems in mind, the easy application of a cell-voltage constraint would be useful for Li-ion scheduling, where low and high cell-voltages increase cell degradation, and the latter also poses a safety concern.
-In the case of Li-ion batteries, where coulombic losses are negligible, the present assumption of a fixed coulombic efficiency would be entirely appropriate.
-The introduced treatment of voltaic losses is an improvement over that of \cite{Sarker2017} as the constraints are all linear.
-
-\section{Acknowledgements}
-
-This work was supported by the UK Engineering and Physical Science Research Council via grant EP/L016818/1 which funds the Centre for Doctoral Training in Energy Storage and its Applications.
-The authors are also grateful to Drax Power Ltd.
-for their financial support.
-
-\section*{References}
-\bibliographystyle{elsarticle-harv}
-\bibliography{Elsevier_article.bib}
-
+A similar approach may be employed to deal with other fixed losses, such as the pumping power loss for a constant output pump.
 \end{document}
